# Digital road infrastructure and connectivity {#digital}

## Vehicle to everything communication

### Synonyms {-}
*Connected Vehicle (CV), Connected Vehicle technologies (CVT), Vehicle-to-x (car and infrastructure) (C2x/V2x), Cooperative Intelligent Transport Systems (C-ITS), Cellular-V2X technology (C-V2X)*

### Definition {-}
Through Intelligent Transportation Systems (ITS), various Connected Vehicle (CV) technologies have been developed in recent years to contribute to safer roads through cooperative situational awareness and hazard avoidance. Two main types of communication have been proposed: vehicle-to-vehicle (V2V) and vehicle-to-infrastructure (V2I) communication (Outay et al., 2019). 
C2X (car to everything) or more broadly V2X (vehicle to everything) is the new technology that enables both communication between vehicles (car-to-car) and information exchange with infrastructure (car-to-infrastructure) (ADAC, 2021). <br/><br/>
V2V offers benefits in terms of increased safety, as it can prevent accidents by allowing a vehicle to exchange real-time information about speed, location and direction with other vehicles in the surrounding area. In addition to their safety applications, V2V and V2I communications can potentially help reduce fuel consumption and emissions due to the fact that excessive pollutant emissions are often associated with heavy braking, changing traveling speeds and acceleration/deceleration, especially at signalised intersections. In the context of smart cities, many researchers are exploring the potential use of connected vehicles to support eco-friendly driving by reducing CO~2~ emissions. This is often achieved through vehicle-to-vehicle (V2V) and vehicle-to-roadside unit RSU (V2I) inter-connectivity to harmonise vehicle speeds by maintaining traffic flow and reducing unnecessary stops and starts (Outay et al., 2019).<br/><br/>
In terms of the communication technology used for V2X, according to Dey et al. (2016) Dedicated Short Range Communication (DSRC) was considered the primary option for CVT security applications in 2016. However, the use of other radio technologies such as Wi-Fi, LTE or WiMAX enables communication with greater range and provides higher throughput requirements that could not be supported by DSRC alone. In addition, the use of other radio technologies potentially reduces the need for costly DSRC infrastructure.


### Key stakeholders {-}

- **Affected**: Car drivers, Insurers
- **Responsible**: National Governments, Private Companies, Car Manufacturers, Infrastructure operators

### Current state of art in research {-}
Many research papers focus on the technical performance of this technology, the comparison of V2V with V2I and the mixing of V2V vehicles with non-V2V equipped vehicles. As in 2019, the idea of combining V2V and V2I communications into a hybrid V2X alert system has already became reality (Outay et al., 2019).<br/><br/>
Moreover, research is conducted on the comparison of available communication technology to ensure the fastest, most efficient and error-free solution. Currently, there are two possibilities under discussion for Car2X communication. Firstly, the IEEE 802.11p Dedicated Short Range Communication (DSRC) and secondly, the Cellular-V2X technology (C-V2X). The former is based on the IEEE 802.11 WiFi standard, while C-V2X is based on 4G LTE, with a roadmap towards 5G C-V-to-X. While China and USA primarily rely on C-V2X, Europe is still undecided as to whether car networking should take place via pWLAN or via C-V2X. This creates an international confusion of languages. As a result, vehicles may not be able to communicate without errors because they use different languages (Köllner, 2020).<br/><br/>
IEEE 802.11p is technically very advanced and operates with minimal latencies. But the advantage of 5GAA technology is that 5G is to be introduced globally on a massive scale in the near future and the fast-mobile radio standard will be in the car anyway, if only to transmit the huge amounts of data generated by autonomous driving. Moreover, in a dense 5G network, far fewer of the 5.9 GHz units needed for direct communication have to be integrated into the infrastructure than with the 802.11p solution. Some of the tasks of these units will then be taken over by the 5G network (Knecht, 2018).

### Current state of art in practice {-}
So far, C2X is only available from some German manufacturers. Only Ford models, the Volkswagen Golf 8 and all Volvo models always have C2X on board as standard. This is desirable from the point of view of consumer protection, because more cars are equipped with C2X, more accurate this system is in preventing accidents. Again, only Golf and Volvo models have C2X free of charge after purchase. All other manufacturers charge money for C2X after one to three years. Moreover, C2X is never available on its own, but only in a package with other Connect services that do not necessarily have anything to do with road safety (ADAC, 2021).<br/><br/>
Several important technical specificities should be considered for V2V communication:

-	The time it takes for a warning to reach another car. This varies greatly between manufacturers, ranging from 0.1 seconds to 2 minutes - with the latter value being far too slow for many situations (e.g. the end of a traffic jam behind a bend). 
-	Various transmission techniques currently prevent cars from all manufacturers from warning each other.
-	With many manufacturers, warnings cannot be passed on if the car is in a cellular network dead zone.
-	The number of dangerous situations that are warned about varies from manufacturer to manufacturer.

In the table below, we compare how the warnings specification differ between three car manufactures: 

```{r table46, echo=FALSE, results='asis'}
cat('|Audi|Ford|Mercedes|
|:----:|:----:|:----:|
|broken-down vehicles|broken-down vehicles|breakdown|
|accidents|general traffic warning|accident|
|end of traffic jam|end of traffic jam|-|
|fog, black ice|dangerous road conditions (icy, heavy rain, oil, etc.) |heavy rain, fog,crosswind and icy roads|
|online traffic sign information|-|hazard warning lights switched on|
|-|-|additional hazard manually reported by the driver through the navigation menu|
|display of the probability of free parking spaces along roads incl. additional information such as prices|-|-|
|speed recommendation to reach the next traffic light in a green phase|-|-|
|-|road works|-|
|-|objects, animas, people on the road|-|
|-|wrong-way drivers|-|')
```

Moreover, the ADAC's recommendations to the manufacturers are the following (ADAC, 2021):

-	The manufacturers should quickly agree on a transmission technology.
-	C2X should quickly become standard equipment
-	Safety-relevant C2X functions should not cause any follow-up costs

A basis for cooperative systems is currently being established in Europe. The procedures for testing under real traffic conditions are being defined and coordinated among the partners involved. A large part of the technical solutions for data communication is standardised. The non-technical aspects (e.g. organisational structures, safety concept) are currently being worked out in preparation for the market launch in a public-private partnership.<br/><br/>
On this basis, German, Dutch and Austrian road operators, together with partners from industry, are starting the step-by-step introduction of cooperative systems in Europe within the framework of the C-ITS corridor from Rotterdam to Frankfurt am Main and Vienna
(Cooperative ITS Corridor, no date). <br/><br/>
In Austria, with the award of a comprehensive framework contract, ASFINAG has now become the first infrastructure provider in Europe to reach a further milestone in networking vehicles and roads. The total volume of the framework contract is 14.5 million euros. This will make it possible to equip the entire motorway network in Austria with C-ITS in the coming years.The equipment, which will be installed step by step along the motorways starting in November, includes up to 525 so-called road units as well as a control centre. The first C-ITS services for hazard warning are expected to go into operation within the next 16 months. Further expansion will focus on the support of automated driving and networked traffic management. The C-ITS equipment is part of the digitalisation of the road infrastructure and is funded by the Climate and Energy Fund and the EU (Močnik, 2020).

### Relevant initiatives in Austria {-}

- [infothek.bmk.gv.at](https://infothek.bmk.gv.at/fahrer-assistenzsysteme-verkehrssicherheit-vernetzung/) 
- [c-its-korridor.de](https://c-its-korridor.de/?menuId=1&sp=en) 
- [asfinag.at](https://www.asfinag.at/ueber-uns/newsroom/pressemeldungen/2020/wlan-ausbau-cooperative-intelligent-transport-systems/) 
- [kununu.com](https://www.kununu.com/de/automotive-safety-technologies/news/car2x-projekt-in-oesterreich-praesentiert) 
- [hitech.at](https://www.hitech.at/mobilitaet/wohin-geht-die-fahrt) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table44, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Improvement of road safety|**+**|Health & Wellbeing (*3*)|Filippi et al., 2016|
|Individual|V2X communications will come at no cost to the end user|**+**|Sustainable economic development (*8,11*)|Hainen et al., 2019|
|Systemic|Emissions reduced|**+**|Environmental sustainability (*7,12,13,15*)|Outay et al., 2019|
|Systemic|Increased efficiency of transport systems|**+**|Sustainable economic development (*8,11*)|Filippi et al., 2016|')
```

### Technology and societal readiness level {-}

```{r table45, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|6-8|5-7|')
```

### Open questions {-}
1. Which combination of the different communication options is the best?
2. Which communication technology is the most suitable for Europe?
3. Are infrastructure operators already taking care of making data internationally compatible so that cars can communicate with it?


### Further links {-}

- [c-its-korridor.de](https://c-its-korridor.de/?menuId=1&sp=en) 
-	[nhtsa.gov](https://www.nhtsa.gov/technology-innovation/vehicle-vehicle-communication) 

### References {-}
- ADAC (2021) ‘Welche Hersteller bieten bereits C2X an ? Datenquelle Original-Rückmeldungen’.
- Cooperative ITS Corridor (no date) Cooperative ITS Corridor. Available at: https://c-its-korridor.de/?menuId=1&sp=de (Accessed: 18 February 2021).
- Dey, K. C. et al. (2016) ‘Vehicle-to-vehicle (V2V) and vehicle-to-infrastructure (V2I) communication in a heterogeneous wireless network - Performance evaluation’, Transportation Research Part C: Emerging Technologies. Elsevier Ltd, 68, pp. 168–184. doi: 10.1016/j.trc.2016.03.008.
- Filippi, A. et al. (2016) ‘Ready to roll: Why 802.11p beats LTE and 5G for V2x’, pp. 1–14. Available at: https://www.siemens.com/content/dam/webassetpool/mam/tag-siemens-com/smdb/mobility/road/connected-mobility-solutions/documents/its-g5-ready-to-roll-en.pdf.
- Hainen, A. et al. (2019) ‘Co-Deployment of DSRC Radio and Cellular Connected Vehicle Technology in Tuscaloosa, AL and Northport, AL’, pp. 1–9.
- Knecht, J. (2018) C-V2X Europapremiere: Wie Autos sprechen – mit wem auch immer | AUTO MOTOR UND SPORT. Available at: https://www.auto-motor-und-sport.de/technik/vernetzung-cv2x-car-to-car-europapremiere/ (Accessed: 15 February 2021).
- Köllner, C. (2020) Car-to-X | Fahrzeugvernetzung per C-V2X oder pWLAN? | springerprofessional.de. Available at: https://www.springerprofessional.de/car-to-x/automatisiertes-fahren/fahrzeugvernetzung-per-c-v2x-oder-pwlan-/17822610 (Accessed: 15 February 2021).
- Močnik, W. (2020) ASFINAG startet als erster Autobahnbetreiber Europas Vernetzung von Straße und Fahrzeug. Available at: https://www.asfinag.at/ueber-uns/newsroom/pressemeldungen/2020/wlan-ausbau-cooperative-intelligent-transport-systems/ (Accessed: 18 February 2021).
- Outay, F. et al. (2019) ‘V2V and V2I communications for traffic safety and CO2 emission reduction: A performance evaluation’, Procedia Computer Science. Elsevier B.V., 151(2018), pp. 353–360. doi: 10.1016/j.procs.2019.04.049.


## Infrastructure support levels for automated driving 

## Wireless communication