# Passenger information system {#passenger}

## Digital journey planner {#djp}

**Updated: 17th May 2022**

### Synonyms {-}
*DJP, Personal travel planner*

### Definition {-}
Digital Journey Planner (DJP) is an app, which based on the travellers' preferences, recommends the fastest and most convenient route to get from point A to point B including different means of transport such as public transport, cycling, walking or car-sharing. Digital Journey Planner provides diverse range of features such as individual or collective transport, scheduled time, availability, numbers of passengers or length of the route. Digital Journey Planner apps are developed and sponsored by public authorities as well as private stakeholders in order to increase sustainable urban mobility, intensify the model of “smart city” in metropolises with an ultimate aim to improve the quality of life in the cities. 
There are some commonly used Digital Journey Planners available worldwide such as [*GoogleMaps.com*](https://www.google.com/maps) or [*mylike*](https://mylike.io/personal-travel-planner/). Furthermore, it is crucial to mention that all of these apps work on the basis of data importer and validator, which could be divided into two categories such as (Jakob et al., 2014): 

-	Databases of roads, footpaths, cycle lanes, which are collected in order to plan routes in a safe and effective way, for both cyclist, car-users and passengers in general;
-	Databases of schedules of public transport and public transport’s stops, which are collected to implement journeys by means of public transport services. 

The function phases of Digital Journey Planners can be categorized into the following phases:

-	Pre-trip phase
-	On-trip phase
-	Post-trip phase


### Key stakeholders {-}
- **Affected**: Digital journey planner users, Travellers
- **Responsible**: Public transport authorities, Private developers, Transport service operators 

### Current state of art in research {-}
Firstly, Liebig et al. (2014) conducted the research, which aimed at creating the system based on individual trip planning, considering the future traffic. The data was gathered based on the real-time sensors reading. The systems consist of three main elements: an interactive web-based user interface similar to OpenTripPlanner app, real-time back-end engine, which imports data and dynamic traffic model, which estimates the traffic flow. This trip planner was implemented in the real-life case in Dublin, Ireland. <br/>

Furthermore, Shoshany-Tavory et al. (2014) determined how to increase transferability of Journey Planners through Multi-Attribute Tradespace Exploration (MATE) methodology. The goal of the research was to show how the methodology could be used in order to design Journey Planner. The research concluded that MATE can be used to simplify the search of the solutions for different stakeholder’s such as public authorities, private developers and passengers (Shoshany-Tavory et al., 2014).<br/> 

Solecka et al. (2022) compared and ranked digital journey planners in Krakow, based on nine criteria, to adapt journey planners to the needs of their users. The findings showed that applications should be easily accessible and adaptable to all traffic users, including people with reduced mobility. <br/> 

Further, Esztergár-Kiss (2019) provided an overview of European Journey planners considering the perspective of the users, by scoring the planners first and weighting them after, to quantify the evaluation of DJP and support the choice of suitable DJP for the user.<br/> 

Mobility as a Service (MaaS) plays a key role within digital journey planners and is discussed in more detail [here](#maas).

### Current state of art in practice {-}
Nowadays, the journey planners exist in majority of the European countries. They differ from each other based on the data they collect and services they provide. However, their main goal is the same, to provide passengers with reliable information to increase the efficiency and safety of the trip. The examples of trip planners around the world include Dutch [*9292*](https://9292.nl/en), Portuguese [*transporlis*](https://www.transporlis.pt/Default.aspx?tabid=36) and Romanian *tpltm*. They offer similar features to users, however they differ in specificity of the data. For instance, some of them work around the whole country and others trip planners are just regional facility (Ştefănescu, et al., 2014). ). After completing the [LinkingDanube Interreg project](https://www.interreg-danube.eu/approved-projects/linking-danube), AustriaTech is currently working on the [LinkingAlps](https://www.alpine-space.eu/projects/linkingalps/en/home) which is, similarily to the LinkingDanube project, a transnational travel information service for a multimodal and environmentally-friendly mobility in the Alpine region running until end of June 2022. Currently in practice, the VOR A to B routing service, which is a multimodal app, initiated by the ITS Vienna Region and VOR (Verkehrsverbund Ost-Region), comparing routes and travel times for various travel modes in the eastern regions of Austria.

### Relevant initiatives in Austria {-}

- [austriatech.at-1](https://www.austriatech.at/de/grenzueberschreitende-reiseplanung-auf-dem-vormarsch/) 
- [austriatech.at-2](https://www.austriatech.at/en/projects//showprojekt/38/LinkingAlps) 
- [LinkingAlps](https://www.alpine-space.eu/projects/linkingalps/en/home) 
- [splitboarding.eu](https://www.splitboarding.eu/en/splitboard-routes/ski-route-planner) 
- [transit.navitime.com](https://transit.navitime.com/en/at/) 
- [its-viennaregion.at](https://www.its-viennaregion.at/en/products.html)
- [LinkingDanube Interreg project](https://www.interreg-danube.eu/approved-projects/linking-danube)

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table74, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Higher accessibility to public transport|**+**|Equality *(5,10)* |Stefanescu et al., 2014|
|Individual|Increasing usage of public transport|**+**| Environmental sustainability (*7,12,13,15*)|Stefanescu et al., 2014|
|Systemic|Decreased emissions, use of paper maps and tickets|**+**|Sustainable economic development (*8,11*)|Stefanescu et al., 2014|
|Systemic|Digitalisation of transport facilities |**+**|Innovation & Infrastructure (*9*)|Liebig et al., 2017|
|Systemic|Collaboration between private app developers and public transport sector|**+**|Partnership & collaborations (*17*)|Jakob et al., 2014; Shoshany-Tavory et al., 2014|')
```

### Technology and societal readiness level {-}

```{r table75, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|7-9|')
```

### Open questions {-}

1. How to improve accuracy of arrival and departure of public transport in Digital Journey Planners in relation to real-life situation using real-time data? 
2. How to implement diverse types of data in DJP such as number of passengers or cost of ride to improve the quality of travel in public transport for passengers? 

### Further links {-}

- [opentransportdata.swiss](https://opentransportdata.swiss/en/cookbook/open-journey-planner-ojp/)
- [Personalized Fully Multimodal Journey Planner](https://upcommons.upc.edu/bitstream/handle/2117/127282/FAIA263-1225%20%282%29.pdf?sequence=3&isAllowed=y)

### References {-}
- Amrani, A., Pasini, K., & Khouadjia, M. (2020, November). Enhance Journey Planner with Predictive Travel Information for Smart City Routing Services. In 2020 Forum on Integrated and Sustainable Transportation Systems (FISTS) (pp. 304-308). IEEE.
-	Borole, N., Rout, D., Goel, N., Vedagiri, P. and Mathew, T., (2013). Multimodal Public Transit Trip Planner with Real-time Transit Data. Procedia - Social and Behavioral Sciences,  104, pp.775-784. Available at: <https://reader.elsevier.com/reader/sd/pii/S1877042813045631?token=42DB6D6836ED2B503BB611823B60EDAC7087D7DA77A894C14E04BDBDE0AD81BAC6E335407F8059EC5DC8DC9DA0C7C345>.
- Esztergár-Kiss, D., (2019). Framework of Aspects for the Evaluation of Multimodal Journey Planners. Sustainability, 11 (18), 4960.
-	Ştefănescu, P., Mocan, M., Ştefănescu, W. and Neculai, P., (2014). Trip Planners Used in Public Transportation. Case Study on the City of Timişoara. Procedia - Social and Behavioral Sciences,  124, pp.142-148. Available at: <https://reader.elsevier.com/reader/sd/pii/S1877042814020187?token=4E35403DB645FDA2D21DC662D8ADD8BBEA08E3794B321355E3DE9478F87F219256238D04D8D6AAF89E2ABEBFE1FBBDB5>.
-	Jakob, M., Hrncir, J., Oliva, L., Ronzano, F., Zilecky, P. and Finnegan, J., 2014. Personalized Fully Multimodal Journey Planner. ECAI 2014,  Available at: <https://upcommons.upc.edu/bitstream/handle/2117/127282/FAIA263-1225%20%282%29.pdf?sequence=3&isAllowed=y> [Accessed: 13 March 2021].
-	Liebig, T., Piatkowski, N., Bockermann, C. and Morik, K., (2017). Dynamic route planning with real-time traffic predictions. Information Systems,  64, pp.258-265. Available at: <https://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.688.687&rep=rep1&type=pdf> [Accessed: 13 March 2021].
- Solecka, K., Kikinski, M., (2022). A Multi-Kriteria Analysis of Applications Supporting Public Transport Users. Energies, 15, 3493.
-	Shoshany-Tavory, S., Gal-Tzur, A. and Eden, N., (2014). Incorporating Systems Engineering Methodologies to Increase the Transferability of Journey Planners. Transportation Research Procedia, 3, pp.631-640.
-	Puiu, D., Bischof, S., Serbanescu,B., Nechifor,S., Parreira, J. and Schreiner, H. (2017). A public transportation journey planner enabled by IoT data analytics. 20th Conference on Innovations in Clouds, Internet and Networks (ICIN), Paris, France, 2017, pp. 355-359, doi: 10.1109/ICIN.2017.7899440.

## Multimodal information and route planning {#info_and_route_planning}

**Updated: 17th May 2022**

### Synonyms {-}
*Multi-Modal Transport Information Services (MMTIS), Real Time Passanger Information (RTPI)*

### Definition {-}
Information systems that ensure reliable, efficient operations and widely accessible, accurate passenger information are essential to the use of public transportation. These systems are being used for a number of specific purposes such as setting schedules and timetables, managing vehicle fleets, issuing tickets and receipts, providing real time information on service running (European Commission, n.d. b). It is particularly crucial for multimodal transport, where passengers use several transport modes or systems operators within a single trip. Therefore, it is essential to provide accurate information to ensure smooth transfer and minimise waiting times. <br/>
Since 2006 the Transmodel (European Standard Public Transport Reference Data Model (EN 12896)) has been developed in Europe. It provides a framework for defining and agreeing data models to cover the whole area of public transport operations. Among other things, the standard facilitates interoperability between the information processing systems of transport operators and agencies by using matching definitions, structures and data formats for the component systems. Therefore, it is possible for operators, authorities and software suppliers to work together more easily towards integrated systems. Beyond, future system developments can be accommodated with the minimum difficulty (European Commission, n.d. b). Furthermore, Transmodel is particularly suitable as *(1)* a specification of an organization's "information architecture," *(2)* a specification of a database, and *(3)* a specification of a data exchange interface (European Commission, n.d. a) and has been a fundamental input for the design of the following EU standards:

-	DVC (Data Communication on Vehicles);
-	IFOPT (Identification of Fixed Objects in Public Transport);
-	SIRI (Standard Interface for Real-Time Information);
-	DJP/OJP (Open API for distributed journey planning);
-	NeTEx (Network Timetable Exchange);
-	OpRa (Operating Raw Data and statistics exchange). 

Belgium, Denmark, Finland, France, Germany, Hungary, Italy, Netherlands, Norway, Spain, Slovenia, Sweden Switzerland and the UK are involved in the implementation of Transmodel at a certain level. The latest research shows that AustriaTech initiated action towards the implementation of the Transmodel in Austria with the project [DATA4PT](https://www.austriatech.at/de/projekte//showprojekt/40/DATA4PT), that is aimed at advancing data-sharing practices in the public transport sector by supporting the development of data exchange standards and models, to fulfil the needs of multimodal travel information service providers.<br/>

The *Commission Delegated Regulation (EU) 2017/1926 of 31 May 2017 supplementing Directive 2010/40/EU of the European Parliament and of the Council with regard to the provision of EU-wide multimodal travel information services* required each EU Member State, among other things, to establish the National Access Points (NAP) by 1st December 2019 (European Commission, 2021a). NAPs were aimed at facilitating the access, easy exchange and reuse of transport-related data to enable the provision of EU-wide interoperable travel and transport services to end-users (European Commission, 2021b). In Austria the NPA is managed by [AustriaTech](https://www.mobilitydata.gv.at/). 

### Key stakeholders {-}
- **Affected**: Public Transport Users 
- **Responsible**: European Commission, NPAs, Transport service providers and public transport operators, National Governments, IT System Providers

### Current state of art in research {-}
Presently, the research aims at improving compatibility, connectivity and flexibility of the system. For example, Dinko et al. (2020) looked at what real-time information would be needed in the future to make the trip planner system resilient and easily adaptable to changing needs, considering the changes and challenges that will bring post-pandemic transportation requirements. Further Chang et al. (2020) conducted research on the accessible indoor navigation application for smartphones, combining information of floor plans, Bluetooth beacons, Wi-Fi/cellular data connectivity, 2D/3D visual models, and user preferences. The aim was to provide visual, audio, and haptic feedback for targeted users to find the optimal route to their destination within a building. Moreover, Huang et al. (2018) looked at the approaches to better connect static (e.g. public transport) and dynamic (e.g. carpooling) networks without limiting flexibility. They found that a merged network, based on the concept of drive-time areas and points of action, could enable multimodal route planning that can provide users with trips from a starting point to a destination using different combinations of modes. Namoun et al. (2021), analysed a route guidance system connecting multiple data sources, such as life traffic data, which were collected from multiple data sources (e.g. road sensors, SCOOT). Advantages of using different types of agents and the empowerment of management authorities to monitor traffic flow and conditions of the city in real time are once again highlighted.<br/>

What is more, the additional research addresses Mobility as a Service (MaaS) solutions. For further information see section [Mobility as a service (Maas)](#maas) of this work.


### Current state of art in practice {-}
*Verkehrsauskunft Österreich (VAO)*  is a nationwide multimodal routing platform authorized by Austrian transport infrastructure, means of transport and transport editorial operators, coordinated among themselves. Highly up-to-date, real-time routing information is provided for most modes of transport and their linking options, such as: car, public transport, bicycle, bike & ride, park & ride, rental bikes, car sharing, etc. The basis for the VAO is the *Graph Integration Platform (GIP)*, a routable and intermodal graph for the entire Austrian transport network and *Basemap*, an Austrian-wide, detailed administrative map based on the geodata of the Austrian federal provinces and the GIP (Verkehrsauskunft Österreich, n.d.).<br/>

*Rome2rio* is an international door-to-door travel information and booking engine which provides flight, train, bus, ferry, rideshare or rental car info as well as estimated prices, journey duration and booking details. The platform uses data from over 5000 companies of more than 160 countries (Rome2rio Pty Ltd, 2021a) and provides a simple XML/JSON interface for integrating multi-modal search capability to any website, app or service (Rome2rio Pty Ltd, 2021b). <br/>

With *Google Routs*, integrated in *Google Maps*, routes can also be called up for journeys by public transport, bicycle and car, as well as for journeys by foot. Based on real-time information on traffic conditions, current or future travel times are determined (Google, n.d.). Furthermore, the public transportation planning feature *Google Transit* combines the latest agency data (provided with *Google Routs*) with the power of *Google Maps*. It integrates transit stop, route, schedule, and fare information to make trip planning quick and easy for users. For the integration public transit agencies need to meet a few basic requirements, as providing their static and real-time data in the general transit feed specification (GTFS) format (Google, 2021). <br/>

Finally, in 2020 *Apple* launched a new online map, including the features of real-time transit information and indoor maps for airports and malls, but not yet worldwide (Apple, 2020). <br/>

Siemens Mobility, together with Wiener Linien, will launch an innovative digital passenger information and guidance system for the Vienna metro system. The dynamic and visual presentation of information through screens inside the metro provide passengers with real-time information about further routes and connections at the next station. The system will debut on the U1-U4 lines of the Vienna metro system along with the launch of the first new [X-car](https://press.siemens.com/global/en/pressrelease/wiener-linien-presents-innovative-digital-passenger-information-system-new-x-car) in 2022.

### Relevant initiatives in Austria {-}

- [verkehrsauskunft.at](https://www.verkehrsauskunft.at/) 
- [gip.gv.at](https://www.gip.gv.at/) 
- [basemap.at](https://www.basemap.at/) 
- [mobilitydata.gv.at](https://www.mobilitydata.gv.at/) 
- [maps.google.com](http://maps.google.com/landing/transit/cities/index.html#Europe)
- [Siemens Mobility](https://energyindustryreview.com/tech/worlds-first-digital-passenger-information-system-by-siemens-mobility/)
- [DATA4PT](https://data4pt-project.eu)
- [X-car](https://press.siemens.com/global/en/pressrelease/wiener-linien-presents-innovative-digital-passenger-information-system-new-x-car)

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table40, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Higher accessibility to mobility|**+**|Equality *(5,10)* |Dibbelt et al., 2017|
|Individual|Facilitated use of public transport to increase its use|**+**| Environmental sustainability (*7,12,13,15*)|European Commission, 2021a|
|Systemic|Accessible data on a non-discriminatory basis| **+**|Equality *(5,10)* |European Commission, 2021b |
|Systemic|More efficient transport system|**+**|Sustainable economic development (*8,11*)|European Commission, 2021a|
|Systemic|Implementation of NAPs, framework for defining and agreeing data models |**+**|Innovation & Infrastructure (*9*)|European Commission, 2021b; European Commission, n.d. b|
|Systemic|Collaborations of private and public sectors & global partnerships|**+**|Partnership & collaborations (*17*)|European Commission, n.d. b|')
```

### Technology and societal readiness level {-}

```{r table41, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|3-7|3-7|')
```

### Open questions {-}

1. How can bus and rail travel be made as convenient to book as flights? 
2. What effect has multimodal information & route planning on energy consumption? 
3. What key features are needed in multimodal trip planning apps to be accessible for disabled or elderly and how can they be implemented to be inclusive for these specific groups?
 
### Further links {-}

-	[transmodel-cen.eu](http://www.transmodel-cen.eu/)
-	[trafi.com](https://www.trafi.com/)
-	[eur-lex.europa.eu](https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX%3A32017R1926)
-	[ec.europa.eu](https://ec.europa.eu/transport/sites/transport/files/2020-02-implementation-handbook-delegated-reg20171926.pdf) 
-	[mobilitydata.gv.at](https://www.mobilitydata.gv.at/) 
-	[apple.com](https://www.apple.com/maps/) 
-	[maps.google.com](http://maps.google.com/landing/transit/cities/index.html#Europe) 
- [RTPI](https://www.railway-technology.com/news/the-real-time-passenger-information-systems-providing-reassurance-for-passengers/)

### References {-}
- Apple. (2020). Apple delivers a new redesigned Maps for all users in the United States - Apple. Available at: https://www.apple.com/newsroom/2020/01/apple-delivers-a-new-redesigned-maps-for-all-users-in-the-united-states/ [Accessed: 15 February 2021]
- Chang, Y., Chen, J., Franklin, T., Zhang, L., Ruci, A., Tang, H., & Zhu, Z. (2020, August). Multimodal Information Integration for Indoor Navigation Using a Smartphone. In 2020 IEEE 21st International Conference on Information Reuse and Integration for Data Science (IRI) (pp. 59-66). IEEE. 
- Dibbelt, J., Konstantopoulos, C., Wagner, D., Gavalas, D., Kontogiannis, S., Zaroliagis, C., ... & Pantziou, G. (2017, July). Multimodal route and tour planning in urban environments. In 2017 IEEE Symposium on Computers and Communications (ISCC) (pp. 214-219). IEEE.
- Dinko, A., Yatskiv, I., & Budilovich, E. (2020, October). Trip Planner Challenges in the Era of Fast Changing Requirements. In International Conference on Reliability and Statistics in Transportation and Communication (pp. 485-496). Springer, Cham.
- European Commission. (n.d. a). Applicability of Transmodel – Transmodel. Available at: http://www.transmodel-cen.eu/overview/applicability-of-the-transmodel/ [Accessed: 8 February 2021]
- European Commission. (n.d. b). Purpose of the Transmodel – Transmodel. Available at: http://www.transmodel-cen.eu/overview/purpose-of-the-transmodel/ [Accessed: 8 February 2021]
- European Commission. (n.d. c). Transmodel countries – Transmodel. Available at: http://www.transmodel-cen.eu/implementations/countries/ [Accessed: 12 February 2021]
- European Commission. (2021a). Multimodal Travel Information | Mobility and Transport. Available at: https://ec.europa.eu/transport/themes/its/road/action_plan/multimodal-travel-information_en [Accessed: 12 February 2021]
- European Commission. (2021b). National Access Points | Mobility and Transport. Available at: https://ec.europa.eu/transport/themes/its/road/action_plan/nap_en [Accessed: 12 February 2021]
- Google. (2021). Get started with Google Transit - Transit Partners Help. Available at: https://support.google.com/transitpartners/answer/1111481?hl=en&ref_topic=3521043 [Accessed: 8 February 2021]
- Google. (n.d.). Routes & Directions  |  Google Maps Platform  |  Google Cloud. Available at: https://cloud.google.com/maps-platform/routes [Accessed: 12 February 2021]
- Namoun, A., Tufail, A., Mehandjiev, N., Alrehaili, A., Akhalghinia, J., Peytchev, E., (2021). An Eco-Friendly Multimodal Route Guidance System for Urban Areas Using Multi-Agent Technology. Applied Sciences, 11, 2057.
- Huang, H., Bucher, D., Kissling, J., Weibel, R., & Raubal, M. (2018). Multimodal route planning with public transport and carpooling. IEEE Transactions on Intelligent Transportation Systems, 20(9), 3513-3525.
- Rome2rio Pty Ltd. (2021a). About Rome2rio - Rome2rio.Available at: https://www.rome2rio.com/about/ [Accessed: 12 February 2021]
- Rome2rio Pty Ltd. (2021b). Rome2rio API - Rome2rio. Available at: https://www.rome2rio.com/documentation/ [Accessed: 12 February 2021]
- Siemens. (2022). Worlds First Digital Passenger Information System by Siemens Mobility. Available at: https://energyindustryreview.com/tech/worlds-first-digital-passenger-information-system-by-siemens-mobility/ [Accessed: 14 May 2022]
- Verkehrsauskunft Österreich. (n.d.). Verkehrsauskunft Österreich: VAO Heute. Available at: https://www.verkehrsauskunft.at/ [Accessed: 8 February 2021]
