# Automated driving {#connected}

## Automated passenger cars {#av}
### Synonyms {-}
*Automated Vehicle (AV), Shared Automated Vehicles (SAVs)* 

### Definition {-}
The potential of automated cars is great - for society, for safety and for Europe as a business location. For society, the opportunity lies in better integrating elderly or disabled people and offering people the possibility to use the time during the journey productively or for recreation. Automated taxis or buses may run so cheaply that rural areas can also be made more accessible. Traffic runs more smoothly and goods can be transported in a more rational and environmentally friendly way. Depending on the degree of automation, the number of accidents could also be further reduced, because human error is the cause of 90 % of all accidents. However, this process will take a long time because conventional and automated vehicles will still be driving in mixed traffic for many years to come. Failures of technical systems or misjudgement of traffic situations must be avoided, henceforth, the development of the best technology for autonomous driving is immensely important. According to surveys in Germany, however, 45 % of drivers currently do not believe in the reliability of vehicle technology or are afraid of hackers (Rudschies & Kroher, 2020).<br/>

The levels of automation are currently divided into 5 levels (Paulsen, 2018):

**First level: assisted driving**<br/>
Individual assistance systems provide support for specific driving tasks. Assisted driving is already a reality in many cars today. (automatic adaptive cruise control and automatic lane departure warning).

**Second level: partially automated driving**<br/>
With semi-automated driving, the car/truck can temporarily perform some tasks itself - without any human intervention. For this purpose, various individual systems are combined with each other - in this case, the automatic adaptive cruise control with the emergency braking and lane departure warning systems. Further level 2 functions (overtaking assistant, automatic parking)

**Third level: highly automated driving**<br/>
Highly automated cars/trucks (Level 3) can perform certain driving tasks autonomously and without human intervention, but only for a limited period of time and under suitable conditions specified by the manufacturer. Level 3 cars will probably be on the road on motorways first: There is no oncoming traffic there, the lane markings are usually in order, and the roads are continuously recorded as digital maps. Since 2017, there has also been a legal framework for Level 3 cars in Germany: as soon as the driver puts his car into highly automated mode, he is allowed to turn his attention away from the road traffic. This means that the driver is allowed to read the newspaper. 
However, if the system detects a problem and sends a signal, the driver must take over the wheel immediately. 

**Fourth level: fully automated driving**<br/>
In the development departments of the large car companies, but also at Apple, Google or Uber, engineers and computer scientists are working flat out on the full automation of the car, i.e. level 4 on the way to autonomous driving. In this level, the technical systems carry out all driving tasks automatically, and the car can also cover longer distances without intervention.
So far, however, there is no legal framework for fully automated vehicles.

**Fifth level: autonomous driving**<br/>
The fifth is the final level of autonomous driving. The car is now completely guided by the system and performs all the necessary tasks automatically. Even complex situations - such as crossing an intersection, driving through a roundabout or the correct behaviour at a zebra crossing - can be handled by the autonomous car. 

So far, however, there is no legal framework for autonomous vehicles. Therefore, the rights and obligations of the manufacturers of the cars/trucks, software developers and the insurance companies, in the automated mode of operation, are still completely unclear. The ADAC experts suggest only *3 operating modes* instead of 5 levels (Paulsen, 2018), where at the moment the car supports us (1st operating mode), soon it will drive itself (2nd operating mode) and in future driverless (3rd operating mode). They can be described as follows:

**First mode: assisted driving**<br/>
The driver is always and at all times responsible for driving. As far as possible, the vehicle keeps in lane, brakes and accelerates. But: feedback is not mandatory and a sudden stop is always possible. 

**Second mode: automated driving**<br/>
The vehicle only drives itself in the application case specified by the manufacturer (e.g. stop-and-go in a traffic jam). The driver may temporarily turn away from the driving task and the traffic situation. Outside employment is possible. The driver is only liable if he has not complied with the request to take over.

**Third mode: autonomous driving**<br/>
The driver becomes a passenger, or the vehicle can be operated without occupants. The autonomous mode can be limited to defined routes. An operator (not the driver) must constantly monitor the vehicle in order to be able to react to operational problems (flat tyres, etc.).<br/>
In a forecast of the introduction of automation functions in the passenger car fleet by 2050 in Germany, 3 different automation functions are considered that build on each other (Altenburg et al., 2018):

-	Motorway pilot: The motorway pilot enables level 4 highly automated driving on all motorways. This automation level enables the vehicle to move on this type of road completely without monitoring or even intervention by the driver. When leaving the motorway, the vehicle hands control back to the human driver.
-	City Pilot: The City Pilot can take control not only on motorways, but also in the entire urban environment up to a speed of 50 km/h. On rural roads that do not have a structural separation of the directional lanes and that are not subject to the urban speed limit, the driving task must be taken over by a human.
-	Door-to-door pilot: This most far-reaching automation function allows vehicles to drive the entire road network in automated mode without driver intervention. Here, too, automation will encompass level 4.


### Key stakeholders {-}

- **Affected**: Citizens, drivers, cyclists, public transport users and drivers, pedestrians
- **Responsible**: National governments, City government, Car manufacturers

### Current state of art in research {-}
The main focus in research is currently concentrated on safety, social acceptance and external effects. Much of the ongoing research is considered to be privately owned by car manufacturers regarding safety and efficiency.<br/>

Although fully automated cars are seen as a means of improving network capacity and network reliability, making car use more accessible to a wider range of people and reducing accidents, several researchers warn that the rebound effect is that they could lead to a significant increase in car use, with a simultaneous decrease in pedestrians, cyclists and public transport, and that this could more than offset the benefits of increased capacity and lead to urban sprawl. May et al. (2020), following a review of the literature and qualitative analysis using causality diagrams, suggest that the key triggers of such change will be the market share of automated vehicles, whether private or shared, the extent to which capacity is increased, the potential to reduce time spent on parking and access, the potential reduction in the value of time spent in vehicles, and the expansion of the number of people who can drive. Their findings indicate that the introduction of automated cars is likely to have negative impacts on the environment, accessibility, health, urban sprawl and overall sustainability. They also suggest that these negative impacts can be mitigated by requiring that automated vehicles are provided in cities as shared vehicles rather than private vehicles. However, it seems that the fee needs to be higher than the one currently charged by car sharing companies to effectively avoid these negative impacts.<br/>

Zhou et al. (2020) investigates preference heterogeneity in mode choice for car-sharing and shared automated vehicles. They show that the use of AVs in car-sharing schemes does not currently receive much research focus, mainly due to the novelty of these technologies. Although the emergence of AVs could help drive car-sharing adoption by removing various barriers such as lack of parking and undesirable access costs. Proponents argue that the combination of carsharing and AVs (shared automated vehicles [SAVs]) offers unique opportunities for policy makers to positively influence and shape outcomes for society, for example SAVs could provide safe and accessible mobility services for the elderly, non-drivers, disabled and children (Howard, 2014; Yang & Coughlin, 2014).<br/>

Nordhoff et al. (2020) investigated the public acceptance of partially automated (SAE Level 3) passenger cars by means of a questionnaire among 9,118 drivers in eight European countries within the European L3Pilot project. Respondents considered partially automated cars easy to use, but were less likely to consider buying one. Slightly less than the majority could imagine engaging in off-road activities, such as talking to fellow drivers, surfing the Internet, watching videos or TV shows, observing the landscape or working. The positive effects of hedonic motivation, social influence and performance expectancy on the behavioural intention to use conditionally automated cars suggest that the benefits of conditionally automated cars need to be clearly demonstrated and promoted by public (e.g. media, policy) and private decision makers (e.g. manufacturers) through established communication channels and in social networks. Car dealers could be trained as real experts on automated driving functions to explain the strengths and limitations of the systems and to show the systems to customers, e.g. by offering test drives. Social networks (both online and offline) could play a more important role in promoting the benefits of conditionally automated cars, as friends, family and colleagues are an important and trusted source of information. Public authorities could form alliances with private organisations (e.g. automobile clubs, road construction companies) to launch joint awareness campaigns on the benefits of automated cars and to conduct test drives to familiarise the general public with automated cars.<br/>

Kaye et al. (2020) also examines the a priori acceptance of highly automated cars through the Theory of Planned Behaviour (TPB) and the Unified Theory of Acceptance and Use of Technology (UTAUT) in Australia, France and Sweden. For France and Sweden, participants who had prior knowledge of AVs reported significantly higher intentions to use automated cars in the future than those who had no prior knowledge. However, it is important to note that the majority of individuals in each country (i.e. over 80% of participants in Australia and France and 99% of participants in Sweden) reported pre-existing knowledge about AVs. In addition, individuals residing in France indicated a significantly greater intention to use highly automated vehicles if they were publicly available than individuals residing in Australia and Sweden. While it is not known why participants in France indicated significantly greater intentions to use automated cars when they are publicly available than participants in Australia and Sweden, these results support previous research that has shown that perceptions of AVs can vary across countries (e.g. Schoettle & Sivak, 2014) and highlight the need to consider country of residence when assessing AV uptake.<br/>

Vlakveld et al. (2020) investigates cyclists' intentions to yield to automated cars at intersections when they have the right of way using an experiment with high-quality video animations. The results indicate that when cyclists have the right of way at intersections, they still sometimes swerve for cars. They swerve more often when the approaching car is an automated car than when the approaching car is a conventional car. However, when automated cars communicate to cyclists that the system has noticed them and that the car will act in accordance with traffic rules, cyclists tend to swerve less often than in similar conflict situations with a traditional car. It is important to note that the fact that cyclists are more inclined to continue when the automated car can indicate its intention to swerve does not necessarily improve road safety. It is beneficial for road safety if road users always correctly predict what other road users will do. Since the driving behaviour of automated cars can differ from the driving behaviour of cars driven by humans, indicating their intentions can help other road users to make the correct predictions because they know what to expect. However, this will only improve road safety if the displayed intentions are always correct.<br/>

Galich & Stark (2021) investigated the impact of the introduction of automated vehicles in private car ownership. The results show that a reduction or increase in private car ownership, depends to a large extent on the actual implementation of automated vehicles in terms of pricing structures, legal regulations, mobility services, etc. Despite the significantly higher waiting times and operating costs that other studies predict for less densely populated areas, the results showed that residents of rural areas are quite willing to accept higher costs and longer waiting times than their urban counterparts if they are offered reliable mobility alternatives to their own car. Rural residents were even willing to book automated ride-sharing services up to a day in advance for certain trips, such as commuting to work. Interviewed participants stated that they are willing to send their children to kindergarten or school alone in an automated vehicle, but that there should be kindergarten staff or school teachers to pick up the children from the vehicle and accompany them for the last few metres. Another very important point of the study is that automated vehicles could be a very attractive option for older people or people with mobility impairments and thus create a new social group of car owners. This could be socially desirable, as the mobility needs of this group often cannot be met by the currently available means of transport. However, this new social group of car owners can have a negative impact on land use, hence, it is still more favourable that public transport operators and car and ridesharing providers consider offering special services that address the specific needs of this user group. In this way, the mobility needs of this group could be met without leading to an increase in car ownership.


### Current state of art in practice {-}
Successes in research and development of automated vehicles differ from practical developments and reality on the roads. A recent study by the Prognos Research Institute on autonomous driving for the ADAC shows that automated driving will only slowly become established. This is mainly due to the fact that cars are in use for up to twenty years on average, and new technologies increase their presence very gradually in the overall stock. According to the forecast, the share of new vehicles in which the driver can completely turn away from the driving task on all motorways will increase in the "optimistic" case from 2.4 % in 2020 to as much as 70 % in 2050. From 2030 onwards, cars with Citypilot, i.e. the ability to drive alone both on the motorway and in the city, will gradually appear on the roads. And only after 2040 will a larger number of cars be available that can go from door to door fully autonomously, i.e. that no longer need a driver even on country roads. This means that for a long time to come, completely normal vehicles will be on the road alongside fully automated ones. This also puts the hope of rapid safety gains through autonomous cars in the next few decades. The fact that fewer and fewer people will die or be injured in road traffic will be due to the spread of efficient assistance systems, such as the emergency brake assistant (Rudschies & Kroher, 2020).<br/>

At the forecast horizon of 2050, about half of the vehicles will already have an automation function, but in most cases this will only be usable on motorways. Significant penetration of vehicles capable of automated driving throughout the network is not expected until after 2050. Vehicles with automation features are likely to account for significant shares of driving by 2050, but the limitations of the driving features delivered, mean that only a fraction of this will take place on road types for which suitable automation is also available. In addition, it cannot be assumed that the functions will be fully deployed immediately. Consequently, by 2050 a maximum of one in five vehicle km could be provided by automation. This value varies greatly between the different types of roads: while on motorways a good 40% of the mileage could be automated, on rural roads it is still less than 4%. The low proportion of automated driving also influences the safety effects. Automated driving can only make significant contributions slowly towards the end of the forecasted horizon. Since serious consequences of accidents occur particularly strongly on rural roads, where automation will hardly have any effect by 2050, the effect of automation on traffic fatalities will still be marginal by that time (Altenburg et al., 2018).<br/>

Driverless cars are to be able to participate in road traffic in Germany as early as 2022. Autonomous motor vehicles of the so-called "level four" will then be allowed to travel on certain defined routes on public roads. In fully automated driving, the computer can take complete control of the car in certain applications without being monitored by a human driver. In emergencies, the system should also bring the vehicle to a halt at the side of the road. According to the German Ministry of Transport, this technology could be used for shuttle services or freight transport, for example (Rudschies & Kroher, 2020).<br/>

The competition for the best technology for autonomous driving is currently ongoing worldwide. According to VW boss Herbert Diess, however, German car manufacturers are "one to two years behind" in developing the technology for autonomous driving. At the forefront of know-how, almost all experts locate companies in the USA - especially the US company Waymo, which belongs to Google. In the US state of Arizona, Waymo has been running robot cars as taxis for a while now - but still with a safety driver behind the wheel. Customers can order their ride via app and are picked up wherever they happen to be. On average in the vehicles run by Waymo the security guard only has to intervene 0.09 times per 1000 miles driven.<br/>

The development gap between Germany and the Americans could be due to the fact that German car manufacturers work with a higher degree of thoroughness. Christian Weiß, system developer at Daimler, emphasises: "Basically, safety comes before speed for us." For him, the Tesla accidents are a good example of what can happen when immature technology is used in production vehicles. Daimler expert Timo Winterling adds: "To understand a scene, we have to get the last out of the sensors."<br/>

BMW is currently heavily involved in collecting and analysing data in real traffic. At the BMW Campus in Unterschleissheim near Munich, which was founded specifically for this purpose, around 1700 specialists are now working on developing the necessary software algorithms for highly automated driving. As test vehicles, 40 BMWs are on the road collecting huge amounts of data and images in road traffic. For storage, BMW has built two data centres with a capacity of 500 petabytes (PB), a storage size that would hold all the words ever written and printed in human history about five times over (Rudschies & Kroher, 2020).


### Relevant initiatives in Austria {-}
In December 2016, a legal framework for automated vehicles and their systems was created in Austria. According to this, test drives by research institutions and vehicle manufacturers were allowed to assign certain driving tasks, such as keeping a distance, accelerating, braking or keeping in lane, to an assistance system or automated or connected driving system present in the vehicle, in compliance with the other legal regulations.

- [Oesterreich.gv.at](https://www.oesterreich.gv.at/themen/freizeit_und_strassenverkehr/kfz/Seite.061910.html)

Since 11 March 2019, hands-free driving on motorways and with 'motorway assistant with automatic lane guidance' and automatic parking with parking assistant have been permitted. "However, one must be able to intervene immediately if an unexpected situation arises. Therefore, distracting activities, such as operating a mobile phone, remain prohibited. The system may also not be used in construction areas.

- [Öamtc.at](https://www.oeamtc.at/thema/vorschriften-strafen/automatisiertes-fahren-30601779)

Truly autonomous driving is currently only allowed in Austria in test operation on certain stretches of road and requires a separate permit. This is primarily to allow companies to test self-driving cars.

- [Unterkaerntner.at](https://unterkaerntner.at/index.php?id=5488)

With the action package Automated Mobility (2019-2022), the government programme 2017-2022 has set out to develop Austria into a pioneering country and thus also into a research, development and production location for automated driving in close cooperation with the automotive industry and research. In particular, the ministry will continue to promote test tracks and related research projects.

- [Bmk.gv.at](https://www.bmk.gv.at/themen/mobilitaet/alternative_verkehrskonzepte/automatisiertesFahren/aktionsplan.html)


### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Potential for increasing (access to) mobility in elderly and other disadvantaged groups|**+**|Equality (*5,10*)|Howard, 2014; Yang & Coughlin, 2014|
|Systemic|Potential for reduction in accidents |**+**|Health & Wellbeing (*3*)|Rudschies & Kroher, 2020|
|Systemic|Potential for increase in car ownership|**-**|Environmental sustainability (*7,12-13,15*)|May et al., 2020|')
```

### Technology and societal readiness level {-}
```{r echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|6-8|5-7|')
```

### Open questions {-}
1. Can ownership of automated cars satisfy the mobility needs of a multi-person household that currently relies on two, three or even more cars?

### References {-}
-	Altenburg, S., Kienzler, H.-P., & Maur, A. A. der. (2018). Einführung von Automatisierungsfunktionen in der Pkw-Flotte.
-	European Commission. (2019). Automated road transport.
-	Galich, A., & Stark, K. (2021). How will the introduction of automated vehicles impact private car ownership? Case Studies on Transport Policy, 9(2), 578–589. https://doi.org/10.1016/j.cstp.2021.02.012
-	Howard, D. (2014). Public Perceptions of Self-driving Cars: The Case of Berkeley, California. MS Transportation Engineering, 2014(1), 21.
-	Kaye, S. A., Lewis, I., Forward, S., & Delhomme, P. (2020). A priori acceptance of highly automated cars in Australia, France, and Sweden: A theoretically-informed investigation guided by the TPB and UTAUT. Accident Analysis and Prevention, 137, 105441. https://doi.org/10.1016/j.aap.2020.105441
-	May, A. D., Shepherd, S., Pfaffenbichler, P., & Emberger, G. (2020). The potential impacts of automated cars on urban transport: An exploratory analysis. Transport Policy, 98, 127–138. https://doi.org/10.1016/j.tranpol.2020.05.007
-	Nordhoff, S., Louw, T., Innamaa, S., Lehtonen, E., Beuster, A., Torrao, G., Bjorvatn, A., Kessel, T., Malin, F., Happee, R., & Merat, N. (2020). Using the UTAUT2 model to explain public acceptance of conditionally automated (L3) cars: A questionnaire study among 9,118 car drivers from eight European countries. Transportation Research Part F: Traffic Psychology and Behaviour, 74, 280–297. https://doi.org/10.1016/j.trf.2020.07.015
-	Paulsen, T. (2018, November 7). Autonomes Fahren: 5 Level zum selbstfahrenden Auto | ADAC. ADAC. https://www.adac.de/rund-ums-fahrzeug/ausstattung-technik-zubehoer/autonomes-fahren/grundlagen/autonomes-fahren-5-stufen/
-	Rudschies, W., & Kroher, T. (2020, May 21). Autonomes Fahren: Der aktuelle Stand der Technik | ADAC. 20.11.2020. https://www.adac.de/rund-ums-fahrzeug/ausstattung-technik-zubehoer/autonomes-fahren/technik-vernetzung/aktuelle-technik/
-	Schoettle, B., & Sivak, M. (2014). A survey of public opinion about autonomous and self-driving vehicles in the US, UK and Australia. UMTRI, Transportation Research Institute, July, 1–38.
-	Vlakveld, W., van der Kint, S., & Hagenzieker, M. P. (2020). Cyclists’ intentions to yield for automated cars at intersections when they have right of way: Results of an experiment using high-quality video animations. Transportation Research Part F: Traffic Psychology and Behaviour, 71, 288–307. https://doi.org/10.1016/j.trf.2020.04.012
-	Yang, J., & Coughlin, J. F. (2014). IN-VEHICLE TECHNOLOGY FOR SELF-DRIVING CARS: ADVANTAGES AND CHALLENGES FOR AGING DRIVERS J. International Journal of Automotive Technology, 15(2), 333−340.
-	Zhou, F., Zheng, Z., Whitehead, J., Washington, S., Perrons, R. K., & Page, L. (2020). Preference heterogeneity in mode choice for car-sharing and shared automated vehicles. Transportation Research Part A: Policy and Practice, 132, 633–650. https://doi.org/10.1016/j.tra.2019.12.004



## Parking infrastructure for autonomous vehicles {#parking_av}

## Automated road freight {#automated_road_freight}

### Synonyms {-}
*automated driving systems (ADS), combustion-powered heavy vehicles (CHV), battery electric heavy vehicles (BEHVs), cooperative and automated vehicles (CAVs)* 

### Definition {-}
The transport sector will soon experience a revolution in automated driving systems (ADS), as reported by Wadud et al. (2016) and Flämig (2016). Some benefits in terms of user experience, efficiency, safety, mobility, productivity, energy, environment and economy have been reported with ADS by many studies (Alessandrini et al., 2015; Anderson et al., 2016; Brown et al., 2014; Chan, 2017; Harper et al., 2016; Wadud, 2017; Kopelias et al., 2020 and Khan et al., 2019), although significant increases in road safety from highly or fully automated vehicles are not certain, as shown by Kalra & Paddock (2016).<br/>

The goals and motivations of automated driving for passenger cars and freight transport differ greatly (Wadud, 2017; Nowakowski et al., 2015). For passenger cars, the main motivations are user experience and the environment, while for freight, the main drivers are productivity and profitability. For example, the increase in the cost of a vehicle related to automation hardware is less important in freight transport due to the lower share of automation hardware costs in the total acquisition costs compared to those of passenger cars. In addition, ADS enable higher profitability in freight transport, mainly due to a reduction in labour costs as well as facilitated logistics and increased utilisation and efficiency (Wadud et al., 2016). In addition, driver costs decrease significantly with high or full automation of driving, leading to early adoption of ADS in the freight sector (Ghandriz et al., 2020a).<br/>

The levels of automation are currently divided into 5 levels (Paulsen, 2018):

**First level: assisted driving**

Individual assistance systems provide support for specific driving tasks. Assisted driving is already a reality in many cars today. (see automatic adaptive cruise control and automatic lane departure warning).

**Second level: partially automated driving**

With semi-automated driving, the car/truck can temporarily perform some tasks itself - without any human intervention. For this purpose, various individual systems are combined with each other - in this case, the automatic adaptive cruise control with the emergency braking and lane departure warning systems. Further level 2 functions (overtaking assistant, automatic parking)

**Third level: highly automated driving**

Highly automated cars/trucks (Level 3) can perform certain driving tasks autonomously and without human intervention, but only for a limited period of time and under suitable conditions specified by the manufacturer. Level 3 cars will probably be on the road on motorways first: There is no oncoming traffic there, the lane markings are usually in order, and the roads are continuously recorded as digital maps. Since 2017, there has also been a legal framework for Level 3 cars in Germany: as soon as the driver puts his car into highly automated mode, he is allowed to turn his attention away from the road traffic. This means that the driver is allowed to read the newspaper. 
However, if the system detects a problem and sends a signal, the driver must take over the wheel immediately. 

**Level four: fully automated driving**

In the development departments of the large car companies, but also at Apple, Google or Uber, engineers and computer scientists are working flat out on the full automation of the car, i.e. level 4 on the way to autonomous driving. In this level, the technical systems carry out all driving tasks automatically, and the car can also cover longer distances without intervention.
So far, however, there is no legal framework for fully automated vehicles.

**Fifth level: autonomous driving**

The fifth is the final level of autonomous driving. The car is now completely guided by the system and performs all the necessary tasks automatically. Even complex situations - such as crossing an intersection, driving through a roundabout or the correct behaviour at a zebra crossing - can be handled by the autonomous car. <br/>

For a regulatory framework, seamless and harmonised cross-border EU and national regulation of automated road transport is seen by the European Commission, 2019 as a key factor for market introduction. Common procedures for testing, validation and certification enable the standardisation of solutions. Standards should also apply to data exchange and communication solutions. Ethical, legal, environmental and safety aspects should be regulated. The liability and responsibility of each actor in the road system must be defined before the introduction of automated driving.


### Key stakeholders {-}

- **Affected**: Truck drivers, Freight forwarders, Delivery companies
- **Responsible**: National Governments, City government, Freight forwarders, Delivery companies

### Current state of art in research {-}
The reduction in fuel consumption achieved with ADS due to improved vehicle use and controlled energy management is only up to 10% for passenger cars compared to when a human driver is involved, depending on the traffic scenario (Mersky & Samaras, 2016). The reduction in fuel consumption achieved with ADS is likely to be higher in freight transport, considering that heavy vehicles can form platoons, which theoretically reduce the energy intensity of the following vehicle by up to 25% when the distance between vehicles reaches zero, according to (Wadud et al., 2016); while measurements showed average fuel savings of 8% and 15% for 10 m and 4 m distance of the trucks in the platoon, respectively, according to (Tsugawa et al., 2016). However, as the European Commission (2016) points out, heavy-duty vehicles contribute to about 25 % of CO2 emissions from road transport in Europe, which are rising due to increasing road freight transport despite improved fuel efficiency. These figures suggest that the efficiency improvements offered by ADS, although a profitable business, cannot lead to a reduction in CO2 emissions in the long term; therefore, a further reduction of emissions in the transport sector is needed, motivating the development of greener solutions (Ghandriz et al., 2020a).<br/>

In a slightly simplified study of ADS in heavy vehicles, a reduction in the total cost of ownership (TCO) of trucks of between 27% and 46% for battery electric heavy vehicles (BEHVs) and between 11% and 41% for combustion-powered heavy vehicles (CHVs) makes BEHVs more profitable than BEHVs with human drivers by a factor of four over longer distances. It was observed that Level 4-5 automated BEHVs tend to have lower optimal speeds than vehicles with human drivers. Furthermore, it was shown that reducing speed for safety reasons is less costly to implement for Level 4-5 automated BEHVs than for Level 4-5 automated CHVs; in many scenarios, low speeds up to 60 km/h even reduced the TCO. The reduction in TCO in Level 4-5 automated trucks was mainly achieved by eliminating driver pay (35%-55%), increasing road travel time (0%-15%), increasing payload by eliminating the cab (1.2%-5%), and optimising the drive system (0%-20%) (Ghandriz et al., 2020a). All the data and figures obtained were made available in (Ghandriz et al., 2020b) and provide valuable information on the feasibility and profitability of a planned freight transport operation with automation and electrification for a specific use case.<br/>

Pribyl et al., (2020) are investigating the effects of lower CAV penetrations. According to this, even with a low number of CAVs on the roads, the impact on transport and environmental performance indicators will be significant. However, this integration needs to be combined with other measures and policies such as the promotion of alternative powertrains if EU emission targets are to be met. However, results have also shown that CAVs can negatively affect other traffic parameters and lead, for example, to delays and other traffic flow phenomena.<br/>

Shin et al., (2018) analyse some of the technical trends related to intermodal automated freight transport systems (ATFS) that aim to reduce greenhouse gases and particulate matter, minimise infrastructure costs and logistics costs, and alleviate traffic congestion. Countries such as the US, Germany, the Netherlands and Japan have been actively developing technologies for ATFS.
Gan et al., 2021 capture the characteristics of the swarm intelligence of truck drivers as a solid basis for the design, control and management of the future driverless truck swarm robotics system. The detection of the short-term location of truck drivers in advance, especially for areas with high freight demand, and thus the optimal use of freight capacity on the motorway, is the key component of freight matching. For city-level forecasting, it can help to determine the city points that truckers will be driving to, making it easier to plan an optimal route for them from the city's perspective. It can also predict the internal transport capacity of the city, predict the bottleneck caused by the truck and provide decisions for inner-city traffic management.

Mulholland et al., (2018) define a global assessment of the decarbonisation of road freight transport by 2050. Overall, global road freight transport in tonnes-kilometer (tkm) is projected to grow 2.4-fold over the period 2015-2050 in their Reference Technology Scenario (RTS), with developing countries accounting for most of the growth, in line with economic growth. Developing countries are projected to account for 75% of road freight transport in 2050. In order to reduce dependence on fossil fuels, three key requirements are proposed:

-	Fuel economy standards (minimum efficiency).
-	Support for comprehensive data collection and information exchange
-	Promoting the use of alternative fuels and the vehicles that use them

### Current state of art in practice {-}
The initial introduction of autonomous driving in closed traffic systems (e.g. traffic in a system network) and manageable scenarios (e.g. motorway, airport grounds, ports) helps to overcome the reluctance towards the technologies as well as the lack of standardisation. The gradual introduction of platooning, starting with manned lead and follow vehicles, could build up the necessary acceptance among the population. It should be examined whether the simultaneous introduction of overhead electric trucks and thus the creation of a dedicated lane would make sense in terms of capacity and safety requirements (Flämig, 2016).<br/>

A European Commission, 2019 report presents many automated road projects, but only one that deals specifically with freight transport:

-	Ensemble (01/06/2018- 31/05/2021)

The project aims to support the implementation of multi-brand truck platooning on European roads, so that a single truck can form a platoon with any other truck.

Other projects may approach the issue in a different way, but the focus is on all vehicles. For example:

-	Headstart (01/01/2019- 31/12/2021)

The project aims to define test and validation procedures for specific functionalities of Connected and Automated Driving (CAD), including key technologies such as communication, cyber security and positioning. The tests will take place both in simulation and in the real world to validate safety and security performance.

-	Levitate (01/12/2018- 30/11/2021)

The project will develop a wide-ranging assessment framework to evaluate the impact of connected and automated transport (CAT) on all aspects of transport and individual mobility, as well as at a societal level. The project addresses the needs of local authorities, regional authorities and national governments to prepare for the increasing use of connected and automated systems, to understand the implications for mobility policy and to take the most effective measures to achieve broader societal goals.<br/>

In the US four manufacturers that dominate the US long-haul truck market have all begun to look at a driverless future. In 2019-2021, each of them has signed contracts with one of the three leading developers of self-driving vehicles (Alphabet's Waymo, TuSimple Holdings and Aurora Innovation). Aurora plans to sell its driverless technology as a service, similar to the way jet engine makers charge airlines for hours spent on the throttle. TuSimple and Waymo Via have similar plans and are also operating cargo services while they test their technology. None of the partnerships with manufacturers are exclusive. The flurry of activity around trucking reflects a growing consensus that self-driving cars are still a long way from reaching most roads, and that 18-wheel trucks might be a better place to establish the technology. Long-haul trucks drive straight most of the time on highways without traffic lights, cyclists or pedestrians crossing the road (Boudway, 2021).

Self-driving vehicles could have an immense impact on logistics and supply chain management. This is because autonomous trucks, last-mile delivery robots and other AVs fit into an increasingly automated distribution process. The automation of high-volume freight routes has the potential to deliver huge efficiency gains in terms of improved safety, increased capacity, lower operating costs and reduced environmental impact. An analyst at Gartner predicts that while self-driving trucks won't be commercially available until 2024 at the earliest, supply chain executives responsible for transportation should start preparing for an autonomous future now. After all, autonomous trucks are just one part to fully automating the entire supply chain, from autonomous inventory planning to autonomous shipping to autonomous distribution. In a report (March 2021), a market research firm predicted that autonomous trucks will be more widely deployed by 2030, accounting for more than 10% of new heavy-duty truck sales on the road. By 2028, one-fifth of the world's countries will have active regulations allowing the legal operation of production-ready autonomous vehicles, up from zero in 2019 (Eddy, 2021).

**Societal aspect**

Despite the imminent technical possibility of implementation, society may not be quite ready for it and political opposition because of the potential for large-scale unemployment among commercial drivers is expected (Wadud, 2017). Whereas, according to the IRU (International Road Transport Union), the European road transport sector is facing the biggest shortage of professional drivers in decades in 2019. According to a survey, the reasons are: the lack of attracting women to work in transport (female drivers make up only 2 per cent of employees in the European transport sector), attracting young people to work in the industry, the poor image of the profession, poor working conditions and being away from home and family for too long. In addition, there is the demographic ageing of the workforce. The average age of a truck driver in Europe is 44 (Kulikowska-Wielgus, 2019).<br/>

In the US, federal agencies have filed plans with the Office of Management and Budget (OMB) for a survey study to assess public appetite for automated driving systems (ADS) in the trucking sector. However, the project, Trucking Fleet Concept of Operations (CONOPS) for Managing Mixed Fleets, has already received public pressure. While the Federal Motor Carrier Safety Administration (FMCSA) received only eight comments on the proposal when it was first requested in November, all opposed it. They cited the "potential loss of jobs for truck drivers and safety concerns arising from ADS testing in the real world" before enough simulator testing has been done. "People who rely on truck driving jobs will become unemployed and will no longer have a way to earn a living and provide for themselves and their families," one comment stated. If OMB approves the survey, a questionnaire will be completed before and after participants have the opportunity to have hands-on experience with ADS truck technology to assess a change in sentiment. The plan is to share these experiences at four "road shows" at existing trucking conferences around the country (Gallagher, 2021).<br/>

Since 2019, a fully electric autonomous truck "T-Pod" from Einride has been driving on a public road in Sweden on behalf of logistics service provider DB Schenker. However, the approval, which was valid until the end of 2020, stipulated that the T-Pod, which weighs up to 26 tonnes, may only travel at 5 km/h, that the approval was only valid for a specific public road in an industrial area and that one employee was assigned to each vehicle as a supervisor who can intervene from a distance via joystick in the event of problems. The truck was highly technical did not have a driver's cab. According to Einride, the e-truck offers space for 15 Euro pallets instead. Its regular maximum speed is capped at 85 km/h. A 280 kWh battery on board is supposed to guarantee a range of 200 km (Of-Allinger et al., 2020; Werwitzke, 2019). <br/>

In manageable, closed areas such as quarries or mines driverless vehicles are already being tested in practice. Volvo Trucks, for example, has six self-driving trucks in operation at a Norwegian limestone plant (Böhm, 2019).


### Relevant initiatives in Austria {-}
In Upper Austria (Gunskirchen), a driverless truck will soon also be on the road in public areas. Over the next three years, it will regularly cover the approximately 600-metre route between the logistics centre of DB Schenker Austria and the engine manufacturer BRP-Rotax. The route runs partly on factory premises and partly in public areas and is to be completed in all weathers. After completion of the test phase, the driverless electric transporter will then cover the route between DB Schenker's logistics centre and the BRP-Rotax company, which is mainly on the public road, completely autonomously from 2023. The electric vehicle is to complete the route in all weathers, whether wind, rain or snow. For freight transport, a classic diesel truck currently travels the same route several times a day (APA, 2021; DB, 2021).

- [Land-oberoesterreich.gv.at](https://www.land-oberoesterreich.gv.at/248885.htm)

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table111, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|ADS in heavy vehicles lead to a reduction in the total cost of ownership of trucks of between 11% and 46%|**+**|Sustainable economic development  (*8,11*)|Ghandriz et al., 2020a|
|Systemic|Reduction in the environmental impact|**+**|Environmental sustainability (*7,12-13,15*)|Kopelias et al., 2020|
|Systemic|ADS enable higher profitability in freight transport|**+**|Sustainable economic development  (*8,11*)|Wadud et al., 2016|
|Systemic|The C-ROADS platform was established to harmonise the use of C-ITS activities across Europe|**+**|Partnership & collaborations (*17*)|European Commission, 2021|')
```

### Technology and societal readiness level {-}
```{r table112, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|6-8|5-7|')
```

### Open questions {-}
1. How significant will the rebound effect be that results from the improvement in freight efficiency?
2. Who is responsible if the technology in semi-automated mode violates a speed limit or other traffic regulations?
3. How should commercial CAVs best be integrated with conventional trucks?

### Further links {-}

- [ec.europa.eu](https://ec.europa.eu/inea/en/horizon-2020/automated-road-transport)

### References {-}

-	Alessandrini, A., Campagna, A., Site, P. D., Filippi, F., & Persia, L. (2015). Automated vehicles and the rethinking of mobility and cities. Transportation Research Procedia, 5, 145–160. https://doi.org/10.1016/j.trpro.2015.01.002
-	Anderson, J., Kalra, N., Stanley, K., Sorensen, P., Samaras, C., & Oluwatola, O. (2016). Autonomous Vehicle Technology: A Guide for Policymakers. In Autonomous Vehicle Technology: A Guide for Policymakers. https://doi.org/10.7249/rr443-2
-	APA. (2021). Selbstfahrender Lkw in OÖ unterwegs | SN.at. Salzburger Nachrichten. https://www.sn.at/wirtschaft/oesterreich/selbstfahrender-lkw-in-ooe-unterwegs-99060673
-	Böhm, M. (2019, November 5). Selbstfahrende Lkws: Die Autonomen rollen an - Transport & Logistik - derStandard.at › Wirtschaft. https://www.derstandard.at/story/2000110697906/selbstfahrende-lkws-die-autonomen-rollen-an
-	Boudway, I. (2021, May 5). Supply Chains Latest: Driverless Trucks a Deal Closer to Freeway. - Bloomberg. Bloomberg. https://www.bloomberg.com/news/newsletters/2021-05-05/supply-chains-latest-driverless-trucks-a-deal-closer-to-freeway
-	Brown, A., Gonder, J., & Repac, B. (2014). An Analysis of Possible Energy Impacts of Automated Vehicles BT  - Road Vehicle Automation (G. Meyer & S. Beiker (eds.); pp. 137–153). Springer International Publishing. https://doi.org/10.1007/978-3-319-05990-7_13
-	Chan, C. Y. (2017). Advancements, prospects, and impacts of automated driving systems. International Journal of Transportation Science and Technology, 6(3), 208–216. https://doi.org/10.1016/j.ijtst.2017.07.008
-	DB. (2021, March 16). Autonomes Fahren: DB Schenker und BRP-Rotax starten Testbetrieb in Österreich. https://www.dbschenker.com/at-de/ueber-uns/presse/corporate-news/autonomes-fahren--db-schenker-und-brp-rotax-starten-testbetrieb-in-oesterreich-688986
-	Eddy, N. (2021, May 8). Driverless trucks worth the long-term investment – Urgent Comms. https://urgentcomm.com/2021/05/10/driverless-trucks-worth-the-long-term-investment/
-	European Commission. (2019). Automated road transport.
-	European Commission. (2021, April 30). Connected and automated mobility | Shaping Europe’s digital future. https://digital-strategy.ec.europa.eu/en/policies/connected-and-automated-mobility
-	Flämig, H. (2016). Autonomous vehicles and autonomous driving in freight transport. In Autonomous Driving: Technical, Legal and Social Aspects (pp. 365–385). Springer Berlin Heidelberg. https://doi.org/10.1007/978-3-662-48847-8_18
-	Gallagher, J. (2021, April 23). FMCSA automated trucks project headed to White House - FreightWaves. Freightwaves. https://www.freightwaves.com/news/fmcsa-automated-trucks-project-headed-to-white-house
-	Gan, M., Qian, Q., Li, D., Ai, Y., & Liu, X. (2021). Capturing the swarm intelligence in truckers: The foundation analysis for future swarm robotics in road freight. Swarm and Evolutionary Computation, 62, 100845. https://doi.org/10.1016/j.swevo.2021.100845
-	Ghandriz, T., Jacobson, B., Laine, L., & Hellgren, J. (2020a). Impact of automated driving systems on road freight transport and electrified propulsion of heavy vehicles. Transportation Research Part C: Emerging Technologies, 115, 102610. https://doi.org/10.1016/j.trc.2020.102610
-	Ghandriz, T., Jacobson, B., Laine, L., & Hellgren, J. (2020b). Optimization data on total cost of ownership for conventional and battery electric heavy vehicles driven by humans and by automated driving systems. Data in Brief, 30. https://doi.org/10.1016/j.dib.2020.105566
-	Harper, C. D., Hendrickson, C. T., & Samaras, C. (2016). Cost and benefit estimates of partially-automated vehicle collision avoidance technologies. Accident Analysis and Prevention, 95, 104–115. https://doi.org/10.1016/j.aap.2016.06.017
-	Kalra, N., & Paddock, S. M. (2016). Driving to safety: How many miles of driving would it take to demonstrate autonomous vehicle reliability? Transportation Research Part A: Policy and Practice, 94, 182–193. https://doi.org/10.1016/j.tra.2016.09.010
-	Khan, A., Harper, C. D., Hendrickson, C. T., & Samaras, C. (2019). Net-societal and net-private benefits of some existing vehicle crash avoidance technologies. Accident Analysis and Prevention, 125, 207–216. https://doi.org/10.1016/j.aap.2019.02.003
-	Kopelias, P., Demiridi, E., Vogiatzis, K., Skabardonis, A., & Zafiropoulou, V. (2020). Connected & autonomous vehicles – Environmental impacts – A review. Science of the Total Environment, 712, 135237. https://doi.org/10.1016/j.scitotenv.2019.135237
-	Kulikowska-Wielgus, A. (2019, March 28). Every fifth truck driver position in Europe is vacant. Soon, there could be twice as many. - Trans.INFO. Trans.Info. https://trans.info/en/every-fifth-truck-driver-position-in-europe-is-vacant-soon-there-could-be-twice-as-many-131212
-	Mersky, A. C., & Samaras, C. (2016). Fuel economy testing of autonomous vehicles. Transportation Research Part C: Emerging Technologies, 65, 31–48. https://doi.org/10.1016/j.trc.2016.01.001
-	Mulholland, E., Teter, J., Cazzola, P., McDonald, Z., & Ó Gallachóir, B. P. (2018). The long haul towards decarbonising road freight – A global assessment to 2050. Applied Energy, 216, 678–693. https://doi.org/10.1016/j.apenergy.2018.01.058
-	Nowakowski, C., Shladover, S. E., & Tan, H. S. (2015). Heavy Vehicle Automation: Human Factors Lessons Learned. Procedia Manufacturing, 3, 2945–2952. https://doi.org/10.1016/j.promfg.2015.07.824
-	Of-Allinger, A., Knecht, J., & Conrad, B. (2020, April 30). Automomer Elektro-Lkw Einride T-Pod: Zulassung in Schweden | AUTO MOTOR UND SPORT. Auto Motr Sport. https://www.auto-motor-und-sport.de/elektroauto/automomer-elektro-lkw-einride-t-pod-strassenzulassung/
-	Paulsen, T. (2018, November 7). Autonomes Fahren: 5 Level zum selbstfahrenden Auto | ADAC. ADAC. https://www.adac.de/rund-ums-fahrzeug/ausstattung-technik-zubehoer/autonomes-fahren/grundlagen/autonomes-fahren-5-stufen/
-	Pribyl, O., Blokpoel, R., & Matowicki, M. (2020). Addressing EU climate targets: Reducing CO2 emissions using cooperative and automated vehicles. Transportation Research Part D: Transport and Environment, 86, 102437. https://doi.org/10.1016/j.trd.2020.102437
-	Shin, S., Roh, H. S., & Hur, S. H. (2018). Technical Trends Related to Intermodal Automated Freight Transport Systems (AFTS) *. Asian Journal of Shipping and Logistics, 34(2), 161–169. https://doi.org/10.1016/j.ajsl.2018.06.013
-	Tsugawa, S., Jeschke, S., & Shladovers, S. E. (2016). A review of truck platooning projects for energy savings. IEEE Transactions on Intelligent Vehicles, 1(1), 68–77. https://doi.org/10.1109/TIV.2016.2577499
-	Wadud, Z. (2017). Fully automated vehicles: A cost of ownership analysis to inform early adoption. Transportation Research Part A: Policy and Practice, 101, 163–176. https://doi.org/10.1016/j.tra.2017.05.005
-	Wadud, Z., MacKenzie, D., & Leiby, P. (2016). Help or hindrance? The travel, energy and carbon impacts of highly automated vehicles. Transportation Research Part A: Policy and Practice, 86, 1–18. https://doi.org/10.1016/j.tra.2015.12.001
-	Werwitzke, C. (2019, May 17). Schweden: T-Pod erhält Zulassung für öffentliche Straße - electrive.net. Electrive.Net. https://www.electrive.net/2019/05/17/schweden-t-pod-erhaelt-zulassung-fuer-oeffentliche-strasse/

