# On-board technology for connected and automated vehicles {#onboard}

## Advanced driver assistance system

## Parking assistance system

### Definition {-}
Nowadays, parking a car becomes an increasingly mundane task due to the growing size of the vehicles which consequently reduces visibility to the rear and front. Moreover, the number of vehicles on the road continues to increase which makes finding parking spaces more difficult. To overcome this challenge, efficient and advanced parking techniques are needed, such as finding the right parking space and parking cars efficiently (Khalid et al., 2021). <br/><br/>
While in the 1990s parking assistance systems were not considered necessary, systems that support the parking process are now standard in new vehicles and have a high acceptance rate. While parking assistance systems of the first generations were mainly informative systems, today these systems help to find a suitable parking space and support the steering during the parking manoeuvre. Future systems will become even more autonomous, up to a point where they will find their parking space without human intervention (Valet Parking) (Gotzig, 2016). <br/><br/>
With respect to safety issues, parking assistance systems can reduce insurance claims that occur during parking manoeuvres by 30% (ADAC, 2020). They also provide a useful application in the form of emergency brake assistant that prevents accidents with other road users when reversing (ADAC, 2019). <br/><br/>
Modern Parking assistants work with two sensor concepts: ultrasonic sensors for the close range at the rear (often already installed as "parking beepers") and radar sensors with a longer range at the side in the bumper (ADAC, 2019). Additionally, there are also new releases with 360° view around the car (Land Rover, no date; Mercedes-Benz, no date). However, there are no tests for this at the moment.

### Key stakeholders {-}
- **Affected**: Car Drivers, Traffic participants, Insurers
- **Responsible**: Car manufacturers

### Current state of art in research {-}
The research on on-board parking features is almost entirely owned by the car manufacturers. Research tends to focus on larger-scale issues such as car parking location tracking, space booking systems using IoT and congestion caused by *parking cruises*.<br/> <br/>
Additionally, tests are carried out on the merger between emergency brake assistant and parking assistant to overcome the difficulties with obstacles detection and reduce the collision rates in this specific situation. The ADAC tested the AEB (Autonomous Emergency Braking) systems from Mercedes, Volvo, BMW, Seat and Skoda in three test scenarios: *(1)* a pedestrian dummy stands behind a car or walks past, *(2)* a car parks in the direction of travel and *(3)* cyclists as well as cars drive past crosswise. <br/> <br/>
BMW was the best at reacting to all situations with radar and ultrasound - with some drop-outs, especially with moving pedestrians or cross-traffic. The Mercedes, on the other hand, only used its side radar sensors for reverse braking and thus did not recognise stationary vehicles at all. The VW system from Skoda and Seat had radar and ultrasound, but moving pedestrians were only detected randomly or not at all.<br/><br/> 
Overall, these tests showed that the automatic braking parking assistants have a lot of potential, but are far from optimal. Even the system of the front-runner does not yet work 100% reliably. Even the low-cost ultrasonic sensors, however, can be very effective and even prevent pedestrian collisions, as the BMW showed in the test. Therefore, it is crucial that manufacturers equip their vehicles with an effective AEB system as a standard. The necessary technology is already available in most passenger cars where the rear ultrasonic sensors would need to be linked to the braking function (ADAC, 2019).
The greatest difficulties are encountered in pedestrian detection, in the scenarios in which the risk of personal injury is potentially the highest. Some of the tested vehicles recognised the dangerous situation too late or not at all.

### Current state of art in practice {-}
Nearly 50% of the cars in Germany are now equipped with parking assistance. Nonetheless, the results of a study by HUK-Coburg, Germany's largest motor insurer with eleven million insured cars, show that the number of fender-benders has not decreased and the damage costs have even risen slightly. The reason for this is the damage to the expensive Park Distance Control (PDC) sensor embedded in the bumper when vehicle collides with an obstacle while parking (Focus Online, 2017).<br/> <br/>
In 2017, 570 accidents with personal injury occurred in Austria when reversing with passenger cars. There were no fatalities, but around 290 people were injured, 60 of them seriously. In addition, there is a large amount of property damage due to obstacles being overlooked (ÖAMTC, 2019).<br/> <br/>
According to different sources, between 23% and 46% of the cars in Germany are equipped with an on-board parking assistant (Focus Online, 2017). In contrast, only 13% of the vehicles are equipped with an emergency brake assistant that automatically brakes in the event of an imminent collision with the vehicle in front or even detects pedestrians and cyclists. Further, a survey conducted on 1000 respondents by German Road Safety Council showed that 85% of the sample found the emergency brake assistant as highly useful and 65% considered parking assistant highly beneficial (Handelsblatt, 2016).<br/> <br/>
Currently, the most advanced technology with respect to parking assistance is offered by Tesla S, where the system is fully autonomous and allows the car to drive itself out of a tight space without the driver’s intervention. But it is expected that the technology will be available from other manufacturers too and will improve rapidly in the coming years. One reason for this is that parking assistants with brake intervention will become part of the European vehicle test programme *Euro NCAP* from 2020. In the past, it has been shown that the inclusion of active and passive passenger car safety systems in this programme quickly increases the rate at which vehicles are equipped (ÖAMTC, 2019).

 
### Relevant initiatives in Austria {-}

- [oeamtc.at](https://www.oeamtc.at/tests/assistenzsystemtest/fuenf-parkassistenten-mit-notbremssystem-im-test-31717317) 


### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table56, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Road accidents when parking reduced|**+**|Health & Wellbeing (*3*)|ADAC, 2019|
|Individual|Higher costs in case of collision|**-**|Sustainable economic development (*8,11*)|Focus Online, 2017|
|Systemic|New designs tested |          **+**|Innovation & Infrastructure  (*9*)|Euro NCAP, 2021|')
```

### Technology and societal readiness level {-}

```{r table57, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|8-9|')
```

### Open questions {-}
1. How can acceptance and usability rates be increased in drivers especially in elderly group?
2. What are the legal implications for the use of assisted parking for both drivers and car manufacturers?

### References {-}
- ADAC (2019) Parkassistenten im Test: Noch nicht gut genug. Available at: https://presse.adac.de/meldungen/adac-ev/tests/parkassistent.html (Accessed: 24 February 2021).
- ADAC (2020) Fahrerassistenzsysteme im Überblick | ADAC. Available at: https://www.adac.de/rund-ums-fahrzeug/ausstattung-technik-zubehoer/assistenzsysteme/fahrerassistenzsysteme/ (Accessed: 24 February 2021).
- Euro NCAP. (2021). http://www.euroncap.com (Accessed: 26 February 2021) 
- Focus Online (2017) Einparkhilfen bringen nichts - sondern verursachen mehr Schaden - FOCUS Online. Available at: https://www.focus.de/auto/news/untersuchung-der-huk-coburg-studie-einparkhilfe-hilft-nicht-sie-versucht-eher-mehr-schaden_id_7035248.html (Accessed: 24 February 2021).
- Gotzig, H. (2016) ‘Parking Assistance’, in Winner, H. et al. (eds) Handbook of Driver Assistance Systems: Basic Information, Components and Systems for Active Safety and Comfort. Cham: Springer International Publishing, pp. 1077–1092. doi: 10.1007/978-3-319-12352-3_45.
- Handelsblatt (2016) Umfrage zu Assistenzsystemen : Einparkassistent nützlich, Notbremsassistent nützlicher. Available at: https://www.handelsblatt.com/auto/nachrichten/umfrage-zu-assistenzsystemen-einparkassistent-nuetzlich-notbremsassistent-nuetzlicher/13404708.html?ticket=ST-4955434-naHBLAxI66h66nbvrBXF-ap1 (Accessed: 24 February 2021).
- Khalid, M. et al. (2021) ‘From smart parking towards autonomous valet parking: A survey, challenges and future Works’, - Journal of Network and Computer Applications, 175, p. 102935. doi: https://doi.org/10.1016/j.jnca.2020.102935.
- Land Rover (no date) Parking Assistance | InControl | Land Rover UK. Available at: https://www.landrover.co.uk/incontrol/driver-safety-and-assistance/parking-assistance.html (Accessed: 25 February 2021).
- Margreiter, M., Mayer, P., Alpas, M., & Vlahogianni, E. (2017). Driver’s Willingness to Use Parking Assistance Tools and their Expectations: A Case Study for the Cities of Munich and Athens. In 8th International Congress on Transportation Research.
- Mercedes-Benz (no date) Mercedes-Benz X-Klasse: Park-Paket mit 360°-Kamera. Available at: https://www.mercedes-benz.at/passengercars/mercedes-benz-cars/models/x-class/x-class-pickup/facts-and-lines/equipment-packages/360-camera.html (Accessed: 25 February 2021).
- Tsugawa, S. (2006). Trends and issues in safe driver assistance systems: Driver acceptance and assistance for elderly drivers. IATSS research, 30(2), 6-18.
- ÖAMTC (2019) ÖAMTC: Fünf Parkassistenten mit Notbremssystem im Test (+ Fotos, + Grafik) | ÖAMTC, 14.05.2019. Available at: https://www.ots.at/presseaussendung/OTS_20190514_OTS0021/oeamtc-fuenf-parkassistenten-mit-notbremssystem-im-test-fotos-grafik (Accessed: 25 February 2021).


## Lane keeping

## Distane keeping

## Crash avoidance

## Mainteinance assistance

## Digital maps 

## Electronic Horizon

### Synonyms {-}
*Connected horizon, e-horizon, EH*

### Definition {-}
Modern vehicles are currently equipped with different sensors to increase the safety and comfort of driving. Nonetheless, these built-in sensors can usually operate for a short distance of a few hundred meters. On the other hand, digital maps within the system provide information about the road such as its geometry, number of lanes, speed limits or traffic signals. Therefore, the merger between the functions of the sensors and the data provided by the map gives rise to so-called *Electronic Horizon* applications, that are able to provide information from the sensors within an extended range and allow for earlier preparation for a situation ahead (Ress et al., 2008).<br/><br/>

The benefits of the e-horizon include increased comfort due to partially automated functions, reduced accident risk thanks to increased information about road situations ahead, lower fuel consumption through predictive driving and longer driving ranges in the case of electric and hybrid vehicles due to optimized energy management (Bosch-mobility-solutions.com, 2021). Nonetheless, the e-horizon still faces certain challenges such as (Grewe et al., 2017):

-	*Speed of data*: in the high-speed mobility data or service request may not be answered or provided fast enough before the vehicle reconnects to another network entry point;
-	*Amount of data*: fully automated driving is predicted to produce 4 TB of data daily (Nelson, 2016);
-	*Scalability*: the system design is relatively simple for small number of participating cars but it may experience difficulties as soon as number of vehicles and/or applications rise; 
-	*Security and privacy issues*: as with other connected systems, the issue with data protection arises to ensure security and quality for customers.
 
### Key stakeholders {-}
- **Affected**: Car drivers, Traffic participants, Insurers
- **Responsible**: Car manufacturers

### Current state of art in research {-}
Profound test of software and hardware are performed under various condition such as different speed and journey breaks to gather data on the entire information chain from GPS positioning to e-Horizon generation. Based on the collected information the engineers optimize the workflow before moving to real test drive. In real test drives, the performance of the applications is frequently checked in the borderline scenarios or with deliberate errors (Ludwig, 2013). Moreover, driving simulation is used to test the effects of sensors failures (Elgharbawy et al., 2019)

### Current state of art in practice {-}
Electronic horizon is relatively well-established technology in terms of both passenger and commercial vehciles and it is offered by several manufacturers including [BOSCH](https://www.bosch-mobility-solutions.com/en/products-and-services/passenger-cars-and-light-commercial-vehicles/connectivity-solutions/connected-horizon/), [TomTom]( https://www.tomtom.com/products/autostream/) in cooperation with [Electrobit]( https://www.elektrobit.com/products/automated-driving/eb-robinos/predictor/) and [Continental](https://www.continental.com/en/press/press-releases/continental-ehorizon-and-previewesc-systems-180356) together with [IBM]( https://www.continental.com/en/press/press-releases/continental-and-ibm-enter-connected-vehicle-collaboration-8438) and [HERE]( https://360.here.com/2017/01/04/introducing-here-electronic-horizon/) (Continental, 2019). 
 
### Relevant initiatives in Austria {-}

- [gsv.co.at](https://gsv.co.at/wp-content/uploads/2015_05_06_Foersterling.pdf ) 
 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table63, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Accident risk reduced|**+**|Health & Wellbeing (*3*)|Bosch-mobility-solutions.com, 2021; Continental, 2019|
|Individual|Lower fuel consumption|**+**|Sustainable economic development (*7,12-13,15*)|Continental, 2019|
|Systemic|Increased driving range for hybrid and electric vehicles|**+**|Sustainable economic development (*7,12-13,15*)|Bosch-mobility-solutions.com, 2021|
|Systemic|Innovative research towards full automation|**+**|Innovation & Infrastructure (*9*) |Foersterling, 2015|
|Systemic|Collaborations between automotive and technology companies|**+**|Partnership & collaborations (*17*)|Continental, 2019; Foersterling, 2015|')
```

### Technology and societal readiness level {-}

```{r table64, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|7-9|')
```

### Open questions {-}

1. How to handle massive mobility of vehicles?
2. How do current approaches scale when the number of participants and services within this system rise?
3. How to fulfill application- or user-specific quality requirements?
4. How to ensure availability of services and data when these are deployed outside of the vehicle? (Grewe et al., 2017)


### Further links {-}

- [autotechreview.com](https://autotechreview.com/media/attachments/Technology_1_1.pdf)
- [here.com](https://www.here.com/sites/g/files/odxslz166/files/2018-11/HERE_Electronic_Horizon_one_pager.pdf)
- [scitepress.org](https://www.scitepress.org/Papers/2008/15080/15080.pdf)
- [bosch-mobility-solutions.com](https://www.bosch-mobility-solutions.com/en/products-and-services/passenger-cars-and-light-commercial-vehicles/connectivity-solutions/connected-horizon/)
- [elektrobit.com](https://www.elektrobit.com/newsroom/tomtom-elektrobit-join-forces-electronic-horizon-automated-driving/)
- [itsinternational.com](https://www.itsinternational.com/its10/products/continental-debuts-electronic-horizon)
- [strategic-partnering.net](http://strategic-partnering.net/continental-ibm-enter-connected-vehicle-collaboration/)


### References {-}
- Bosch-mobility-solutions.com. (2021). Connected Horizon. [online] Available at: <https://www.bosch-mobility-solutions.com/en/products-and-services/passenger-cars-and-light-commercial-vehicles/connectivity-solutions/connected-horizon/> [Accessed 1 March 2021].
- Continental. (2019). Increased Safety Thanks to Anticipatory Technology: The Continental eHorizon and PreviewESC Systems. [online] Available at: <https://www.continental.com/en/press/press-releases/continental-ehorizon-and-previewesc-systems-180356> [Accessed 1 March 2021].
- Elgharbawy, M., Schwarzhaupt, A., Arenskrieger, R., Elsayed, H., Frey, M., & Gauterin, F. (2019). A testing framework for predictive driving features with an electronic Horizon. Transportation research part F: traffic psychology and behaviour, 61, 291-304.
- Försterling, F. (2015) Electronic Horizon How the Cloud improves the connected vehicle. [online] Available at: https://gsv.co.at/wp-content/uploads/2015_05_06_Foersterling.pdf [Accessed 1 March 2021].
- Grewe, D., Wagner, M., Arumaithurai, M., Psaras, I., & Kutscher, D. (2017, August). Information-centric mobile edge computing for connected vehicle environments: Challenges and research directions. In Proceedings of the Workshop on Mobile Edge Communications (pp. 7-12).
- Ludwig, J. (2013). Electronic Horizon—Forward-Looking Safety Systems. Auto Tech Review, 2(6), 44-48.
- Nelson, P. (2016). Just one autonomous car will use 4,000 GB of data/day. (December 2016). http://www.networkworld.com/article/3147892/internet/ one-autonomous-car-will-use-4000-gb-of-dataday.html
- Ress, C., Etemad, A., Kuck, D., & Requejo, J. (2008). Electronic horizon-providing digital map data for ADAS applications. In Proceedings of the 2nd International Workshop on Intelligent Vehicle Control Systems (IVCS) (pp. 40-49).




## Emergency call

### Synonyms {-}
*Public-Safety Answering-Point (PSAP), E-call*

### Definition {-}
The E-Call is a system that provides an automated message to the emergency services following a road crash which includes the precise crash location. The in-vehicle e-Call is an emergency call generated either manually by the vehicle occupants or automatically via activation of in-vehicle sensors after a crash. The aim of this system is to reduce the time between the accident and the provision of medical services.As soon as the in-vehicle eCall device is activated, it establishes an emergency call carrying both voice and data directly to the nearest emergency services (normally the nearest 112 Public Safety Answering Point, PSAP). The voice call enables vehicle occupants to communicate with the trained eCall operator. At the same time, a minimum set of data is sent to the eCall operator receiving the voice call (European Commission, 2020a). <br/>
The minimum data provided during an E-call are (Kroher, 2020):

-	time of accident
-	the exact coordinates of the accident location 
-	direction of travel (important on motorways and in tunnels) 
-	the last two vehicle positions
-	vehicle ID and vehicle class
-	type of drive (e.g. petrol, electric)
-	service provider ID
-	number of occupants (based on seat belts worn) 
-	whether the emergency call was triggered automatically or manually 

### Key stakeholders {-}
- **Affected**: Car Drivers, Emergency Services
- **Responsible**: Road Infrastructure Agencies, Local and National Governments, Automotive Companies, Policy Makers

### Current state of art in research {-}
The research on the emergency calls looks at its efficiency in fatalities reduction as a consequence of road accidents. For example, a Swedish study concluded that 49% of those who died in fatal road accidents could have survived. In particular, 5% of them would have survived if they had been located more quickly, 12% if they had been transported to hospital more quickly and 32% if they had been taken to an advanced trauma centre more quickly (Henriksson et al., 2001). Therefore, the evidence shows that the potential of e-calls in saving lives is significant. Moreover, a study by Virtanen et al. (2006) estimated that the use of an e-Call could reduce between 4-8% of road fatalities and 5-10% of motor vehicle occupant deaths in Finland. In other European countries, these estimates fluctuate between 2%-7%, whereas the estimate for whole EU area with 25 member states reaches up to 15% in fatalities reduction. The European Commission estimates that a pan-European eCall system has the potential to prevent up to 2500 deaths per year in the EU-25 if fully implemented (European Commission, 2020a). 
Another strand of research focuses on the testing of different designs of e-Call systems and aims at its standardisation across European Union member countries. Uniform legislation concerns in particular, the communication protocol and the amount of data that is provided by an e-call, as well as its content and format (European Commission, 2020a).

### Current state of art in practice {-}
Since the end of March 2018, manufacturers in the EU have to include E-Call in all new models of passenger cars and light commercial vehicles (Bundesministerium für Verkehr, no date). However, even if the vehicles have an e-Call system installed, some car producers additionally install their own emergency call systems. The use of manufacturer-specific e-call system rather than a standardised one, gives car manufacturers an advantage and creates an opportunity to trade the data to external ‘accident related’ providers such as towing, or offer services in terms of post-accident customer care such as vehicle reparation or car replacement.<br/><br/>
Regardless of the benefits for the car producers the results of the crash tests carried out by the European new car assessment programme *Euro NCAP* at the *ADAC Technik Zentrum* are alarming. According to the tests, manufacturer emergency calls were sometimes answered by the call centre only 58 seconds after the airbags had been deployed. This shows that the use of manufacturer-specific e-call, in fact, delays the medical service provision (relative to 112 e-Call), where the position of the car must first be determined from the transmitted location data in order to then forward it to the actual responsible rescue control centre on site. After that, only the rescue centre sends out the ambulance. In the event of an accident, the valuable time would be lost due to this indirect procedure.<br/><br/>
Nevertheless, many vehicles still use this manufacturer-specific e-call, permitted by the EU, which  first informs the car manufacturer's control centre or its service provider, and not 112, directly. According to an ADAC survey, mainly the German manufacturers prefer to use a manufacturer-specific emergency call instead of 112 e-Call, while other car producers that participated in the survey relied only on 112 e-Call.<br/><br/>
On the other hand, due to data protection concerns, the system is viewed sceptically by many car drivers. However, in the view of the ADAC, there is no reason for this. 112 e-Call only logs into the mobile network after a serious accident and then sends data to the rescue coordination centre, not to the manufacturer. The 112 e-Call also does not record any data in the car. According to the ADAC, there are now the following steps for legislators and manufacturers to take (Kroher, 2020):

-	112-eCall should be mandatory for all new vehicles, not only for new type approvals.
-	The manufacturer-specific emergency call already installed in many models should be convertible to 112-eCall without major effort.
-	In order to better inform the driver about the differences between 112-eCall and the manufacturer's emergency call, a detailed description of the function including the content of the MSD (Minimum Set of Data, the transmitted data) should be available in the on-board manual and also in the display of the vehicle.
-	If 112-eCall and manufacturer emergency call are available in parallel in the vehicle, the driver should have the right to choose his preferred service provider. As many consumers are uncertain about this, it would be advisable to pre-set 112-eCall in the car by default.
-	When using the manufacturer's emergency call, there should be no delay in reporting the accident to the rescue coordination centre in order to enable the fastest possible assistance.
-	The data set (MSD) transmitted in the 112-eCall should be expanded to include information (e.g. acceleration values) that enable rescue control centres to automatically predict the type and severity of injuries and thus adequately alert rescue resources.
-	In order to be able to use an eCall technology installed in the vehicle fleet over the lifetime of the vehicles in an emergency, it is necessary to maintain the 2G/3G networks.
 
### Relevant initiatives in Austria {-}
In Austria, the handling of emergency calls falls under the responsibility of the Ministry of the Interior.
In the project *E-Call Austria*, the E-Call system is to be implemented in Austria. The measures include the testing, implementation and certification of the emergency call answering points in Austria. The new Public-Safety Answering Points (PSAP) in all nine federal states will then be in line with the requirements of EU regulations (Regulation (EU) No 305/2013; provision of an EU-wide e-call service).

- [bmi.gv.at](https://www.bmi.gv.at/209/start.aspx) 
- [oeamtc.at](https://www.oeamtc.at/thema/ecall/) 
- [austriatech.at](https://austriatech.at/de/wenn-es-schnell-gehen-muss-ecall-macht-die-strasse-sicherer/) 
- [oe1.orf.at](https://oe1.orf.at/artikel/644731/eCall-der-automatische-Autonotruf-und-seine-Nachteile) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table34, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Road fatalities reduced|**+**|Health & Wellbeing (*3*)|European Commission, 2020a|
|Systemic|         Disporportionally high fitting costs of e-Call|          **-**|Sustainable economic development (*8,11*)|European Commission, 2020a|
|Systemic|Universal emergency number in EU (112) |          **+**|Partnership & collaborations (*17*)|European Commission, 2020b|')
```

### Technology and societal readiness level {-}

```{r table35, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|6-9|')
```

### Open questions {-}
1. What is the current state of implementation in individual EU member states?
2. Are car buyers sufficiently informed about the difference between manufacturer-specific emergency call and 112 e-Call? And if not, how to tackle this?
3. What is the time horizon to see a full impact of an e-Call, given the proportion of 'older' cars on the roads which do not have e-call fitted?

### Further links {-}

-	[ec.europa.eu](https://ec.europa.eu/transport/road_safety/specialist/knowledge/esave/esafety_measures_unknown_safety_effects/ecall_en) 

### References {-}
- Bundesministerium für Verkehr, I. und T. (BMVIT) (no date) eCall Austria. Available at: https://www.bmi.gv.at/209/start.aspx (Accessed: 10 February 2021).
- European Commission (2020a) Air | Mobility and Transport, European Commission. Available at: https://ec.europa.eu/transport/road_safety/specialist/knowledge/esave/esafety_measures_unknown_safety_effects/ecall_en (Accessed: 10 February 2021).
- European Commission (2020b) eCall – Kraftfahrzeugassistenzsystem für Notrufe an die europäische Notrufnummer 112 - Your Europe. Available at: https://europa.eu/youreurope/citizens/travel/security-and-emergencies/emergency-assistance-vehicles-ecall/index_de.htm (Accessed: 10 February 2021).
- Henriksson, E. M., Oström, M. Eriksson, A. (2001) Preventability of vehicle-related fatalities. Accident Analysis and Prevention, 467-475
- Kroher, T. (2020) eCall: Probleme beim automatischen Notrufsystem | ADAC. Available at: https://www.adac.de/rund-ums-fahrzeug/unfall-schaden-panne/unfall/ecall-herstellernotruf/ (Accessed: 10 February 2021).
- Virtanen, A., Schirokoff, J., Luoma and R. Kulmala (2006) Impacts of an automatic emergency call system on accident consequences, Ministry of Transport and Communications Finland Finnish R&D Programme on Real-Time Transport Information AINO


