# Freight and commercial transport {#freight}

## Automated road freight {#automated_road_freight}

## Tracking and tracing of dangerous goods {#dangerous_goods}

## Intermodal Freight {#intermodal_freight}

### Synonyms {-}
*unaccompanied combined transport (UCT), intermodal transport unit (ITE)* 

### Definition {-}
Intermodal freight refers to the transport of goods in an intermodal transport unit (ITE) with at least two different modes of transport (e.g. road, rail, inland waterway). ITEs are containers or swap bodies, accompanied or unaccompanied road freight vehicles and trailers of road freight vehicles. The essential feature of intermodal transport is that, when the transport unit is changed to another mode of transport, there is no transhipment of goods, i.e. the entire ITE is always loaded onto another mode of transport (Rudlof et al., 2018).<br/>

In terms of the advantages the intermodal freight is crucial in reducing the emissions where transport accounts for almost a quarter of the European Union's green house gases (GHG) emissions, of which road transport accounts for 74%. In particular, freight transport by road is an important source of emissions, as heavy-duty vehicles are responsible for 6% of the European Union's total emissions. Therefore, reducing carbon emissions associated with freight transport by road is a priority to meet current policy emission reduction targets, such as reducing GHG emissions by 80% by 2050 under the terms of the Paris Agreement. Other possible solutions to this problem include technological improvements, such as the development of more efficient and near-zero emission engines, and management improvements, such as reducing and optimising transport distances and shifting traffic from high-emission to low-emission vehicles, for example from trucks to trains or ships (European Commission, no date). Nonetheless, intermodal freight is frequently associated with slower speed, lack of reliability and increased potential for damages of the good.


### Key stakeholders {-}

- **Affected**: Truck drivers, Freight companies, Rail companies, Freight terminals, Highway users
- **Responsible**: National Governments, International authorities

### Current state of art in research {-}
Research focuses on the ecological perspective of intermodal transport (Heinold & Meisel, 2020) as well as automated intermodal freight transport systems that aim to reduce GHG emissions, minimise infrastructure and logistics costs and reduce traffic congestion (Shin et al., 2018). Moreover, some papers, for example Saeedi et al. (2019) look at the technical efficiency of the intermodal transport. 

### Current state of art in practice {-}
Over the last two decades, EU transport policy has promoted intermodal freight transport (by rail or inland waterways). In 2011, the European Commission set a target of shifting 30% of freight transported further than 300 km by road to other modes such as rail or waterways by 2030, and more than 50% by 2050. Despite significant investments (about €28 billion in funding for rail projects between 2007 and 2013) and the priority shift of freight from road to intermodal freight transport (IFT), EU policy for intermodal transport has not achieved significant improvements (European Court of Auditors, 2016). The performance of an IFT service is attributed to two main factors: the performance of the different parts of the chain and the cooperation and harmony between these parts (Saeedi et al., 2019). <br/>

Since this special type of freight transport has become increasingly important in recent years from an economic and political point of view, Eurostat has set up the Task Force on Intermodal Transport Statistics on this topic. The task force, in which Statistics Austria (STAT) also participated, was to evaluate how meaningful and high-quality data on intermodal transport could be compiled at European level for road and rail transport as well as inland and maritime navigation without imposing additional burdens on the member states of the European Union. It was investigated whether in individual countries, in addition to the data on intermodal transport already to be collected according to EU regulations,  information is available that could provide a better picture of transport with ITE. This assessment showed that information on intermodal transport is available in many different forms across countries. Possible indicators can, therefore, only be compiled in a limited form at EU level, as their characteristics differ with regard to the individual modes of transport and also show inconsistencies due to methodological reasons. For example, there are differences in the weights to be collected or the type and size of the ITEs. Furthermore, information on the source and destination regions of transported containers is not available, nor is information on transport chains. This is because the transports on individual modes are considered independent of each other. In the context of the Task Force it also turned out that, compared to other Member States, most information on intermodal transport at national level is available in Germany. The Federal Statistical Office in Germany (Destatis) collects data on the transport of goods in containers, swap bodies, road freight vehicles and trailers of road freight vehicles on roads, railways, rivers and seas. For other countries, one of the basic assumptions is that only a small amount of ITE is transported by road freight vehicles, mainly over short distances (Rudlof et al., 2018). <br/>

Some foreign developments in the field of automated freight transport system technologies have entered pilot operations after finalising their concept designs. In other cases, further developments have been suspended due to economic challenges such as high initial investment. In the case of FSS, an unmanned automated freight transport system was developed in the US and considered the most advanced of its kind in terms of its development progress. The system is currently in test operation after completing the construction of a test bed with a 100 m linear track. Overall, in the advanced countries, it is considered that the developments have reached a point where they will be subjected to the validation processes for commercialisation after the concept phase is completed. Upon completion of further testing, product launches are expected within the next 3 to 4 years (Shin etal., 2018).


### Relevant initiatives in Austria {-}
In Austria, there is a special support programme for unaccompanied combined transport (UCT) by rail, which is intended to contribute to a shift of freight transport to this mode of transport. The Rail Infrastructure Service Company ([SCHIG](https://www.schig.com/)) is responsible for handling these subsidies on behalf of the BMVIT.<br/>

A survey of the terminals as well as the analysis of the SCHIG data sets on the UCT funding programme have shown that no supplementary data on intermodal transport is currently available in Austria. Without new and specially designed surveys, it is currently not possible to compile comprehensive statistics on intermodal transport across all modes of transport, which would above all also allow statements on transport chains. One way of obtaining data could be, for example, to oblige the terminals where intermodal transport takes place to keep records and make them available to the statistics. However, the necessary legal basis for this would first have to be created.<br/>

Intermodal transport in Austria in 2016 was particularly important in the area of rail freight transport with a share of 22.3%. In contrast, only 2.6% of the transport volume on the road was carried in intermodal transport units, and intermodal transport was of no significance for the inland waterway mode of transport, as only empty containers were transported on inland vessels (Rudlof et al., 2018).

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table76, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Potential for reduction in emissions|**+**|Environmental sustainability  (*7,12-13,15*)|European Commission, no date|
|Systemic|Lack of standarisation in information on intermodal transport across EU|**~**|Partnership & collaborations (*17*)|Rudlof et al., 2018|')
```

### Technology and societal readiness level {-}

```{r table77, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|8-9|')
```

### Open questions {-}
1. How can coordination between multi-level decision makers be improved to make intermodal performance more efficient?
2. Given that drayage operations are a large proportion of total cost in intermodal freight, how they can be performed more efficiently?
3. How the standarisation in information collection, storage and sharing procedures can be achieved across the EU conutries?


### Further links {-}

- [eca.europa.eu](https://www.eca.europa.eu/en/Pages/DocItem.aspx?did=36398)
- [transportgeography.org](https://transportgeography.org/contents/chapter5/intermodal-transportation-containerization/)

### References {-}
-	European Commission. (n.d.). Transport emissions | Climate Action. Retrieved 11 March 2021, from https://ec.europa.eu/clima/policies/transport_en
-	European Court of Auditors. (2016). Rail freight transport in the EU: still not on the right track (Issue 08). https://doi.org/10.2865/53961
-	Heinold, A., & Meisel, F. (2020). Emission limits and emission allocation schemes in intermodal freight transportation. Transportation Research Part E: Logistics and Transportation Review, 141, 101963. https://doi.org/10.1016/j.tre.2020.101963
-	Rudlof, M., Karner, T., & Fleck, S. (2018). Intermodaler Verkehr in Österreich. 87–95.
-	Saeedi, H., Behdani, B., Wiegmans, B., & Zuidwijk, R. (2019). Assessing the technical efficiency of intermodal freight transport chains using a modified network DEA approach. Transportation Research Part E: Logistics and Transportation Review, 126, 66–86. https://doi.org/10.1016/j.tre.2019.04.003
-	Shin, S., Roh, H. S., & Hur, S. H. (2018). Technical Trends Related to Intermodal Automated Freight Transport Systems (AFTS) *. Asian Journal of Shipping and Logistics, 34(2), 161–169. https://doi.org/10.1016/j.ajsl.2018.06.013


## Real-time disruption management and route planning {#disruption_management}

## Urban Deliveries {#urban_delivery}

## Intelligent truck parking {#intelligent_truck_park}

### Synonyms {-}
*ITP, Information on Truck Parking*

### Definition {-}
Nowadays, there is a significant increase in transport volume, especially in the international road freight transport. In particular, with the enlargement of the EU member states, the interdependence of the member states has increased as well as the ratio of the exchange of goods and thus the increase in need for transport services (Gnap & Kubíková, 2020). Consequently, the higher number of vehicles and associated traffic lead to difficulties in meeting deadlines for loading and unloading goods as well as the obligation to take mandatory safety breaks and rest periods. It also amplifies the issue with lack of parking spaces along the route for the Heavy Goods Vehicle (HGV) which has a negative impact on logistics and creates assumptions for increasing logistics costs (Lozia & Kulma, 2014). <br/><br/>
Therefore, the European Union is striving to support the development of the Trans-European Transport Network (TEN-T) through structural funds. The TEN-T network is intended to connect the member states from east to west and from north to south. Unfortunately, the rate of road construction in some countries is far below the rate of growth of transport services. The abolishment of in-vehicle resting and other upcoming changes in the so-called "road package" will increase the need for more parking spaces and associated equipment for drivers, especially on motorways. This fact will have a significant impact on the pressure on vehicle parking spaces when the maximum prescribed driving time is observed and there is a need to stop a vehicle and take a safety break or a daily/weekly rest period (Carrese et al., 2014). On top of this, there is a clear need for charging and refuelling stations for alternative fuels at rest areas to ensure smooth transit of vehicles using alternative fuels. These should be in line with the implementation of [Directive 2014/94/EU](https://eur-lex.europa.eu/legal-content/en/TXT/?uri=CELEX%3A32014L0094) of the European Parliament. <br/><br/>
Therefore, to attempt addressing these issues, an Intelligent Truck Parking (ITP) emerged which is an Intelligent Transportation Systems service concerned with the management of information related to HGV parking operations. ITP collects data about parking facilities (services and availability), processes the data, e.g. occupancy level, and communicates the information to drivers (in real time) in order to help them easily access such a facility. ITP enables efficient parking management and potentially improve the needs of transport stakeholders through the provision of information such as available parking spaces and additional services (Sochor & Mbiydzenyuy, 2013). <br/><br/>
Good examples of the efficient use of the information system can be found in Germany or Austria. These systems, as part of the intelligent transport systems, offer services that make transport easier for drivers, e.g. searching for free parking spaces at rest areas located along motorways and motorways (Gnap & Kubíková, 2020).

### Key stakeholders {-}

- **Affected**: Drivers, Transport companies, Freight forwarders, Shippers and Insurers
- **Responsible**: National Governments, Fright Companies, International authorities (eg. EU)

### Current state of art in research {-}
In the literature mostly, regional plans for specific areas in a certain country can be found. For example Carrese et al., (2014) for the Lazio Region in Italy and Gnap & Kubíková, (2020) for the Slovak Republic. However, safety aspects are also highlighted, such as a classification scheme of parking spaces for trucks in long-distance traffic. Again, the data is limited to a specific region (Peel Region in Ontario, Canada) but efforts have been made to develop a generally applicable scheme (Nevland et al., 2020).

### Current state of art in practice {-}
A common European park strategy was developed in 2014 by the European organisation SETPOS within the [SETPOS]( https://ec.europa.eu/transport/sites/transport/files/modes/road/parking/doc/2010_04_28_setpos_project_handbook.pdf) project. The project had four objectives:

-	The assessment and validation of the requirements of stakeholders, i.e. drivers, dispatchers, hauliers, rest area operators, insurers, authorities and shippers;
-	The formulation of a common set of standards for secured parking;
-	The establishment of a number of pilot parking areas in cross-border regions to validate and demonstrate the standard set;
-	To set up an information, guidance and reservation platform for all types of truck parking.

The [LABEL]( https://ec.europa.eu/transport/sites/transport/files/modes/road/parking/doc/handbook_for_labelling.pdf ) project, which follows the SETPOS, aims to increase safety and quality standards for truck parking in Europe by (Carrese et al., 2014):

-	Introducing a European standard certification scheme for truck parking areas;
-	developing an online database for users to provide financial benefits to certified sites.

Every day, billions of euros worth of goods are transported on the trans-European road network, the backbone of the EU economy. Despite the sector's successful performance with respect to volume, it faces a significant number of issues, including drivers’ shortage, skills shortage, ageing workforce and challenges related to safety, security and connectivity. In the road haulage industry, cargo equipment is a frequent target of organised criminal groups. Consequently, cargo thefts and illegal truck boarding cause significant financial and reputational losses for supply chain operators. <br/><br/>
Currently, there is a lack of facilities that allow for safe parking of trucks and that also provide a minimum level of services for the social well-being of drivers. By establishing a denser network of secure truck parking areas (SSTPAs) with a clear definition of security levels, it is possible to address these issues all together. Drivers, transport companies, freight forwarders, shippers and insurers, as well as society as a whole, will benefit from a sufficient supply of these facilities by protecting drivers, cargo and transport equipment. In addition, road safety through well-rested drivers is another area where such facilities can have a significant positive impact. In this context, the European Commission conducted a study on safe parking areas for trucks, which was completed in December 2018. The main findings of this study demonstrated a need for (van Weenen et al., 2019):

-	A common standard and rating system for safety and service;
-	Audit responsibilities and guidelines to ensure reliability;
-	Comprehensive maps showing where SSTPAs are needed;
-	Recommendations for the basis of a common application program interface (API) for the exchange of dynamic data between SSTPAs and information platforms;
-	Manuals to support the creation of business plans for the establishment of SSTPAs.

A gap analysis to compare secure truck parking demand and supply at the European level and in more detail along the core network identified a number of issues, including (van Weenen et al., 2019):

-	The total night parking demand is 400,000 truck parking spaces per night;
-	Only 300,000 truck parking spaces are available, which means a deficit of about 100,000 additional spaces;
-	The deficit in certified secure parking spaces is much greater, as only 7,000 spaces are available in a few countries. In some countries and on certain corridors, motorists cannot rely on the availability of certified secure parking spaces;
-	The current supply of non-secure parking spaces is more evenly distributed across the network. However, these spaces are not certified and do not offer guaranteed services to drivers.

Truck Parking Europe helps transport companies ensure the safety of drivers and freight. Inspection reports provide information on the safety features in truck parking areas. These audit reports help safety and planning managers to approve safe parking areas across Europe (in line with the standardised safety framework created by the EU). They provide a detailed overview of the facilities that each truck parking area offers (Truck Parking Europe, no date).

### Relevant initiatives in Austria {-}

- [truckparkingeurope-austria](https://www.truckparkingeurope.com/secure-truck-parking-overview/austria/) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table70, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Increase safety and security of drivers|**+**|Health & Wellbeing (*3*)|van Weenen et al., 2019|
|Systemic|Reduced loses ralated to cargo thefts|**+**|Sustainable economic development (*8,11*)|van Weenen et al., 2019|
|Systemic|Additional focus on the alternative fuel infrastructure|**+**|Innovation & Infrastructure (*9*)|Gnap & Kubikova, 2020|')
```

### Technology and societal readiness level {-}

```{r table71, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|7-9|')
```

### Open questions {-}
1. How to attract the interest of motorway companies, both private and public, to facilitate data provision, upgrading of parking sites for security and reservation, and encourage the infrastructure construction to capture dynamic data?
2. What software infrastructure is needed to allow for in-vehicle up-to-date information about the occupancy and reservation of truck parking areas?


### Further links {-}

- [truckparkingeurope.com-1](https://www.truckparkingeurope.com/secure-truck-parking-overview/)
- [truckparkingeurope.com-2](https://www.truckparkingeurope.com/how-do-i-reserve-a-secure-truck-parking-spot/)
- [ec.europa.eu-1](https://ec.europa.eu/transport/themes/its/road/action_plan/intelligent-truck-parking_en)
- [ec.europa.eu-2](https://ec.europa.eu/transport/sites/transport/files/modes/road/parking/doc/2010_04_28_setpos_project_handbook.pdf) 
- [ec.europa.eu-3](https://ec.europa.eu/transport/sites/transport/files/2019-study-on-safe-and-secure-parking-places-for-trucks.pdf) 
- [tis-gdv.de](https://www.tis-gdv.de/tis_e/bedingungen/parken/lpp-htm/)
- [unece.org](https://unece.org/DAM/trans/doc/2008/sc1/ECE-TRANS-SC1-103-pres01e.pdf)


### References {-}
-	Carrese, S., Mantovani, S., & Nigro, M. (2014). A security plan procedure for Heavy Goods Vehicles parking areas: An application to the Lazio Region (Italy). Transportation Research Part E: Logistics and Transportation Review, 65(1), 35–49. https://doi.org/10.1016/j.tre.2013.12.011
-	Gnap, J., & Kubíková, S. S. (2020). Possible Effects of Lacking Parking Areas for Road Freight Transport on Logistics and Transport Safety. Transportation Research Procedia, 44(2019), 53–60. https://doi.org/10.1016/j.trpro.2020.02.009
-	Lozia, Z., & Kulma, J. (2014). Simulation method of comparative evaluation of the agility of a passenger car when moving ‘forwards’ and ‘backwards’. Archiwum Motoryzacji, 64, 49–64, 149.
-	Nevland, E. A., Gingerich, K., & Park, P. Y. (2020). A data-driven systematic approach for identifying and classifying long-haul truck parking locations. Transport Policy, 96, 48–59. https://doi.org/10.1016/j.tranpol.2020.04.003
-	Sochor, J., & Mbiydzenyuy, G. (2013). Assessing the benefits of intelligent truck parking. International Journal of Intelligent Transportation Systems Research, 11(2), 43-53.
-	Truck Parking Europe. (n.d.). Audit reports on secure truck parking locations - Truck Parking Europe. Retrieved 3 March 2021, from https://www.truckparkingeurope.com/2020/03/11/audit-reports-on-secure-truck-parking-locations/
-	van Weenen, R. de L., Newton, S., Menist, M., Maas, F., Penasse, D., Nielsen, M., Halatsis, A., Männistö, T., Stamos, I., & Ruschin, P. P. (2019). Study on Safe and Secure Parking Places for Trucks (Issue February).

## Delivery space booking {#space_book} 

### Synonyms {-}
*on-street loading zones (OLZ), Freight Trip Generation (FTG), loading/unloading zones (L/U) *

### Definition {-}
In urban areas, goods deliveries for commercial and residential customers generate a huge flow of vehicles ranging from small vans (for express deliveries) to trucks (for deliveries to larger shops). Delivery services are increasing due to a number of factors such as just-in-time management, the development of e-commerce and the emergence of new customer behaviours  for example home delivery, driving and delivery lockers (Patier et al., 2014).
Urban distribution usually requires delivery vehicles to stop temporarily at the side of the road to allow the driver to complete the last part of the delivery on foot. The stops take place in designated areas called loading/unloading (L/U) zones, each with a certain number of available parking spaces. Distributors often face bottlenecks in the L/U areas due to a variety of factors, such as peaks in delivery demands in terms of time or space, systemic problems in the allocation of L/U areas (e.g. scarcity) or poor placement of spaces (Mor et al., 2020).
All this complicates the movement of goods in city centres, with negative effects on vehicle traffic and public transport, for example by causing congestion. In European cities (especially in France), a significant proportion of double-parked delivery trucks (delivery trucks parked on the street parallel to parked cars) is observed. Such behaviour leads to congestion, pollution and conflicts between road users. In this context, special areas called "delivery zones" have been created to improve the work of delivery drivers and reduce congestion.  Nevertheless, delivery drivers continue to double park as the delivery areas are regularly occupied by unauthorised vehicles (single cars) (Patier et al., 2014).


### Key stakeholders {-}

- **Affected**: Delivery truck drivers in urban areas, Car drivers, Road and/or pavement users in cities, parcel/goods receivers
- **Responsible**: National governments, Local governments, Transport authorities, Delivery companies

### Current state of art in research {-}
Several works address the problem of sizing and locating L/U areas in an urban environment. Muñuzuri et al., 2017 presents several alternative approaches to estimating the need for loading zones. Pinto et al., 2019 propose to determine the number and location of L/U areas using a "coverage principle" based on the longest distance a delivery person is willing to walk and define the number of spaces in each area based on demand. The impact of parking availability on the cost and operation of commercial vehicles is explored in (Dezi et al., 2010).<br/><br/>
Mor et al., 2020 conclude that double parking is unlikely to disappear from urban areas unless more dedicated L/U spaces are made available at peak times. In turn, however, the lack of available parking spaces for commercial vehicles is not only caused by undersized infrastructure, but often also by the misuse of reserved spaces by private vehicles (Aiura & Taniguchi, 2005; Pinto et al., 2019).<br/><br/>
In Nourinejad et al., (2014), a parking choice simulation for the study of truck parking policies is presented that captures various dimensions of parking activity such as walking distance, congestion impacts and parking search times. Two scenarios based on the Toronto area are presented to validate the model.
The issue of pricing L/U spaces has also been considered in the literature. An auction-based approach for allocating time slots of a single area with multiple parking spaces is considered to optimise the performance of a management system while ensuring fair allocation of L/U zones (Nourinejad et al., 2014). The time preferences and service duration of booking requests are considered in finding the best possible allocation.<br/><br/>
Next, a sizing problem considering L/U range management is also looked at. A two-phase algorithm is proposed for the problem (Letnik et al., 2018). In the first phase, the location of the L/U areas is optimised based on the location of the goods receivers. In the second phase, the algorithm optimises deliveries from outside the city to the L/U areas. L/U areas are assigned to vehicles and if none are available, vehicles are queued. The benefits of a booking system for L/U areas are assessed for the case of Winchester High Street (McLeod & Cherrett, 2011). It is assumed that each vehicle arrives from one of eight possible entry points and stops in an L/U area, with the possibility of transfer to another area if the desired one is not available. A control system to deal with operational problems such as vehicles arriving too early or too late is also presented.<br/><br/>
The management of L/U areas through a booking system would require the involvement of all stakeholders (customers, traders, municipality, traffic police) and their acceptance that it is in everyone's interest that the system works well as it allows for less inconvenience both in terms of delivery delays and traffic flow. In the initial phase of implementation, the traffic police would need to oversee the new system and educate any trader coming with an unregistered vehicle or without registration about the new rules for using the L/U areas. Information for stakeholders could be provided via the Internet (and possibly in some meetings) to share some statistics on the adoption of the system, the occupancy of delivery positions across the hours of the day and days of the week, and to review the problems each category has in using the system and explore corrective measures (Mor et al., 2020).

### Current state of art in practice {-}
Efforts are being made in various cities around the world to introduce a management system for L/U areas. For example, in Barcelona's *DUM programme1*, traders are asked to check in and out of any L/U operation using an SMS or an app in a smartphone, but in September 2020 no reservation could be made (Mor et al., 2020).
Companies such as *Smart Parking Systems*, *Coord*, *ParkUnload* and *Cleverciti* offer solutions for delivery area booking. However, many different terms are still used, such as "curb management", "smart loading zones" or "digital loading zones". <br/><br/>
In 2010, Treviso was awarded the title of "First European Smart City". Through Smart Parking Applications, the city has benefited from a significant increase in revenue from paid municipal parking, greater economic availability and a reduction in traffic and pollution within the historic centre. Parking management within Treviso Municipality was already in an excellent situation in 2009, monitored and supervised by an extremely attentive mobility department.
The real problem of the municipality was, therefore, not the economic aspect, but mainly the excessive traffic and the scarce availability of parking spaces in some areas of the city. For this reason, it was considered appropriate to use smart parking technology to influence urban mobility in connection with parking. Since 2010, Treviso has, therefore, embarked on this technological path together with Intercomp, installing Smart Parking Systems® in the historic centre with the long-term goal of improving mobility, air quality and service for citizens (Smart Parking Systems®, no date).<br/><br/>
In September 2020, the City of Omaha (US) launched its Smart Zone pilot project with *Coord* in five Smart Zones. Overall, the rollout has gone well, although the city and fleets still need to learn how to disseminate information. Fleets have embraced the concept well in particular that they do not pay to use the Smart Zones. It is expected that once more data is collected and the concept becomes more established the city will introduce the charge (Brown, 2020).

### Relevant initiatives in Austria {-}
In Vienna, more than 2600 loading zones were recorded and located via geodata as part of a project. This data is now available via app *Ladezonen in Wien*. It is not yet possible to make a booking or check the current occupancy rate. The project *i-Ladezone* was first mentioned in 2010 from FFG Austria. In 2012, the project *i-Ladezone - Intelligent Loading Zone Routing and Management* was mentioned again, but only to detect and record the illegal occupancy of loading zones (Stocker, 2012). In 2016, the project *Urban Loading* was submitted to the VCÖ Mobility Award. From a technical point of view, reservation/booking, information to the road user and control/enforcement were considered (VCÖ, 2016). No further information about this project was found.

- [ots.at](https://www.ots.at/presseaussendung/OTS_20150123_OTS0044/simple-stressfreie-ladezonensuche-wk-wien-praesentiert-neue-app) 
- [wko.at-1](https://www.wko.at/service/verkehr-betriebsstandort/Ladezonen-Nutzung.html) 
- [wko.at-2](https://www.wko.at/service/w/verkehr-betriebsstandort/ladezone-wien-app.html) 
- [oeamtc.at](https://www2.ffg.at/verkehr/projekte.php?id=805&lang=de&browse=programm) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table82, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Congestion, pollution and conflicts between road users can be reduced|**+**|Health & Wellbeing (*3*)|Patier et al., 2014; Alho et al., 2018; Mor et al., 2020|
|Systemic|Digitalised solution for occupancy and availability status|**+**|Innovation & Infrastructure (*9*)|Smart Parking Systems, no date|
|Systemic|L/U booking systems requires engagement of all stakeholders|          **+**|Partnership & collaborations (*17*)|Mor et al., 2020|')
```

### Technology and societal readiness level {-}

```{r table83, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|5-8|')
```

### Open questions {-}
1. How to improve the information dissemination and enable the engagement of all stakeholders to develop an efficient L/U booking system?
2. What are the solutions for enforcement of legislation beyond traffic officers?

### Further links {-}

- [areaverda.cat](https://www.areaverda.cat/index.php/es)
- [smartparkingsystems.com](https://smartparkingsystems.com/en/loading-parking-bay-management/)
- [smartloadingzone.com](https://www.smartloadingzone.com/en_us/)
- [coord.com](https://www.coord.com/blog/comparing-loading-zones)

### References {-}
-	Aiura, N., & Taniguchi, E. (2005). PLANNING ON-STREET LOADING-UNLOADING SPACES CONSIDERING THE BEHAVIOUR OF PICKUP-DELIVERY VEHICLES. Journal of the Eastern Asia Society for Transportation Studies, 6, 2963–2974. https://doi.org/10.11175/easts.6.2963
-	Alho, A. R., de Abreu e Silva, J., de Sousa, J. P., & Blanco, E. (2018). Improving mobility by optimizing the number, location and usage of loading/unloading bays for urban freight vehicles. Transportation Research Part D: Transport and Environment, 61, 3–18. https://doi.org/10.1016/j.trd.2017.05.014
-	Brown, C. (2020, October 13). Should Delivery Fleets Pay to Park in Loading Zones? - Smart Cities - Fleet Forward. https://www.fleetforward.com/10127896/should-delivery-fleets-pay-to-park-in-loading-zones
-	Dezi, G., Dondi, G., & Sangiorgi, C. (2010). Urban freight transport in Bologna: Planning commercial vehicle loading/unloading zones. Procedia - Social and Behavioral Sciences, 2(3), 5990–6001. https://doi.org/10.1016/j.sbspro.2010.04.013
-	Letnik, T., Farina, A., Mencinger, M., Lupi, M., & Božičnik, S. (2018). Dynamic management of loading bays for energy efficient urban freight deliveries. Energy, 159, 916–928. https://doi.org/10.1016/j.energy.2018.06.125
-	McLeod, F., & Cherrett, T. (2011). Loading bay booking and control for urban freight. International Journal of Logistics Research and Applications, 14(6), 385–397. https://doi.org/10.1080/13675567.2011.641525
-	Mor, A., Speranza, M. G., & Viegas, J. M. (2020). Efficient loading and unloading operations via a booking system. Transportation Research Part E: Logistics and Transportation Review, 141, 102040. https://doi.org/10.1016/j.tre.2020.102040
-	Muñuzuri, J., Cuberos, M., Abaurrea, F., & Escudero, A. (2017). Improving the design of urban loading zone systems. Journal of Transport Geography, 59, 1–13. https://doi.org/10.1016/j.jtrangeo.2017.01.004
-	Nourinejad, M., Wenneman, A., Habib, K. N., & Roorda, M. J. (2014). Truck parking in urban areas: Application of choice modelling within traffic microsimulation. Transportation Research Part A: Policy and Practice, 64, 54–64. https://doi.org/10.1016/j.tra.2014.03.006
-	Patier, D., David, B., Chalon, R., & Deslandres, V. (2014). A New Concept for Urban Logistics Delivery Area Booking. Procedia - Social and Behavioral Sciences, 125, 99–110. https://doi.org/10.1016/j.sbspro.2014.01.1459
-	Pinto, R., Lagorio, A., & Golini, R. (2019). The location and sizing of urban freight loading/unloading lay-by areas. International Journal of Production Research, 57(1), 83–99. https://doi.org/10.1080/00207543.2018.1461269
-	Smart Parking Systems®. (n.d.). The economic benefits of Smart Parking Systems® > Smart Parking Systems - Intercomp Innovation. Retrieved 21 March 2021, from https://smartparkingsystems.com/en/the-economic-benefits-of-smart-parking-systems/
-	Stocker, G. (2012, November). Snizek+Partner Verkehrsplanungs GmbH - iLadezone - Intelligentes Ladezonenrouting und Management. https://www.snizek.at/go/de/projects/simulationen/94-iladezonen-intelligentes-ladezonenrouting-und-management
-	VCÖ. (2016). Urban Loading - VCÖ Vorbildhafte Mobilitätsprojekte. https://mobilitaetsprojekte.vcoe.at/urban-loading



## Delivery drones {#delivery_drone}

### Synonyms {-}
*urban air mobility (UAM), vertical take-off and landing (VTOL), unmanned aerial vehicles (UAVs), Unmanned Traffic Management (UTM)/U-Space*

### Definition {-}
In the recent year, online shopping and the demand for same-day deliveries has grown exponentially, with more and more people preferring online shopping instead of buying items stationary (Di Puglia Pugliese et al., 2020). At the same time, the last mile is the most expensive process in distribution logistics, contributing to 13% - 73% of the total distribution cost. As a result there is an increased pressure for more efficient solutions in the field.  The answer to this problem could be drones, or unmanned aerial vehicles (UAVs). Drones combine three key principles of technological modernity - computing, autonomy and limitless mobility. Capabilities that until now could only be used by the military are becoming accessible to most of the population. Potential use for drones ranges from surveillance and reconnaissance missions to novel forms of logistics and personal transport. The commercial use of drones is associated with enormous economic opportunities. Even though drones are already common as surveillance/sensing devices in security services, geodesy or agriculture, their use as a means of transport is still at the beginning. Delivery drones are currently able to lift weights of up to 2-3 kg and carry out flight assignments in an urban space (Kellermann et al., 2020).<br/>
What is more, an accelerated progress is expected in the near future, where the European Commission estimates the economic impact of the wider use of drones at €10 billion annually until 2035 and foresees the creation of more than 100,000 direct jobs. Taking into account indirect macroeconomic effects in drone-related industries, the Commission even projects 250,000 to 400,000 additional jobs (SESAR, 2016).

### Key stakeholders {-}

- **Affected**: Mobile citizen, delivery companies and truck drivers, customers of online shops
- **Responsible**: National governments, city government,  private companies, online shops

### Current state of art in research {-}
Current research shows that drones reduce noise, congestion in urban areas and CO~2~ emissions. Multiple studies showed that the use of drones in logistics processes can reduce the lead time (defined as the latency between the beginning and the completion of the process), time span and transport costs of the delivery process (Di Puglia Pugliese et al., 2020; Wang et al., 2017; Di Puglia Pugliese & Guerriero, 2017). However, it is also important that drones have lower capacity and working time (low flight endurance) than classic vehicles. 
Besides all the benefits that drones can bring, there are also some potential drawbacks associated with an increased use of drones. These include safety and security issues, aerial collisions, crashes, malfunctioning of software and hardware components, misuse of drones for criminal purposes, dangers to wildlife and privacy violations, among others (Kellermann et al., 2020).
Moreover, the media analysis about drones for parcel and passenger transportation found the following thematic priorities in relevant literature (see table below): 

```{r table29, echo=FALSE, results='asis', fig.cap="Adapted from Kellermann, Biehle and Fischer, 2020"}
cat('| Topics| Percentage|Number of studies|
|:------------:|:-----------:|:------------:|
|General Surveys    |   18.9%  |  21   |
|Logistics (general)  |  18.0%   |  20   |
|Attitude and Acceptance Research    | 13.5%    |  15   |
|Law and Regulations    | 11.7%    |   13  |
|Ethics and Technology Assessment    | 10.8%    |  12   |
|Sustainability Assessment    | 8.1%    |  9   |
|Urban and Transportation Planning    |  7.2%   |  8   |
|Political agenda/strategies    |  6.3%   |   7  |
|Passenger Transportation   |  2.7%   |  3   |
|Humanitarian Logistics    |2.7%     |   3  |') 
```

The research proposes two approaches in terms of the incorporation of drones into delivery process, either as a sole delivery option or as a support system for trucks. Calculation results show the disadvantage of the exclusive use of drones in the delivery process. Overall, the vehicle routing that uses a combination of trucks and drones is the transport system offers the best trade-off between transport costs, CO~2~  emissions and congestion (Di Puglia Pugliese et al., 2020). For example, with 2 drones per truck, the delivery completion time can be reduced by 75% in the best case scenario (Wang et al., 2017). <br/> 

Current research also examines the potential of drones in food delivery services due to COVID-19 (Kim et al., 2021). A study on choice-model-based analysis of consumer preference for a drone delivery service concludes the following (Kim, 2020):

-	Drone delivery service competes with truck delivery service for clothing and powder formula, and with motorbike delivery service for urgent documents. 
-	Potential customers' preferences for drone delivery service depend on the price and type of goods. 
-	Consumers are concerned about the reliability of the drone delivery service for expensive items. 
-	In all models, it is observed that young people are more likely to choose drone delivery service than old people.

### Current state of art in practice {-}
Several large companies, such as *Amazon* (Amazon.com, 2021), *DHL* (Di Puglia Pugliese et al., 2020), *UBER* , *Wing* (Bonnington, 2012), *UPS* (UPS, 2021) and some more have started using delivery drones for parcel delivery. However, governments and experts fear the misuse of drones. In particular, after their concerns materialised with the incident at London Airport (Keane, 2018). To protect against misuse, work is being carried out internationally on a digital remote ID so that the drone operators can always be assigned to the drone (Boyle, 2020). <br/> <br/>
In the USA, the Federal Aviation Administration (FAA) has allowed drone flights as long as the drone is within an operator's line of sight. However, companies are working on developing drones to fly without this requirement (Fox Rubin, 2019). Further, UPS is already carrying out drone deliveries between WakeMed's main hospital campus in Raleigh, North Carolina, and its neighbouring laboratory centre to send laboratory samples back and forth. The company is also planning flights between hospitals to provide easier access to life-saving drugs like anti-venom (Fox Rubin, 2019). The fact that the first drone experiments have focused on business-to-business deliveries in the healthcare sector is not a coincidence. Firstly, funding is available in this sector, and secondly, the speed of the drones can bring considerable benefits in this field (Marshall, 2020).<br/> <br/>
Moreover, Amazon announced back in 2013 that they would start delivering with drones within years and in 2019, alongside the unveiling of their new drone, they announced that they would start drone delivery "within months" (Kolodny & Feiner, 2019). Nonetheless, it is currently claimed that even with all the test flights and announcements, it will be several years before a drone can deliver (Marshall, 2020).

### Relevant initiatives in Austria {-}
One of the European Amazon drone development centres is localised in Austria (Marshall, 2020). Moreover, a new drone law came into force in the EU on 31 December 2020 to standardise the rules. It requires a CE marking on the drone and a drone licence for drones weighing 250g or more.

- [www.oesterreich.gv.at-1](https://www.oesterreich.gv.at/themen/dokumente_und_recht/Drohnen.html) 
- [www.oesterreich.gv.at-2](https://www.oesterreich.gv.at/themen/dokumente_und_recht/Drohnen/Drohnenfuehrerschein.html) 
- [www.oesterreich.gv.at-3](https://www.oesterreich.gv.at/themen/dokumente_und_recht/Drohnen/EU-Regelungen-f%C3%BCr-Drohnen-im-%C3%9Cberblick.html) 
- [oeamtc.at](https://www.oeamtc.at/thema/drohnen/drohnen-info-app-26853120) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table30, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Faster provision of time-sensitive goods (eg. pharmaceuticals)|**+**|Health & Wellbeing (*3*)|Fox Rubin, 2019|
|Individual|Drone deliveries offered at no extra cost|          **+**|Sustainable economic development (*8,11*)|Amazon.com, 2021|
|Individual|New delivery options available|          **+**|Innovation & Infrastructure (*9*)|Kim, 2020; Amazon.com, 2021|
|Systemic|Reduced time of completion of logistics processes|          **+**|Sustainable economic development (*8,11*)|Wang et al., 2017|
|Systemic|Standardised drones legislation across the EU  |          **+**|Partnership & collaborations (*17*)|EASA, 2021|')
```

### Technology and societal readiness level {-}

```{r table31, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-8|3-5|')
```

### Open questions {-}
1. Will private companies work together and share their technologies?
2. What other areas of application will there be?
3. What will be the thematic priorities for development in the coming years?
4. Will freight drones automatically communicate with other drones such as surveillance, geodesy and agricultural drones?

### Further links {-}

-	[easa.europa.eu](https://www.easa.europa.eu/sites/default/files/dfu/Easy%20Access%20Rules%20for%20Unmanned%20Aircraft%20Systems%20(Regulations%20(EU)%202019-947%20and%20(EU)%202019-945).pdf) 

### References {-}
- Amazon.com: Prime Air. (n.d.). Retrieved 27 January 2021, from https://www.amazon.com/Amazon-Prime-Air/b?ie=UTF8&node=8037720011
-	Bonnington, C. (2012, March 23). Tacocopter: The Coolest Airborne Taco Delivery System That’s Completely Fake | WIRED. https://www.wired.com/2012/03/qa-with-tacocopter/
-	Boyle, A. (2020, May 6). Amazon and T-Mobile will take part in FAA’s Remote ID program for drones - GeekWire. https://www.geekwire.com/2020/amazon-t-mobile-will-take-part-faas-remote-id-drone-monitoring-program/
-	Di Puglia Pugliese, L., & Guerriero, F. (2017). Last-Mile Deliveries by Using Drones and Classical Vehicles. In A. Sforza & C. Sterle (Eds.), Optimization and Decision Science: Methodologies and Applications (pp. 557–565). Springer International Publishing.
-	Di Puglia Pugliese, L., Guerriero, F., & Macrina, G. (2020). Using drones for parcels delivery process. Procedia Manufacturing, 42, 488–497. https://doi.org/10.1016/j.promfg.2020.02.043
- EASA (2021). Easy Access Rules for Unmanned Aircraft Systems (Regulations (EU) 2019/947 and (EU) 2019/945) Available at: https://www.easa.europa.eu/sites/default/files/dfu/Easy%20Access%20Rules%20for%20Unmanned%20Aircraft%20Systems%20(Regulations%20(EU)%202019-947%20and%20(EU)%202019-945).pdf. (Accessed: 08 February 2021)
-	Fox Rubin, B. (2019). UPS drone delivers medicine from CVS straight to customers’ homes. https://www.cnet.com/news/ups-drone-deliver-medicine-from-cvs-straight-to-customers-homes/
-	Keane, S. (2018). London’s Gatwick Airport closes again after drones force 33-hour shutdown. https://www.cnet.com/news/airport-suspends-all-flights-after-drones-get-too-close-to-airfield/
-	Kellermann, R., Biehle, T., & Fischer, L. (2020). Drones for parcel and passenger transportation: A literature review. Transportation Research Interdisciplinary Perspectives, 4, 100088. https://doi.org/10.1016/j.trip.2019.100088
-	Kim, J. J., Kim, I., & Hwang, J. (2021). A change of perceived innovativeness for contactless food delivery services using drones after the outbreak of COVID-19. International Journal of Hospitality Management, 93(June 2020), 102758. https://doi.org/10.1016/j.ijhm.2020.102758
-	Kim, S. H. (2020). Choice model based analysis of consumer preference for drone delivery service. Journal of Air Transport Management, 84(September 2019), 101785. https://doi.org/10.1016/j.jairtraman.2020.101785
-	Kolodny, L., & Feiner, L. (2019, June 5). Amazon debuts its new delivery drone. https://www.cnbc.com/2019/06/05/amazon-debuts-its-new-delivery-drone.html
-	Marshall, A. (2020, May 8). No, Amazon Won’t Deliver You a Burrito by Drone Anytime Soon | WIRED. https://www.wired.com/story/amazon-wont-deliver-burrito-drone-soon/
-	SESAR. (2016). European Drones Outlook Study. In Single European Sky ATM Research (Issue November). https://doi.org/10.2829/219851
-	UPS. (n.d.). UPS Drone Delivery Service | UPS - United States. Retrieved 4 February 2021, from https://www.ups.com/us/en/services/shipping-services/flight-forward-drones.page
-	Wang, X., Poikonen, S., & Golden, B. (2017). The vehicle routing problem with drones: several worst-case results. Optimization Letters, 11(4), 679–697. https://doi.org/10.1007/s11590-016-1035-3

## Rail telematics for freight services {#rail_telematics_freight}

### Definition {-}
With increasing digitalisation and automation in rail operations, many applications are being developed that will make it possible to significantly increase the attractiveness and efficiency of rail freight transport. Telematics applications are a functional subsystem of the railway system. This subsystem comprises two elements:

-	Applications for [passenger services](#telematics_passenger)
-	Applications for freight transport

In the area of freight transport, this includes the following (European Commission, 2021):

-	information systems (real-time monitoring of goods and trains), 
-	marshalling and allocation systems, 
-	reservation, 
-	payment and settlement systems, 
-	management of connections with other modes of transport and production of electronic accompanying documents 

The basis for a smart and connected freight wagon is, in addition to telematics for geolocation and communication (satellite navigation, inertial and mobile radio systems), the use of sensors for condition recording and intelligent validation of driving and operating conditions (Ußler et al., 2019). In road transport, about 85% of trucks are equipped with telematics systems that monitor the condition of the vehicle, track its position and provide a direct communication with the driver via mobile phone. Fleet management receives real-time information and has redundant ways to communicate with the transport in case of problems. Rail freight traditionally operates its transports with infrastructure-based information sources. Therefore, wagons are usually not equipped with wagon-based monitoring and communication devices such as telematics systems and sensors. The result is a constant loss of time-critical goods to road transport (Behrends et al., 2016).  <br/>

### Key stakeholders {-}

- **Affected**: Mobile citizen, delivery companies and truck drivers, customers of online shops
- **Responsible**: National governments, private companies, public rail provider

### Current state of art in research {-}
[Single wagonload transport](https://blog.railcargo.com/de/artikel/einzelwagenverkehr) (SWL) that refers to a situation where individual wagons or groups of wagons from different consignors are bundled together to form one train, is an important component in the European rail transport system and in the logistics of various industries such as steel, chemicals and automotive. However, changing framework conditions and increasingly demanding market requirements have led to dramatic losses of market share and in some countries even to the complete discontinuation of the SWL business. As this business segment is also seen as an important part of the European co-modal transport system in the future, significant improvements are necessary. <br/> 
With on-board communication technology, freight operators can improve wagon dispatching and rescheduling processes in case of disruptions. Based on reliable online telematics data, dispatchers will be able to inform their customers about changes in the transport plan earlier, which will increase reliability and satisfaction of stakeholders. 
Cost-efficient and intelligent telematics-based information services allow wagons to be tracked in real time and automatically present wagon mileage information. The telematics data service generates the necessary information required for reliable quality recording. This additional information optimises the current six-year life cycle of freight wagons. 


### Current state of art in practice {-}
A group of freight rail industry stakeholders in the US announced the formation of Rail Pulse, a joint venture in late 2020 to create a technology platform which will facilitate and accelerate the adoption of GPS and other telematics technologies in the North American railcar fleet. Rail Pulse partners are focused on promoting the adoption of telematics on two fronts:

-	Rail safety, with the early phase of the platform focused on handbrake and impact data, which they believe could provide important safety data points for railways, car owners and shippers, coupled with a forward-looking approach to telematic features such as on-board bearing temperature and wheel impact detection sensors;
-	Improve the competitive position of rail compared to other modes by improving visibility of the status, location and condition of individual wagons, with telematic capabilities including data collection to support real-time track-level visibility of whether doors or hatches are open, whether the wagon is loaded or partially loaded and other key performance indicators.<br/>

Development of the Rail Pulse platform was expected to begin by the end of 2020, with a full service platform available to the North American railcar industry by the end of 2022 (Berman, 2020).
An investment in wagon telematics results in lower costs and higher turnover. One of the critical indications for a large implementation of telematics seems to be the fact that all three parties - shippers, railways and wagon keepers - will share the benefit of "lower costs". But this distribution of benefits may also lead to a "wait and see" behaviour of many railway stakeholders compared to their competitors on the road (Behrends et al., 2016). <br/>

Thus far, a significant cost reduction was achieved with new GSM and GPS modules. Interfaces for supplementary sensors such as impact detection, digital/analogue inputs/outputs, etc. were integrated. Furthermore, based on the analysis of user requirements, the development of a reliable load-sensing technology for freight wagons was started. This requirement emerged from the fact that today most freight wagons in railway operation do not use their full load capacity, as there is no cost-effective way to measure the load, especially during the filling process, e.g. in the area of bulk goods. If the wagons were to be overloaded and then moved by a train, all wheelsets would have to be replaced during costly inspection in a workshop. As a result, the freight wagon is not filled to its maximum payload, which leads to lower capacity and higher costs. To optimise the dispatching processes, the wagons transmit their loading status so that the dispatchers can re-dispatch the wagon in a short time after unloading, resulting in shorter downtimes and higher efficiency of the wagons (Behrends et al., 2016).

### Relevant initiatives in Austria {-}

- [www.a1.digital](https://www.a1.digital/en-de/about-a1-digital/Press-Releases/OBB-freight-trains-become-smart-with-A1-Digital/) 
- [blog.railcargo.com](https://blog.railcargo.com/en/artikel/smart-cargo-erstmontage) 
- [statistik.at](https://www.statistik.at/web_de/statistiken/energie_umwelt_innovation_mobilitaet/verkehr/schiene/gueterverkehr/index.html) 
- [railcargo.com](https://www.railcargo.com/de/unternehmen/international/oesterreich) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table58, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Rail safety improved|**+**|Health & Wellbeing (*3*)|Berman, 2020|
|Systemic|Potential for lower costs for all involved parties|**+**|Sustainable economic development (*8,11*)|Behrends et al., 2016|
|Systemic|Digitalisation implemented|**+**|Innovation & Infrastructure  (*9*)|Behrends et al., 2016|
|Systemic|Potential for *wait and see* behaviour |**-**|Partnership & collaborations (*17*)|Behrends et al., 2016|')
```

### Technology and societal readiness level {-}

```{r table59, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|8-9|')
```

### Open questions {-}
1. Whiat are the barriers that prevent the implementation of digital solutions?
2. What should be changed to enable an effective use of digital solutions?


### Further links {-}
- [wascosa.ch](https://www.wascosa.ch/infoletter/2016/wascosa-infoletter-25-en.pdf)
- [bearingpoint.com](https://www.bearingpoint.com/files/Digitalization_rail_infrastructure_management_PR.pdf?download=0&itemId=659400)
-	[ec.europa.eu](https://ec.europa.eu/transport/modes/rail/useful-links_de)
- [blog.railcargo.com](https://blog.railcargo.com/en/artikel/einzelwagenverkehr)
- [taf-jsg.info](http://taf-jsg.info/wp-content/uploads/2019/05/RU-IM-Telematics-governance-ToR_V1.3_final-14-05-2019.pdf)

### References {-}
-	Behrends, V., Haunschild, M., & Galonske, N. (2016). Smart Telematics Enabling Efficient Rail Transport - Development of the ViWaS Research and Development Project. Transportation Research Procedia, 14, 4430–4439. https://doi.org/10.1016/j.trpro.2016.05.365
-	Berman, J. (2020, October 22). Freight rail offering aims to boost GPS and telematics adoption across North American railcar fleet - Logistics Management. https://www.logisticsmgmt.com/article/freight_rail_offering_aims_to_boost_gps_and_telematics_adoption_across_nort
-	European Commission. (2021). Telematic applications | Mobility and Transport. https://ec.europa.eu/transport/modes/rail/interoperability/interoperability/telematic_applications_de
-	Ußler, H., Michler, O., & Löffler, G. (2019). Validation of multiple sensor systems based on a telematics platform for intelligent freight wagons. Transportation Research Procedia, 37(September 2018), 187–194. https://doi.org/10.1016/j.trpro.2018.12.182

## Electric vehicle delivery fleets {#electric_delivery_fleets}

## Multimodal transport management systems {#mtms}

## Freight hubs {#freight_hubs}