# Freight and commercial transport {#freight}

## Tracking and tracing of goods {#dangerous_goods}

### Synonyms {-}
*Radio Frequency Identification (RFID)* 

### Definition {-}
Current technology developments offer promising opportunities to address diverse challenges faced by manufacturers in today's dynamic business environment. To effectively manage ever-changing customer demands, managers are looking to technology solutions, such as tracking and tracing technologies, to improve planning, control and performance (Kache & Seuring, 2017). This improved level of service leads to a higher customer satisfaction and is crucial in such a competitive market as it "*increases customer trust, strengthens brand integrity and increases customer loyalty*" (Costa et al., 2013). <br/>

Integration into existing systems and whether to label at item or package level remains a challenge. Perhaps this is one reason why adoption has been strongest in some segments of the retail industry, where finished goods are mostly sold by piece. Therefore, there is a need to shed more light on the challenges of adopting and using tracking and tracing technologies in other industrial settings if the potential benefits are to be fully realised (Høyer et al., 2019). <br/>

**Existing technologies** <br/>
Various tracking and tracing systems have different capabilities and functions, some span the entire chain, from farmer to retailer, while other solutions are limited to a specific area. The level of detail of the information these technologies capture varies. The following technologies were found during the literature review and were considered popular for tracking (Høyer et al., 2019).

**Barcodes** <br/>
Barcodes and barcode scanners are an established technology for identifying products. Barcode technology can often be seen as a simpler form of tracking, yet it is often preferred in industry as it is easier to implement and is a more cost-effective solution that still captures data at the required level of detail and accuracy (Høyer et al., 2019).

**RFID** <br/>
Radio Frequency Identification (RFID) technology consists of two components: an antenna and the chip that contains the electronic product code. The information can be tracked in real-time throughout the chain. With the emergence of RFID technology, research has proven that it has improved inventory handling and warehouse management (Lao et al., 2012). The adoption of RFID technology also leads to a reduction in human errors originally caused by manual data entry (RFID tags transmit the data "by themselves", which largely eliminates any human error in the inventory process (cwi-logistics, 2020)). The significant benefits will be noted most by the distributor and the retailer (Høyer et al., 2019).
Thus, unlike barcodes, RFID tags do not require a scanner to be in the same line of sight. This eliminates the need to manually scan each box because items can be scanned and catalogued even if they are hidden behind other goods (cwi-logistics, 2020). <br/>

However, RFID tags are problematic from a technological point of view, mainly because there is no (global) industry-wide standardisation. Also, because RFID tags and their systems operate on radio frequency, they can easily become jammed or disturbed, reducing their usability. RFID systems can also experience signal problems, such as collisions when signals from two or more readers overlap, or interference is caused by metal, water or other magnetic fields in the environment. An RFID system is time-consuming and labour-intensive to set up. Companies need to test different hardware and tag systems to determine the best fit. This can take months. In addition to the cost of the RFID system itself, such as RFID tags and scanners, an increase in time and labour also means an increase in cost. These types of disadvantages are often avoided by using barcodes, which is why they remain a popular choice for many businesses for data collection and inventory control (Jänisch, 2019). <br/>

**Smart packaging systems and TTIs** <br/>
These packaging systems equipped with time-temperature indicators (TTIs) can sense the environment and detect, record, track and communicate based on stimuli. These functions help in decision making regarding shelf life and quality, they are able to warn in case of deviations, and they support the flow of materials and information (Høyer et al., 2019).

### Key stakeholders {-}

- **Affected**: Online shop consumers, Freight forwarders, Shippers
- **Responsible**: National Governments, City government, Private Companies

### Current state of art in research {-}
The challenges with tracking and tracing technology encountered during the literature review can be categorised as strategic challenges, technical challenges and convenience challenges (Vermesan & Friess, 2014).

**Strategic challenges**<br/>

-	Implementation costs: High implementation costs remain one of the biggest barriers to the adoption of certain tracking and tracing technologies, such as RFID.
-	Low awareness of the benefits and lack of incentives: It is claimed that there is a lack of incentives to adopt the technologies and risk when applying new technologies to old processes. These may be incompatible and lead to increased costs and inefficiencies. 
-	Information sharing: i.e. increased supply chain visibility and integration, can have a significant positive impact on the entire supply chain by improving planning, production and delivery performance (Zhou & Benton, 2007). With the advent of Big Data, it is important to ensure that masses of data are made interpretable and available to all partners in a timely manner (Morgan et al., 2018).
-	Coordination, collaboration and trust: A supply chain can potentially consist of multiple partners and there is always a risk of divergent and misaligned interests that can affect the quality of information shared. The strategic value of some information can inhibit the free exchange of information (Aramyan et al., 2007). There is a tendency to associate the act of sharing information with the loss of power and dependency (Soosay & Hyland, 2015). 
-	Ingrained business practices: If not all key stakeholders are on board, this can lead to the technology not being implemented at all due to, for instance, high one-off investments.

**Technical challenges**<br/>

-	Colliding signal: One risk with these technologies is that signals may overlap.
-	Environmental interference: Environmental factors and materials with high water content affect the performance of tracing technologies (Kumari et al., 2015). These characteristics are critical in food supply chains where food is often affected by high water content, exposed to extreme temperatures, and have dielectric properties that can interfere with signals. 
-	Suboptimal reading:On one hand, errors can occur and on the other hand, the packaging of the product and not the product itself is registered.  
-	Data acquisition: The industry is lagging in comparison to the research that has been done to digitise supply chains. The reality is that manual and paper-based operations are still common practice, the data collected is unstructured and the masses of data generated are difficult to handle as current collection systems are limited and unable to handle large volumes of data (Høyer et al., 2019).

**Convenience challenges**<br/>

-	Waste and recycling: As for recycling the RFID tags, there is no easy way other than to remove them from each box before it is thrown away and then melt down the metal antenna if one is used.
-	Lack of professional skills: It is possibly due to inadequate or poor training of employees in the use of tracking and tracing technology, limits the potential in the supply chain (Ruiz-Garcia & Lunadei, 2011).
-	Privacy and security: Privacy issues prevent companies from taking advantage of the opportunities offered by tracking technologies. The consequences are fake barcodes, hacking, industrial espionage, unwanted customer tracking, virus attacks and malicious intent.
-	Regulations and standards: As transparency increases, so does the need for regulations and traceability standards, with several standards currently co-existing (Kumari et al., 2015). The lack of standards leads to system incompatibilities that make it difficult to share information.
-	Unification and standardisation of data: The process of data collection and transmission varies between supply chain partners. This disparity makes collaboration difficult and leads to greater incompatibilities (Høyer et al., 2019).

### Current state of art in practice {-}
There is no doubt that tracking and tracing technology leads to more real and representative data, improving and facilitating decision-making. Some manufacturers have abandoned RFID pilot projects due to high implementation costs and lack of benefits, but many manufacturers are still interested in exploring other solutions that support operational decision-making. The physical implementation of tracking and tracing technology has not generally posed any major difficulties, instead it appears to be aspects relating to organisational issues and security (Høyer et al., 2019). <br/>

The Digital Container Shipping Association (DCSA), a neutral, non-profit group formed by major shipping lines to digitise and standardise container shipping, has announced that "*the majority of its member shipping lines have adopted the DCSA Track & Trace (T&T) standards and are currently or will soon be providing their customers with access to the standards-based API*".
The DCSA believes this is a ground breaking development that "*provides shippers with a streamlined way to obtain real-time data on the whereabouts of their containers*". The widespread adoption of the DCSA standard will move the industry forward in terms of real-time visibility and responsiveness, leading to greater reliability and a better customer experience. <br/>
Executives from the world's leading shipping companies MSC, ONE, CMA CGM, Yang Ming and Evergreen issued statements supporting the new T&T standards. All executives emphasised the need for standards to support digital initiatives working across multiple shipping companies and highlighted the importance of standardisation to their digitalisation initiatives. <br/>

"*We are very pleased to see the increasing adoption among shipping companies*" said Thomas Bagge, the CEO of DCSA. "*But the digital transformation of the container shipping industry and the resulting improvements in efficiency and customer experience simply cannot happen without even wider adoption of digital standards. DCSA's focus for 2021 is to drive adoption by all stakeholders. Without acceptance, the industry will not benefit from the digital foundation that is being created.*" (Avery, 2021). <br/>

What is now already perceived as live tracking by the customer is tracking based on the position of the delivery driver. The vehicle is equipped with GPS, which means that the driver's position can be tracked but not the parcels themselves. DPD has been using this system for several years. When a shipment is moved, the barcode on the parcel, which is linked to the delivery address, is scanned. As a result, the current shipment data is loaded into a database.Amazon also offers real-time tracking with its own delivery service, Amazon Logistics. Amazon's map tracking service uses its own intelligent route planning for this - but technical details are not disclosed. However, not all parcels are sent via Amazon Logistics, especially since the transport service is not yet available in all regions. The tracking in the map view is, therefore, also not available everywhere. <br/>

In contrast to Germany, DHL Austria does not yet offer live tracking. Here, parcels can still be tracked conventionally by means of physical scans of a barcode on a parcel-by-parcel basis at certain process points (Iosa, 2021).

### Relevant initiatives in Austria {-}
DPD and Amazon already offer live tracking in Austria. Real-time tracking is not yet offered by Austrian Post but is in the planning stage and is expected to become a focus in 2021. According to a company spokeswoman, the aim is not so much to track the delivery route online and live, but rather to continuously update the route and the time of delivery. In this way, the delivery window can be limited to the smallest possible range. At the moment, the tracking of consignments is still done with scan events, which update the status of the parcel with every scan.
In addition, historical data is currently used to calculate the time window for delivery. Data from the past is compared to determine how long it took to reach an address during a particular tour. In this way, the time of delivery can be narrowed down to three hours (Iosa, 2021).

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table97, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Improved planning, control and performance|**+**|Sustainable economic development  (*8,11*)|Kache & Seuring, 2017|
|Systemic|Development in tagging technologies|**+**|Innovation & Infrastructure  (*9*)|cwi-logistics, 2020|
|Systemic|Effort and initatives for global standardisation |**+**|Partnership & collaborations (*17*)|Avery, 2021|')
```

### Technology and societal readiness level {-}

```{r table98, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|7-9|')
```

### Open questions {-}
1. Is the potential of RFID tags to reduce the carbon footprint higher than the waste they can produce?

### Further links {-}

- [cargomon.com-tracking](http://cargomon.com/index.php/language/en/cargoobserver-track/)
- [cargomon.com-gps monitoring](http://cargomon.com/index.php/language/en/industry-sectors-for-montoring-gps/)
- [imaginenext.ingrammicro.com](https://imaginenext.ingrammicro.com/iot/rfid-in-logistics)
- [rand.org](https://www.rand.org/pubs/technical_reports/TR1283.html)


### References {-}
-	Aramyan, L. H., Oude Lansink, A. G. J. M., van der Vorst, J. G. A. J., & van Kooten, O. (2007). Performance measurement in agri‐food supply chains: a case study. Supply Chain Management: An International Journal, 12(4), 304–315. https://doi.org/10.1108/13598540710759826
-	Avery, P. (2021, April 9). WorldCargo News - News - Majority of DCSA members adopt Track and Trace standards. https://www.worldcargonews.com/news/majority-of-dcsa-members-adopt-track-and-trace-standards-65996
-	Costa, C., Antonucci, F., Pallottino, F., Aguzzi, J., Sarriá, D., & Menesatti, P. (2013). A Review on Agri-food Supply Chain Traceability by Means of RFID Technology. Food and Bioprocess Technology, 6(2), 353–366. https://doi.org/10.1007/s11947-012-0958-7
-	cwi-logistics. (2020, January 13). How RFID Is Taking Warehousing to the Next Level | CWI Logistics. https://cwi-logistics.com/news/how-rfid-is-taking-warehousing-to-the-next-level/
-	Høyer, M. R., Oluyisola, O. E., Strandhagen, J. O., & Semini, M. G. (2019). Exploring the challenges with applying tracking and tracing technology in the dairy industry. IFAC-PapersOnLine, 52(13), 1727–1732. https://doi.org/10.1016/j.ifacol.2019.11.450
-	Iosa, A. (2021, January 9). Wie funktioniert Live-Tracking bei Paketen? https://futurezone.at/digital-life/wie-funktioniert-live-tracking-bei-paketen/401110917
-	Jänisch, R. (2019, February 17). Was sind RFID-Tags? Funktionsweise, Einsatz und Nachteile. https://ioxlab.de/de/iot-tech-blog/was-sind-rfid-tags/#Einsatzgebiete_fuer_IOX_RFID
-	Kache, F., & Seuring, S. (2017). Challenges and opportunities of digital information at the intersection of Big Data Analytics and supply chain management. International Journal of Operations & Production Management, 37(1), 10–36. https://doi.org/10.1108/IJOPM-02-2015-0078
-	Kumari, L., Narsaiah, K., Grewal, M. K., & Anurag, R. K. (2015). Application of RFID in agri-food sector. Trends in Food Science and Technology, 43(2), 144–161. https://doi.org/10.1016/j.tifs.2015.02.005
-	Lao, S. I., Choy, K. L., Ho, G. T. S., & Yam, R. C. M. (2012). An RFRS that combines RFID and CBR technologies. Industrial Management & Data Systems, 112(3), 385–404. https://doi.org/10.1108/02635571211210040
-	Morgan, T. R., Richey Jr, R. G., & Ellinger, A. E. (2018). Supplier transparency: scale development and validation. The International Journal of Logistics Management, 29(3), 959–984. https://doi.org/10.1108/IJLM-01-2017-0018
-	Ruiz-Garcia, L., & Lunadei, L. (2011). The role of RFID in agriculture: Applications, limitations and challenges. Computers and Electronics in Agriculture, 79(1), 42–50. https://doi.org/10.1016/j.compag.2011.08.010
-	Soosay, C. A., & Hyland, P. (2015). A decade of supply chain collaboration and directions for future research. Supply Chain Management: An International Journal, 20(6), 613–630. https://doi.org/10.1108/SCM-06-2015-0217
-	Vermesan, O., & Friess, P. (2014). Internet of Things - From Research and Innovation to Market Deployment. In River Publishers. https://doi.org/10.1007/978-3-642-24788-0_19
-	Zhou, H., & Benton, W. C. (2007). Supply chain practice and information sharing. Journal of Operations Management, 25(6), 1348–1365. https://doi.org/10.1016/j.jom.2007.01.009

## Intermodal Freight {#intermodal_freight}

### Synonyms {-}
*unaccompanied combined transport (UCT), intermodal transport unit (ITE)* 

### Definition {-}
Intermodal freight refers to the transport of goods in an intermodal transport unit (ITE) with at least two different modes of transport (e.g. road, rail, inland waterway). ITEs are containers or swap bodies, accompanied or unaccompanied road freight vehicles and trailers of road freight vehicles. The essential feature of intermodal transport is that, when the transport unit is changed to another mode of transport, there is no transhipment of goods, i.e. the entire ITE is always loaded onto another mode of transport (Rudlof et al., 2018).<br/>

In terms of the advantages the intermodal freight is crucial in reducing the emissions where transport accounts for almost a quarter of the European Union's green house gases (GHG) emissions, of which road transport accounts for 74%. In particular, freight transport by road is an important source of emissions, as heavy-duty vehicles are responsible for 6% of the European Union's total emissions. Therefore, reducing carbon emissions associated with freight transport by road is a priority to meet current policy emission reduction targets, such as reducing GHG emissions by 80% by 2050 under the terms of the Paris Agreement. Other possible solutions to this problem include technological improvements, such as the development of more efficient and near-zero emission engines, and management improvements, such as reducing and optimising transport distances and shifting traffic from high-emission to low-emission vehicles, for example from trucks to trains or ships (European Commission, no date). Nonetheless, intermodal freight is frequently associated with slower speed, lack of reliability and increased potential for damages of the good.


### Key stakeholders {-}

- **Affected**: Truck drivers, Freight companies, Rail companies, Freight terminals, Highway users
- **Responsible**: National Governments, International authorities

### Current state of art in research {-}
Research focuses on the ecological perspective of intermodal transport (Heinold & Meisel, 2020) as well as automated intermodal freight transport systems that aim to reduce GHG emissions, minimise infrastructure and logistics costs and reduce traffic congestion (Shin et al., 2018). Moreover, some papers, for example Saeedi et al. (2019) look at the technical efficiency of the intermodal transport. 

### Current state of art in practice {-}
Over the last two decades, EU transport policy has promoted intermodal freight transport (by rail or inland waterways). In 2011, the European Commission set a target of shifting 30% of freight transported further than 300 km by road to other modes such as rail or waterways by 2030, and more than 50% by 2050. Despite significant investments (about €28 billion in funding for rail projects between 2007 and 2013) and the priority shift of freight from road to intermodal freight transport (IFT), EU policy for intermodal transport has not achieved significant improvements (European Court of Auditors, 2016). The performance of an IFT service is attributed to two main factors: the performance of the different parts of the chain and the cooperation and harmony between these parts (Saeedi et al., 2019). <br/>

Since this special type of freight transport has become increasingly important in recent years from an economic and political point of view, Eurostat has set up the Task Force on Intermodal Transport Statistics on this topic. The task force, in which Statistics Austria (STAT) also participated, was to evaluate how meaningful and high-quality data on intermodal transport could be compiled at European level for road and rail transport as well as inland and maritime navigation without imposing additional burdens on the member states of the European Union. It was investigated whether in individual countries, in addition to the data on intermodal transport already to be collected according to EU regulations,  information is available that could provide a better picture of transport with ITE. This assessment showed that information on intermodal transport is available in many different forms across countries. Possible indicators can, therefore, only be compiled in a limited form at EU level, as their characteristics differ with regard to the individual modes of transport and also show inconsistencies due to methodological reasons. For example, there are differences in the weights to be collected or the type and size of the ITEs. Furthermore, information on the source and destination regions of transported containers is not available, nor is information on transport chains. This is because the transports on individual modes are considered independent of each other. In the context of the Task Force it also turned out that, compared to other Member States, most information on intermodal transport at national level is available in Germany. The Federal Statistical Office in Germany (Destatis) collects data on the transport of goods in containers, swap bodies, road freight vehicles and trailers of road freight vehicles on roads, railways, rivers and seas. For other countries, one of the basic assumptions is that only a small amount of ITE is transported by road freight vehicles, mainly over short distances (Rudlof et al., 2018). <br/>

Some foreign developments in the field of automated freight transport system technologies have entered pilot operations after finalising their concept designs. In other cases, further developments have been suspended due to economic challenges such as high initial investment. In the case of FSS, an unmanned automated freight transport system was developed in the US and considered the most advanced of its kind in terms of its development progress. The system is currently in test operation after completing the construction of a test bed with a 100 m linear track. Overall, in the advanced countries, it is considered that the developments have reached a point where they will be subjected to the validation processes for commercialisation after the concept phase is completed. Upon completion of further testing, product launches are expected within the next 3 to 4 years (Shin etal., 2018).


### Relevant initiatives in Austria {-}
In Austria, there is a special support programme for unaccompanied combined transport (UCT) by rail, which is intended to contribute to a shift of freight transport to this mode of transport. The Rail Infrastructure Service Company ([SCHIG](https://www.schig.com/)) is responsible for handling these subsidies on behalf of the BMVIT.<br/>

A survey of the terminals as well as the analysis of the SCHIG data sets on the UCT funding programme have shown that no supplementary data on intermodal transport is currently available in Austria. Without new and specially designed surveys, it is currently not possible to compile comprehensive statistics on intermodal transport across all modes of transport, which would above all also allow statements on transport chains. One way of obtaining data could be, for example, to oblige the terminals where intermodal transport takes place to keep records and make them available to the statistics. However, the necessary legal basis for this would first have to be created.<br/>

Intermodal transport in Austria in 2016 was particularly important in the area of rail freight transport with a share of 22.3%. In contrast, only 2.6% of the transport volume on the road was carried in intermodal transport units, and intermodal transport was of no significance for the inland waterway mode of transport, as only empty containers were transported on inland vessels (Rudlof et al., 2018).

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table76, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Potential for reduction in emissions|**+**|Environmental sustainability  (*7,12-13,15*)|European Commission, no date|
|Systemic|Lack of standarisation in information on intermodal transport across EU|**~**|Partnership & collaborations (*17*)|Rudlof et al., 2018|')
```

### Technology and societal readiness level {-}

```{r table77, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|8-9|')
```

### Open questions {-}
1. How can coordination between multi-level decision makers be improved to make intermodal performance more efficient?
2. Given that drayage operations are a large proportion of total cost in intermodal freight, how they can be performed more efficiently?
3. How the standarisation in information collection, storage and sharing procedures can be achieved across the EU conutries?


### Further links {-}

- [eca.europa.eu](https://www.eca.europa.eu/en/Pages/DocItem.aspx?did=36398)
- [transportgeography.org](https://transportgeography.org/contents/chapter5/intermodal-transportation-containerization/)

### References {-}
-	European Commission. (n.d.). Transport emissions | Climate Action. Retrieved 11 March 2021, from https://ec.europa.eu/clima/policies/transport_en
-	European Court of Auditors. (2016). Rail freight transport in the EU: still not on the right track (Issue 08). https://doi.org/10.2865/53961
-	Heinold, A., & Meisel, F. (2020). Emission limits and emission allocation schemes in intermodal freight transportation. Transportation Research Part E: Logistics and Transportation Review, 141, 101963. https://doi.org/10.1016/j.tre.2020.101963
-	Rudlof, M., Karner, T., & Fleck, S. (2018). Intermodaler Verkehr in Österreich. 87–95.
-	Saeedi, H., Behdani, B., Wiegmans, B., & Zuidwijk, R. (2019). Assessing the technical efficiency of intermodal freight transport chains using a modified network DEA approach. Transportation Research Part E: Logistics and Transportation Review, 126, 66–86. https://doi.org/10.1016/j.tre.2019.04.003
-	Shin, S., Roh, H. S., & Hur, S. H. (2018). Technical Trends Related to Intermodal Automated Freight Transport Systems (AFTS) *. Asian Journal of Shipping and Logistics, 34(2), 161–169. https://doi.org/10.1016/j.ajsl.2018.06.013


## Urban Deliveries {#urban_delivery}

### Synonyms {-}
*city logistics, micro-hubs, "Grätzl"-hubs, City Distribution Centers, micro-depots*

### Definition {-}
Efficient and reliable goods delivery and collection services are essential for the functioning of businesses in the city centre area and the consumption of products and services by residents, visitors and employees (Aljohani & Thompson, 2020). However, limited space, increased traffic and, at the same time, continuously growing demand are increasingly challenging, especially in the urban areas. In addition, freight traffic wears out roads much more than passenger traffic due to the high axle loads and is, therefore, critical for road maintenance planning (Forschungs-Informations-System für Mobilität und Verkehr (FIS), 2003). The concept of city logistics, therefore, seeks to relieve the strain on urban infrastructure while improving the quality of supply in the city (Aljohani & Thompson, 2020). <br/>

There are two primary areas that could provide practical and relevant solutions to address the challenges and efficiencies of last mile freight activity in the city center: *(1)* Freight Demand Management (FDM) measures and *(2)* improving parking and charging infrastructure. Recipients typically have more market power than freight forwarders or shippers, and can significantly influence the timing, size and frequency of deliveries, which typically occur during business hours and high traffic periods. One action would be to push deliveries outside business hours to change the delivery activities of carriers and shippers (Holguín-Veras and Sánchez-Díaz, 2016). Local authorities can oblige large recipients and building managers in the city centre to coordinate and consolidate their deliveries from multiple freight companies. Freight behaviour studies should be conducted to assess the interest of these receivers and determine what policy levers might influence their participation (Aljohani & Thompson, 2020).<br/>

Research shows that both receiver pricing and incentives could play an important role in reducing freight transport, with receiver pricing likely to face significantly more political resistance than incentives. Policymakers must make pragmatic choices, balancing economic optimality with political feasibility. Careful selection of receiver fees and incentives will likely result in a situation where significant sustainability improvements can occur at minimal political cost. A further development of the Economic Order Quantity (EOQ) model shows that to maximize profits, receivers must allocate floor space appropriately between revenue-generating activities and storage areas. Pressure to maximize floor space for revenue-generating activities leads to smaller warehouse areas and greater reliance on small order sizes and frequent deliveries. Optimal space allocation and ordering leads to order sizes and cycle times that are 35% smaller than in the classic EOQ model. This would result in 50% more delivery trips. Applying an optimal receiver charge - to incentivize receivers to internalize the externalities generated - suggests that a reduction in Freight Trip Generation (FTG) of about 10% could be achieved (Holguín-Veras & Sánchez-Díaz, 2016). The results suggest that resorting to frequent small shipments is a rational decision in a profit-maximizing environment. A negative effect that should be noted is that mode choice strategies that require the adoption of large shipment sizes, such as the use of rail, may be rejected by recipients because they assume larger storage areas. In contrast, mode choice strategies that might work with smaller shipments, such as the use of electric bicycles, may be accepted by recipients as long as they do not unduly increase logistics costs.<br/>

Aljohani & Thompson (2020) argue that road loading space regulations and allocations need to be updated to take into account the challenges faced by hauliers in the city centre. Local authorities might argue that there are enough loading spaces in the city centre. However, they are not managed and distributed efficiently as illegal users park in these loading spaces. An on-street loading zone (OLZ) policy should take into account the different parking requirements for the different classes of freight vehicles and the sub-industries of the recipients. In addition, it may be necessary to convert some OLZs in busy locations to serve as temporary staging areas for freight operators, especially in dense retail and commercial areas. Local authorities should also take advantage of advances in Internet-of-Things (IoT) technologies, number plate recognition and smart occupancy signs to enable booking of OLZs and display booking details (see [Delivery space booking](#space_book)). These adaptive displays could act as virtual on-street charging zones that only become active when a booking is made.<br/>

Holguín-Veras & Sánchez-Díaz (2016) argue in their paper that their findings once again illustrate the potential of freight demand management. The authors believe that harnessing the power of recipients as agents of change offers a unique opportunity to improve the economic efficiency of urban freight transport systems, reduce the negative externalities caused by freight transport and improve both quality of life and environmental justice.

### Key stakeholders {-}

- **Affected**: Courier Express Parcel Services, package recipients, Freight Forwarder
- **Responsible**: Courier Express Parcel Services, Logistics companies, City administration, Funding provider, District authorities, Chamber of Commerce, Private companies

### Current state of art in research {-}
Holguín-Veras & Sánchez-Díaz (2016) used survey data in a case study in New York City (NYC) to estimate the potential traffic reduction that could be achieved through a receiver-led consolidation (RLC) programme. The authors estimated a model to show that the main factors that determine this decision are industry sector, factory space (a measure of size) and the number of deliveries received. The industries that tend to participate are retail, accommodation and wholesale. Small businesses tend to be less interested in RLC. The model and freight trip generation (FTG) analyses show that RLC could result in reductions in the range of 3.0% to 8.8% of total delivery traffic in the NYC metropolitan area and between 3.5% and 11.2% in Manhattan. Applying a behavioural-based microsimulation to Manhattan showed that the total savings to carriers could range from $376 906 to $1 186 128 per day, due to 4740 to 15062 hours of operation saved. In terms of vehicle miles traveled (VMT), expected savings range from 33445 to 104255 VMT per day. These estimates equate to a savings of approximately $65 in operating costs, approximately 50 min in travel and processing time saved, and the avoidance of approximately 6 miles of travel per consolidated delivery. In addition to the savings to carriers participating in RLC, a reduction in truck traffic is expected to generate economic benefits. The magnitude of these savings suggests that public agencies should consider RLC programmes as a part of their sustainability efforts. The study also found that policies in the form of prizes and/or incentives are needed to encourage recipients to act. Otherwise, corporate inertia and the natural tendency to change only when circumstances force it will lead recipients to maintain the status quo.<br/>

Beyond, the idea of using existing light rail and suburban rail networks for the distribution and delivery of goods in cities has also been and is increasingly discussed. Several case studies have also been conducted in European cities, such as Amsterdam, Dresden, Paris and Zurich. Marino et. al (2013) concluded that urban freight transport by rail is a feasible concept as it brings several benefits, essentially a reduction in congestion, emissions and traffic in cities. However, the concept has not really caught on so far. An example in Amsterdam showed that financing can be a challenge, as there was no clear consensus among the parties involved.

### Current state of art in practice {-}
To develop the [Logistics 2030+ Action Plan](https://www.logistik2030.at/?page_id=2769) (Popp, et al., 2019), the Province of Lower Austria, the City of Vienna and the Chambers of Commerce of Lower Austria and Vienna launched the "*Sustainable Logistics 2030+ Lower Austria-Vienna*" project. Over a period of three years, more than 300 stakeholders were involved in a multi-stage process to achieve the following project goals: *(1)* Resolution of conflicts of use in flowing and stationary freight traffic, *(2)* CO2 savings, *(3)* reduced traffic volume without loss of performance and quality, *(4)* development of logistics and traffic concepts and *(4)* initiation and implementation of pilot projects. Eight thematic clusters were worked on for this purpose: 

-	Planning and securing logistics areas with foresight
-	Driving forward freight consolidation with the help of new business models
-	Developing and implementing efficient parcel delivery solutions
-	Supporting sustainable logistics concepts at companies and large-scale projects
-	Create incentives for accelerated fleet conversions
-	Use digital information and services to increase efficiency and optimization
-	Establish framework conditions for sustainable development
-	Actively communicate logistics services and costs

Finally, the action plan includes 35 measures with 133 actions (ARGE L2030+, n.d. a). <br/>

In the topic cluster "*Solutions for parcel delivery*" , mentioned above, the measures are: *(1)* Avoid non-delivery: Fruitless first delivery attempts should be avoided by providing alternative addresses, *(2)* use P&R facilities and public transport stations as white label B2C nodes: Create better and cheaper offers for commuters, *(3)* "*Grätzelboxen*" and box/logistics rooms in new buildings and existing properties: Grätzelboxen are microhubs with additional functions and integration of box locations in properties, and due to the Covid 19 crisis also added *(4)* the increase of contactless delivery (Faast & Gregori, 2020).<br/>

Furthermore, loading zones represent a central component of the public logistics infrastructure, which is essential for supplying the city with goods and services. In order to counteract the lack of this infrastructure, which was identified as a major problem for urban logistics by representatives of the Austrian business community during the preparation of the L2030+ work program, the "*Loading Zone Calculator*" project was launched. The loading zone calculator is to consist of a GIS tool and an empirical characteristic value determination for estimating the freight traffic volume for a loading zone management. Based on the preliminary study completed by the TU Wien in April 2020 on the topic of "*Estimation of freight traffic volume for a loading zone management*", the steps of the procedure are to be implemented in a digital GIS tool including a characteristic value matrix for Vienna. Within the scope of the preliminary study, it became apparent that with regard to characteristic values of freight traffic in the literature, there is often a missing or insufficient reference to the use/industry. For this reason, a standardized procedure for the empirical determination of characteristic values including an exemplary survey for one sector shall be implemented in this project.<br/>

In 2017, a City Hub, specialising in e-commerce, was implemented with LogPOINT Logistics Services GmbH and the support of the Vienna Chamber of Commerce on one of the last efficient inner-city logistics sites in Vienna. The comprehensive service portfolio in the field of e-fulfilment offers customers a wide range of options, from full logistics services to the handling of parcels. Parcels are bundled at the location and from then on delivered CO~2~-free to both B2B and B2C addresses by means of cargo bikes and e-vehicles. On the one hand, the pilot project enables the flood of parcels addressed to the inner-city districts to be brought in, handled and delivered directly on the basis of swap body transshipment. On the other hand, the company's team, which is experienced in e-fulfilment, offers a full service product, whereby parcel volumes are produced directly at the location (ARGE L2030+, n.d. b).<br/>

Parcel boxes have evolved from their singular use as a parcel delivery box. In addition to parking deliveries, they can also be used by retailers for "*Click & Collect*", by local service providers, by the administration for documents and ID cards, as C2C lockers, for X2B return shipments or by X2C for a wide variety of applications. X2C can be different users such as "Willhaben" buyers, sales representatives, delicatessen and wine merchants, craftspeople, Airbnb platform users, and so on (Faast & Gregori, 2020). Currently, each box operator has its own IT platform, which they connect with the IT systems of the CEP service providers and those of the suppliers/retailers. Likewise, each box operator has its own receiver interface, which usually enables access to the boxes via a corresponding app. In principle, all the necessary functionalities are mapped, but there are currently no standards and interoperability between the systems is not possible. In the medium term, therefore, appropriate standards or a superordinate platform must be developed which - analogous to mobile communications technology - enable problem-free switching/exchange between the various platforms (Faast & Gregori, 2020).

### Relevant initiatives in Austria {-}

- [logistik2030.at-1](https://www.logistik2030.at/?page_id=268) 
- [infothek.bmk.gv.at](https://infothek.bmk.gv.at/gruene-stadtlogistik-post-testet-city-hubs-in-wien/) 
- [logistik2030.at-2](https://www.logistik2030.at/?page_id=63) 
- [remihub.at](https://www.remihub.at) 
- [logpoint.at](https://logpoint.at/ueber-uns/gruene-logistikwelt-und-standorte/)
- [wu.ac.at](https://www.wu.ac.at/scm/projekte/)
- [post.at](https://www.post.at/p/c/vorzimmer-zustellung) 


### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table113, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Cost savings through efficiency|**+**|Sustainable economic development (*8,11*)|Holguin-Veras & Sanchez-Diaz, 2016|
|Systemic|Reduced negative externalities|**+**|Environmental sustainability (*7,12-13,15*)|Holguin-Veras & Sanchez-Diaz, 2016
|Systemic|Economic benefits due to reduction in truck traffic|**+**|Sustainable economic development (*8,11*)|Holguin-Veras & Sanchez-Diaz, 2016|
|Systemic|Less shipment by rail; more shipment by cargo bikes and electric cars|**~**|Innovation & Infrastructure (*9*)|Holguin-Veras & Sanchez-Diaz, 2016|')
```

### Technology and societal readiness level {-}

```{r table114, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|4-7|4-6|')
```

### Open questions {-}
1. How can the different IT platforms of the individual providers be combined?
2. How can we ensure that facilities such as City Logisic Hubs, parcel boxes, etc. are open to all? 

### Further links {-}

- [railfreight.com](https://www.railfreight.com/railfreight/2017/08/03/light-rail-network-used-for-freight-transport/?gdpr=accept )

### References {-}
-	Aljohani, K., & Thompson, R. G. (2020). Last mile delivery activities in the city centre–Insights into current practices and characteristics of delivery trips. Transportation Research Procedia, 46, 261-268.
-	ARGE L2030+. (n.d. a). Aktionsplan – Logistik 2030+. Retrieved April 29, 2021, from https://www.logistik2030.at/?page_id=63
-	ARGE L2030+. (n.d. b). Pilotprojekte – Logistik 2030+. Central LogPOINT – DER Logistik HUB im Herzen von Wien. Retrieved April 28, 2021, from https://www.logistik2030.at/?page_id=268
-	Faast, A., & Gregori, G. (2020). NACHHALTIGE LOGISTIK 2030+ NIEDERÖSTERREICH WIEN Pilotprojekt Evaluierung von großteils betreiberunabhängigen Paketboxensystemen in Niederösterreich und Wien.
-	Forschungs-Informations-System für Mobilität und Verkehr (FIS). (2003, April 8). Infrastrukturschäden durch den Straßengüterverkehr. https://www.forschungsinformationssystem.de/servlet/is/39816/
-	Holguín-Veras, J., & Sánchez-Díaz, I. (2016). Freight demand management and the potential of receiver-led consolidation programs. Transportation Research Part A: Policy and Practice, 84, 109-130.
-	Marinov, M., Giubilei, F., Gerhardt, M., Özkan, T., Stergiou, E., Papadopol, M., & Cabecinha, L. (2013). Urban freight movement by rail. Journal of Transport Literature, 7(3), 87-116.
-	Popp, C., Winkler, A., Hahn, E., & Faast, A. (2019). Nachhaltige Logistik 2030+ Niederösterreich-Wien Aktionsplan.
-	van Leijen, M. (2017, August 3). Light rail network used for freight transport | RailFreight.com. https://www.railfreight.com/railfreight/2017/08/03/light-rail-network-used-for-freight-transport/?gdpr=accept&gdpr=deny



## Intelligent truck parking {#intelligent_truck_park}

### Synonyms {-}
*ITP, Information on Truck Parking*

### Definition {-}
Nowadays, there is a significant increase in transport volume, especially in the international road freight transport. In particular, with the enlargement of the EU member states, the interdependence of the member states has increased as well as the ratio of the exchange of goods and thus the increase in need for transport services (Gnap & Kubíková, 2020). Consequently, the higher number of vehicles and associated traffic lead to difficulties in meeting deadlines for loading and unloading goods as well as the obligation to take mandatory safety breaks and rest periods. It also amplifies the issue with lack of parking spaces along the route for the Heavy Goods Vehicle (HGV) which has a negative impact on logistics and creates assumptions for increasing logistics costs (Lozia & Kulma, 2014). <br/><br/>
Therefore, the European Union is striving to support the development of the Trans-European Transport Network (TEN-T) through structural funds. The TEN-T network is intended to connect the member states from east to west and from north to south. Unfortunately, the rate of road construction in some countries is far below the rate of growth of transport services. The abolishment of in-vehicle resting and other upcoming changes in the so-called "road package" will increase the need for more parking spaces and associated equipment for drivers, especially on motorways. This fact will have a significant impact on the pressure on vehicle parking spaces when the maximum prescribed driving time is observed and there is a need to stop a vehicle and take a safety break or a daily/weekly rest period (Carrese et al., 2014). On top of this, there is a clear need for charging and refuelling stations for alternative fuels at rest areas to ensure smooth transit of vehicles using alternative fuels. These should be in line with the implementation of [Directive 2014/94/EU](https://eur-lex.europa.eu/legal-content/en/TXT/?uri=CELEX%3A32014L0094) of the European Parliament. <br/><br/>
Therefore, to attempt addressing these issues, an Intelligent Truck Parking (ITP) emerged which is an Intelligent Transportation Systems service concerned with the management of information related to HGV parking operations. ITP collects data about parking facilities (services and availability), processes the data, e.g. occupancy level, and communicates the information to drivers (in real time) in order to help them easily access such a facility. ITP enables efficient parking management and potentially improve the needs of transport stakeholders through the provision of information such as available parking spaces and additional services (Sochor & Mbiydzenyuy, 2013). <br/><br/>
Good examples of the efficient use of the information system can be found in Germany or Austria. These systems, as part of the intelligent transport systems, offer services that make transport easier for drivers, e.g. searching for free parking spaces at rest areas located along motorways and motorways (Gnap & Kubíková, 2020).

### Key stakeholders {-}

- **Affected**: Drivers, Transport companies, Freight forwarders, Shippers and Insurers
- **Responsible**: National Governments, Fright Companies, International authorities (eg. EU)

### Current state of art in research {-}
In the literature mostly, regional plans for specific areas in a certain country can be found. For example Carrese et al., (2014) for the Lazio Region in Italy and Gnap & Kubíková, (2020) for the Slovak Republic. However, safety aspects are also highlighted, such as a classification scheme of parking spaces for trucks in long-distance traffic. Again, the data is limited to a specific region (Peel Region in Ontario, Canada) but efforts have been made to develop a generally applicable scheme (Nevland et al., 2020).

### Current state of art in practice {-}
A common European park strategy was developed in 2014 by the European organisation SETPOS within the [SETPOS]( https://ec.europa.eu/transport/sites/transport/files/modes/road/parking/doc/2010_04_28_setpos_project_handbook.pdf) project. The project had four objectives:

-	The assessment and validation of the requirements of stakeholders, i.e. drivers, dispatchers, hauliers, rest area operators, insurers, authorities and shippers;
-	The formulation of a common set of standards for secured parking;
-	The establishment of a number of pilot parking areas in cross-border regions to validate and demonstrate the standard set;
-	To set up an information, guidance and reservation platform for all types of truck parking.

The [LABEL]( https://ec.europa.eu/transport/sites/transport/files/modes/road/parking/doc/handbook_for_labelling.pdf ) project, which follows the SETPOS, aims to increase safety and quality standards for truck parking in Europe by (Carrese et al., 2014):

-	Introducing a European standard certification scheme for truck parking areas;
-	developing an online database for users to provide financial benefits to certified sites.

Every day, billions of euros worth of goods are transported on the trans-European road network, the backbone of the EU economy. Despite the sector's successful performance with respect to volume, it faces a significant number of issues, including drivers’ shortage, skills shortage, ageing workforce and challenges related to safety, security and connectivity. In the road haulage industry, cargo equipment is a frequent target of organised criminal groups. Consequently, cargo thefts and illegal truck boarding cause significant financial and reputational losses for supply chain operators. <br/><br/>
Currently, there is a lack of facilities that allow for safe parking of trucks and that also provide a minimum level of services for the social well-being of drivers. By establishing a denser network of secure truck parking areas (SSTPAs) with a clear definition of security levels, it is possible to address these issues all together. Drivers, transport companies, freight forwarders, shippers and insurers, as well as society as a whole, will benefit from a sufficient supply of these facilities by protecting drivers, cargo and transport equipment. In addition, road safety through well-rested drivers is another area where such facilities can have a significant positive impact. In this context, the European Commission conducted a study on safe parking areas for trucks, which was completed in December 2018. The main findings of this study demonstrated a need for (van Weenen et al., 2019):

-	A common standard and rating system for safety and service;
-	Audit responsibilities and guidelines to ensure reliability;
-	Comprehensive maps showing where SSTPAs are needed;
-	Recommendations for the basis of a common application program interface (API) for the exchange of dynamic data between SSTPAs and information platforms;
-	Manuals to support the creation of business plans for the establishment of SSTPAs.

A gap analysis to compare secure truck parking demand and supply at the European level and in more detail along the core network identified a number of issues, including (van Weenen et al., 2019):

-	The total night parking demand is 400,000 truck parking spaces per night;
-	Only 300,000 truck parking spaces are available, which means a deficit of about 100,000 additional spaces;
-	The deficit in certified secure parking spaces is much greater, as only 7,000 spaces are available in a few countries. In some countries and on certain corridors, motorists cannot rely on the availability of certified secure parking spaces;
-	The current supply of non-secure parking spaces is more evenly distributed across the network. However, these spaces are not certified and do not offer guaranteed services to drivers.

Truck Parking Europe helps transport companies ensure the safety of drivers and freight. Inspection reports provide information on the safety features in truck parking areas. These audit reports help safety and planning managers to approve safe parking areas across Europe (in line with the standardised safety framework created by the EU). They provide a detailed overview of the facilities that each truck parking area offers (Truck Parking Europe, no date).

### Relevant initiatives in Austria {-}

- [truckparkingeurope-austria](https://www.truckparkingeurope.com/secure-truck-parking-overview/austria/) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table70, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Increase safety and security of drivers|**+**|Health & Wellbeing (*3*)|van Weenen et al., 2019|
|Systemic|Reduced loses ralated to cargo thefts|**+**|Sustainable economic development (*8,11*)|van Weenen et al., 2019|
|Systemic|Additional focus on the alternative fuel infrastructure|**+**|Innovation & Infrastructure (*9*)|Gnap & Kubikova, 2020|')
```

### Technology and societal readiness level {-}

```{r table71, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|7-9|')
```

### Open questions {-}
1. How to attract the interest of motorway companies, both private and public, to facilitate data provision, upgrading of parking sites for security and reservation, and encourage the infrastructure construction to capture dynamic data?
2. What software infrastructure is needed to allow for in-vehicle up-to-date information about the occupancy and reservation of truck parking areas?


### Further links {-}

- [truckparkingeurope.com-1](https://www.truckparkingeurope.com/secure-truck-parking-overview/)
- [truckparkingeurope.com-2](https://www.truckparkingeurope.com/how-do-i-reserve-a-secure-truck-parking-spot/)
- [ec.europa.eu-1](https://ec.europa.eu/transport/themes/its/road/action_plan/intelligent-truck-parking_en)
- [ec.europa.eu-2](https://ec.europa.eu/transport/sites/transport/files/modes/road/parking/doc/2010_04_28_setpos_project_handbook.pdf) 
- [ec.europa.eu-3](https://ec.europa.eu/transport/sites/transport/files/2019-study-on-safe-and-secure-parking-places-for-trucks.pdf) 
- [tis-gdv.de](https://www.tis-gdv.de/tis_e/bedingungen/parken/lpp-htm/)
- [unece.org](https://unece.org/DAM/trans/doc/2008/sc1/ECE-TRANS-SC1-103-pres01e.pdf)


### References {-}
-	Carrese, S., Mantovani, S., & Nigro, M. (2014). A security plan procedure for Heavy Goods Vehicles parking areas: An application to the Lazio Region (Italy). Transportation Research Part E: Logistics and Transportation Review, 65(1), 35–49. https://doi.org/10.1016/j.tre.2013.12.011
-	Gnap, J., & Kubíková, S. S. (2020). Possible Effects of Lacking Parking Areas for Road Freight Transport on Logistics and Transport Safety. Transportation Research Procedia, 44(2019), 53–60. https://doi.org/10.1016/j.trpro.2020.02.009
-	Lozia, Z., & Kulma, J. (2014). Simulation method of comparative evaluation of the agility of a passenger car when moving ‘forwards’ and ‘backwards’. Archiwum Motoryzacji, 64, 49–64, 149.
-	Nevland, E. A., Gingerich, K., & Park, P. Y. (2020). A data-driven systematic approach for identifying and classifying long-haul truck parking locations. Transport Policy, 96, 48–59. https://doi.org/10.1016/j.tranpol.2020.04.003
-	Sochor, J., & Mbiydzenyuy, G. (2013). Assessing the benefits of intelligent truck parking. International Journal of Intelligent Transportation Systems Research, 11(2), 43-53.
-	Truck Parking Europe. (n.d.). Audit reports on secure truck parking locations - Truck Parking Europe. Retrieved 3 March 2021, from https://www.truckparkingeurope.com/2020/03/11/audit-reports-on-secure-truck-parking-locations/
-	van Weenen, R. de L., Newton, S., Menist, M., Maas, F., Penasse, D., Nielsen, M., Halatsis, A., Männistö, T., Stamos, I., & Ruschin, P. P. (2019). Study on Safe and Secure Parking Places for Trucks (Issue February).

## Smart delivery space booking {#space_book} 

### Synonyms {-}
*on-street loading zones (OLZ), Freight Trip Generation (FTG), loading/unloading zones (L/U) *

### Definition {-}
In urban areas, goods deliveries for commercial and residential customers generate a huge flow of vehicles ranging from small vans (for express deliveries) to trucks (for deliveries to larger shops). Delivery services are increasing due to a number of factors such as just-in-time management, the development of e-commerce and the emergence of new customer behaviours  for example home delivery, driving and delivery lockers (Patier et al., 2014).
Urban distribution usually requires delivery vehicles to stop temporarily at the side of the road to allow the driver to complete the last part of the delivery on foot. The stops take place in designated areas called loading/unloading (L/U) zones, each with a certain number of available parking spaces. Distributors often face bottlenecks in the L/U areas due to a variety of factors, such as peaks in delivery demands in terms of time or space, systemic problems in the allocation of L/U areas (e.g. scarcity) or poor placement of spaces (Mor et al., 2020).
All this complicates the movement of goods in city centres, with negative effects on vehicle traffic and public transport, for example by causing congestion. In European cities (especially in France), a significant proportion of double-parked delivery trucks (delivery trucks parked on the street parallel to parked cars) is observed. Such behaviour leads to congestion, pollution and conflicts between road users. In this context, special areas called "delivery zones" have been created to improve the work of delivery drivers and reduce congestion.  Nevertheless, delivery drivers continue to double park as the delivery areas are regularly occupied by unauthorised vehicles (single cars) (Patier et al., 2014).


### Key stakeholders {-}

- **Affected**: Delivery truck drivers in urban areas, Car drivers, Road and/or pavement users in cities, parcel/goods receivers
- **Responsible**: National governments, Local governments, Transport authorities, Delivery companies

### Current state of art in research {-}
Several works address the problem of sizing and locating L/U areas in an urban environment. Muñuzuri et al., 2017 presents several alternative approaches to estimating the need for loading zones. Pinto et al., 2019 propose to determine the number and location of L/U areas using a "coverage principle" based on the longest distance a delivery person is willing to walk and define the number of spaces in each area based on demand. The impact of parking availability on the cost and operation of commercial vehicles is explored in (Dezi et al., 2010).<br/><br/>
Mor et al., 2020 conclude that double parking is unlikely to disappear from urban areas unless more dedicated L/U spaces are made available at peak times. In turn, however, the lack of available parking spaces for commercial vehicles is not only caused by undersized infrastructure, but often also by the misuse of reserved spaces by private vehicles (Aiura & Taniguchi, 2005; Pinto et al., 2019).<br/><br/>
In Nourinejad et al., (2014), a parking choice simulation for the study of truck parking policies is presented that captures various dimensions of parking activity such as walking distance, congestion impacts and parking search times. Two scenarios based on the Toronto area are presented to validate the model.
The issue of pricing L/U spaces has also been considered in the literature. An auction-based approach for allocating time slots of a single area with multiple parking spaces is considered to optimise the performance of a management system while ensuring fair allocation of L/U zones (Nourinejad et al., 2014). The time preferences and service duration of booking requests are considered in finding the best possible allocation.<br/><br/>
Next, a sizing problem considering L/U range management is also looked at. A two-phase algorithm is proposed for the problem (Letnik et al., 2018). In the first phase, the location of the L/U areas is optimised based on the location of the goods receivers. In the second phase, the algorithm optimises deliveries from outside the city to the L/U areas. L/U areas are assigned to vehicles and if none are available, vehicles are queued. The benefits of a booking system for L/U areas are assessed for the case of Winchester High Street (McLeod & Cherrett, 2011). It is assumed that each vehicle arrives from one of eight possible entry points and stops in an L/U area, with the possibility of transfer to another area if the desired one is not available. A control system to deal with operational problems such as vehicles arriving too early or too late is also presented.<br/><br/>
The management of L/U areas through a booking system would require the involvement of all stakeholders (customers, traders, municipality, traffic police) and their acceptance that it is in everyone's interest that the system works well as it allows for less inconvenience both in terms of delivery delays and traffic flow. In the initial phase of implementation, the traffic police would need to oversee the new system and educate any trader coming with an unregistered vehicle or without registration about the new rules for using the L/U areas. Information for stakeholders could be provided via the Internet (and possibly in some meetings) to share some statistics on the adoption of the system, the occupancy of delivery positions across the hours of the day and days of the week, and to review the problems each category has in using the system and explore corrective measures (Mor et al., 2020).

### Current state of art in practice {-}
Efforts are being made in various cities around the world to introduce a management system for L/U areas. For example, in Barcelona's *DUM programme1*, traders are asked to check in and out of any L/U operation using an SMS or an app in a smartphone, but in September 2020 no reservation could be made (Mor et al., 2020).
Companies such as *Smart Parking Systems*, *Coord*, *ParkUnload* and *Cleverciti* offer solutions for delivery area booking. However, many different terms are still used, such as "curb management", "smart loading zones" or "digital loading zones". <br/><br/>
In 2010, Treviso was awarded the title of "First European Smart City". Through Smart Parking Applications, the city has benefited from a significant increase in revenue from paid municipal parking, greater economic availability and a reduction in traffic and pollution within the historic centre. Parking management within Treviso Municipality was already in an excellent situation in 2009, monitored and supervised by an extremely attentive mobility department.
The real problem of the municipality was, therefore, not the economic aspect, but mainly the excessive traffic and the scarce availability of parking spaces in some areas of the city. For this reason, it was considered appropriate to use smart parking technology to influence urban mobility in connection with parking. Since 2010, Treviso has, therefore, embarked on this technological path together with Intercomp, installing Smart Parking Systems® in the historic centre with the long-term goal of improving mobility, air quality and service for citizens (Smart Parking Systems®, no date).<br/><br/>
In September 2020, the City of Omaha (US) launched its Smart Zone pilot project with *Coord* in five Smart Zones. Overall, the rollout has gone well, although the city and fleets still need to learn how to disseminate information. Fleets have embraced the concept well in particular that they do not pay to use the Smart Zones. It is expected that once more data is collected and the concept becomes more established the city will introduce the charge (Brown, 2020).

### Relevant initiatives in Austria {-}
In Vienna, more than 2600 loading zones were recorded and located via geodata as part of a project. This data is now available via app *Ladezonen in Wien*. It is not yet possible to make a booking or check the current occupancy rate. The project *i-Ladezone* was first mentioned in 2010 from FFG Austria. In 2012, the project *i-Ladezone - Intelligent Loading Zone Routing and Management* was mentioned again, but only to detect and record the illegal occupancy of loading zones (Stocker, 2012). In 2016, the project *Urban Loading* was submitted to the VCÖ Mobility Award. From a technical point of view, reservation/booking, information to the road user and control/enforcement were considered (VCÖ, 2016). No further information about this project was found.

- [ots.at](https://www.ots.at/presseaussendung/OTS_20150123_OTS0044/simple-stressfreie-ladezonensuche-wk-wien-praesentiert-neue-app) 
- [wko.at-1](https://www.wko.at/service/verkehr-betriebsstandort/Ladezonen-Nutzung.html) 
- [wko.at-2](https://www.wko.at/service/w/verkehr-betriebsstandort/ladezone-wien-app.html) 
- [oeamtc.at](https://www2.ffg.at/verkehr/projekte.php?id=805&lang=de&browse=programm) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table82, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Congestion, pollution and conflicts between road users can be reduced|**+**|Health & Wellbeing (*3*)|Patier et al., 2014; Alho et al., 2018; Mor et al., 2020|
|Systemic|Digitalised solution for occupancy and availability status|**+**|Innovation & Infrastructure (*9*)|Smart Parking Systems, no date|
|Systemic|L/U booking systems requires engagement of all stakeholders|          **+**|Partnership & collaborations (*17*)|Mor et al., 2020|')
```

### Technology and societal readiness level {-}

```{r table83, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|5-8|')
```

### Open questions {-}
1. How to improve the information dissemination and enable the engagement of all stakeholders to develop an efficient L/U booking system?
2. What are the solutions for enforcement of legislation beyond traffic officers?

### Further links {-}

- [areaverda.cat](https://www.areaverda.cat/index.php/es)
- [smartparkingsystems.com](https://smartparkingsystems.com/en/loading-parking-bay-management/)
- [smartloadingzone.com](https://www.smartloadingzone.com/en_us/)
- [coord.com](https://www.coord.com/blog/comparing-loading-zones)

### References {-}
-	Aiura, N., & Taniguchi, E. (2005). PLANNING ON-STREET LOADING-UNLOADING SPACES CONSIDERING THE BEHAVIOUR OF PICKUP-DELIVERY VEHICLES. Journal of the Eastern Asia Society for Transportation Studies, 6, 2963–2974. https://doi.org/10.11175/easts.6.2963
-	Alho, A. R., de Abreu e Silva, J., de Sousa, J. P., & Blanco, E. (2018). Improving mobility by optimizing the number, location and usage of loading/unloading bays for urban freight vehicles. Transportation Research Part D: Transport and Environment, 61, 3–18. https://doi.org/10.1016/j.trd.2017.05.014
-	Brown, C. (2020, October 13). Should Delivery Fleets Pay to Park in Loading Zones? - Smart Cities - Fleet Forward. https://www.fleetforward.com/10127896/should-delivery-fleets-pay-to-park-in-loading-zones
-	Dezi, G., Dondi, G., & Sangiorgi, C. (2010). Urban freight transport in Bologna: Planning commercial vehicle loading/unloading zones. Procedia - Social and Behavioral Sciences, 2(3), 5990–6001. https://doi.org/10.1016/j.sbspro.2010.04.013
-	Letnik, T., Farina, A., Mencinger, M., Lupi, M., & Božičnik, S. (2018). Dynamic management of loading bays for energy efficient urban freight deliveries. Energy, 159, 916–928. https://doi.org/10.1016/j.energy.2018.06.125
-	McLeod, F., & Cherrett, T. (2011). Loading bay booking and control for urban freight. International Journal of Logistics Research and Applications, 14(6), 385–397. https://doi.org/10.1080/13675567.2011.641525
-	Mor, A., Speranza, M. G., & Viegas, J. M. (2020). Efficient loading and unloading operations via a booking system. Transportation Research Part E: Logistics and Transportation Review, 141, 102040. https://doi.org/10.1016/j.tre.2020.102040
-	Muñuzuri, J., Cuberos, M., Abaurrea, F., & Escudero, A. (2017). Improving the design of urban loading zone systems. Journal of Transport Geography, 59, 1–13. https://doi.org/10.1016/j.jtrangeo.2017.01.004
-	Nourinejad, M., Wenneman, A., Habib, K. N., & Roorda, M. J. (2014). Truck parking in urban areas: Application of choice modelling within traffic microsimulation. Transportation Research Part A: Policy and Practice, 64, 54–64. https://doi.org/10.1016/j.tra.2014.03.006
-	Patier, D., David, B., Chalon, R., & Deslandres, V. (2014). A New Concept for Urban Logistics Delivery Area Booking. Procedia - Social and Behavioral Sciences, 125, 99–110. https://doi.org/10.1016/j.sbspro.2014.01.1459
-	Pinto, R., Lagorio, A., & Golini, R. (2019). The location and sizing of urban freight loading/unloading lay-by areas. International Journal of Production Research, 57(1), 83–99. https://doi.org/10.1080/00207543.2018.1461269
-	Smart Parking Systems®. (n.d.). The economic benefits of Smart Parking Systems® > Smart Parking Systems - Intercomp Innovation. Retrieved 21 March 2021, from https://smartparkingsystems.com/en/the-economic-benefits-of-smart-parking-systems/
-	Stocker, G. (2012, November). Snizek+Partner Verkehrsplanungs GmbH - iLadezone - Intelligentes Ladezonenrouting und Management. https://www.snizek.at/go/de/projects/simulationen/94-iladezonen-intelligentes-ladezonenrouting-und-management
-	VCÖ. (2016). Urban Loading - VCÖ Vorbildhafte Mobilitätsprojekte. https://mobilitaetsprojekte.vcoe.at/urban-loading



## Delivery drones {#delivery_drone}

### Synonyms {-}
*urban air mobility (UAM), vertical take-off and landing (VTOL), unmanned aerial vehicles (UAVs), Unmanned Traffic Management (UTM)/U-Space*

### Definition {-}
In the recent year, online shopping and the demand for same-day deliveries has grown exponentially, with more and more people preferring online shopping instead of buying items stationary (Di Puglia Pugliese et al., 2020). At the same time, the last mile is the most expensive process in distribution logistics, contributing to 13% - 73% of the total distribution cost. As a result there is an increased pressure for more efficient solutions in the field.  The answer to this problem could be drones, or unmanned aerial vehicles (UAVs). Drones combine three key principles of technological modernity - computing, autonomy and limitless mobility. Capabilities that until now could only be used by the military are becoming accessible to most of the population. Potential use for drones ranges from surveillance and reconnaissance missions to novel forms of logistics and personal transport. The commercial use of drones is associated with enormous economic opportunities. Even though drones are already common as surveillance/sensing devices in security services, geodesy or agriculture, their use as a means of transport is still at the beginning. Delivery drones are currently able to lift weights of up to 2-3 kg and carry out flight assignments in an urban space (Kellermann et al., 2020).<br/>
What is more, an accelerated progress is expected in the near future, where the European Commission estimates the economic impact of the wider use of drones at €10 billion annually until 2035 and foresees the creation of more than 100,000 direct jobs. Taking into account indirect macroeconomic effects in drone-related industries, the Commission even projects 250,000 to 400,000 additional jobs (SESAR, 2016).

### Key stakeholders {-}

- **Affected**: Mobile citizen, delivery companies and truck drivers, customers of online shops
- **Responsible**: National governments, city government,  private companies, online shops

### Current state of art in research {-}
Current research shows that drones reduce noise, congestion in urban areas and CO~2~ emissions. Multiple studies showed that the use of drones in logistics processes can reduce the lead time (defined as the latency between the beginning and the completion of the process), time span and transport costs of the delivery process (Di Puglia Pugliese et al., 2020; Wang et al., 2017; Di Puglia Pugliese & Guerriero, 2017). However, it is also important that drones have lower capacity and working time (low flight endurance) than classic vehicles. 
Besides all the benefits that drones can bring, there are also some potential drawbacks associated with an increased use of drones. These include safety and security issues, aerial collisions, crashes, malfunctioning of software and hardware components, misuse of drones for criminal purposes, dangers to wildlife and privacy violations, among others (Kellermann et al., 2020).
Moreover, the media analysis about drones for parcel and passenger transportation found the following thematic priorities in relevant literature (see table below): 

```{r table29, echo=FALSE, results='asis', fig.cap="Adapted from Kellermann, Biehle and Fischer, 2020"}
cat('| Topics| Percentage|Number of studies|
|:------------:|:-----------:|:------------:|
|General Surveys    |   18.9%  |  21   |
|Logistics (general)  |  18.0%   |  20   |
|Attitude and Acceptance Research    | 13.5%    |  15   |
|Law and Regulations    | 11.7%    |   13  |
|Ethics and Technology Assessment    | 10.8%    |  12   |
|Sustainability Assessment    | 8.1%    |  9   |
|Urban and Transportation Planning    |  7.2%   |  8   |
|Political agenda/strategies    |  6.3%   |   7  |
|Passenger Transportation   |  2.7%   |  3   |
|Humanitarian Logistics    |2.7%     |   3  |') 
```

The research proposes two approaches in terms of the incorporation of drones into delivery process, either as a sole delivery option or as a support system for trucks. Calculation results show the disadvantage of the exclusive use of drones in the delivery process. Overall, the vehicle routing that uses a combination of trucks and drones is the transport system offers the best trade-off between transport costs, CO~2~  emissions and congestion (Di Puglia Pugliese et al., 2020). For example, with 2 drones per truck, the delivery completion time can be reduced by 75% in the best case scenario (Wang et al., 2017). <br/> 

Current research also examines the potential of drones in food delivery services due to COVID-19 (Kim et al., 2021). A study on choice-model-based analysis of consumer preference for a drone delivery service concludes the following (Kim, 2020):

-	Drone delivery service competes with truck delivery service for clothing and powder formula, and with motorbike delivery service for urgent documents. 
-	Potential customers' preferences for drone delivery service depend on the price and type of goods. 
-	Consumers are concerned about the reliability of the drone delivery service for expensive items. 
-	In all models, it is observed that young people are more likely to choose drone delivery service than old people.

### Current state of art in practice {-}
Several large companies, such as *Amazon* (Amazon.com, 2021), *DHL* (Di Puglia Pugliese et al., 2020), *UBER* , *Wing* (Bonnington, 2012), *UPS* (UPS, 2021) and some more have started using delivery drones for parcel delivery. However, governments and experts fear the misuse of drones. In particular, after their concerns materialised with the incident at London Airport (Keane, 2018). To protect against misuse, work is being carried out internationally on a digital remote ID so that the drone operators can always be assigned to the drone (Boyle, 2020). <br/> <br/>
In the USA, the Federal Aviation Administration (FAA) has allowed drone flights as long as the drone is within an operator's line of sight. However, companies are working on developing drones to fly without this requirement (Fox Rubin, 2019). Further, UPS is already carrying out drone deliveries between WakeMed's main hospital campus in Raleigh, North Carolina, and its neighbouring laboratory centre to send laboratory samples back and forth. The company is also planning flights between hospitals to provide easier access to life-saving drugs like anti-venom (Fox Rubin, 2019). The fact that the first drone experiments have focused on business-to-business deliveries in the healthcare sector is not a coincidence. Firstly, funding is available in this sector, and secondly, the speed of the drones can bring considerable benefits in this field (Marshall, 2020).<br/> <br/>
Moreover, Amazon announced back in 2013 that they would start delivering with drones within years and in 2019, alongside the unveiling of their new drone, they announced that they would start drone delivery "within months" (Kolodny & Feiner, 2019). Nonetheless, it is currently claimed that even with all the test flights and announcements, it will be several years before a drone can deliver (Marshall, 2020).

### Relevant initiatives in Austria {-}
One of the European Amazon drone development centres is localised in Austria (Marshall, 2020). Moreover, a new drone law came into force in the EU on 31 December 2020 to standardise the rules. It requires a CE marking on the drone and a drone licence for drones weighing 250g or more.

- [www.oesterreich.gv.at-1](https://www.oesterreich.gv.at/themen/dokumente_und_recht/Drohnen.html) 
- [www.oesterreich.gv.at-2](https://www.oesterreich.gv.at/themen/dokumente_und_recht/Drohnen/Drohnenfuehrerschein.html) 
- [www.oesterreich.gv.at-3](https://www.oesterreich.gv.at/themen/dokumente_und_recht/Drohnen/EU-Regelungen-f%C3%BCr-Drohnen-im-%C3%9Cberblick.html) 
- [oeamtc.at](https://www.oeamtc.at/thema/drohnen/drohnen-info-app-26853120) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table30, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Faster provision of time-sensitive goods (eg. pharmaceuticals)|**+**|Health & Wellbeing (*3*)|Fox Rubin, 2019|
|Individual|Drone deliveries offered at no extra cost|          **+**|Sustainable economic development (*8,11*)|Amazon.com, 2021|
|Individual|New delivery options available|          **+**|Innovation & Infrastructure (*9*)|Kim, 2020; Amazon.com, 2021|
|Systemic|Reduced time of completion of logistics processes|          **+**|Sustainable economic development (*8,11*)|Wang et al., 2017|
|Systemic|Standardised drones legislation across the EU  |          **+**|Partnership & collaborations (*17*)|EASA, 2021|')
```

### Technology and societal readiness level {-}

```{r table31, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-8|3-5|')
```

### Open questions {-}
1. Will private companies work together and share their technologies?
2. What other areas of application will there be?
3. What will be the thematic priorities for development in the coming years?
4. Will freight drones automatically communicate with other drones such as surveillance, geodesy and agricultural drones?

### Further links {-}

-	[easa.europa.eu](https://www.easa.europa.eu/sites/default/files/dfu/Easy%20Access%20Rules%20for%20Unmanned%20Aircraft%20Systems%20(Regulations%20(EU)%202019-947%20and%20(EU)%202019-945).pdf) 

### References {-}
- Amazon.com: Prime Air. (n.d.). Retrieved 27 January 2021, from https://www.amazon.com/Amazon-Prime-Air/b?ie=UTF8&node=8037720011
-	Bonnington, C. (2012, March 23). Tacocopter: The Coolest Airborne Taco Delivery System That’s Completely Fake | WIRED. https://www.wired.com/2012/03/qa-with-tacocopter/
-	Boyle, A. (2020, May 6). Amazon and T-Mobile will take part in FAA’s Remote ID program for drones - GeekWire. https://www.geekwire.com/2020/amazon-t-mobile-will-take-part-faas-remote-id-drone-monitoring-program/
-	Di Puglia Pugliese, L., & Guerriero, F. (2017). Last-Mile Deliveries by Using Drones and Classical Vehicles. In A. Sforza & C. Sterle (Eds.), Optimization and Decision Science: Methodologies and Applications (pp. 557–565). Springer International Publishing.
-	Di Puglia Pugliese, L., Guerriero, F., & Macrina, G. (2020). Using drones for parcels delivery process. Procedia Manufacturing, 42, 488–497. https://doi.org/10.1016/j.promfg.2020.02.043
- EASA (2021). Easy Access Rules for Unmanned Aircraft Systems (Regulations (EU) 2019/947 and (EU) 2019/945) Available at: https://www.easa.europa.eu/sites/default/files/dfu/Easy%20Access%20Rules%20for%20Unmanned%20Aircraft%20Systems%20(Regulations%20(EU)%202019-947%20and%20(EU)%202019-945).pdf. (Accessed: 08 February 2021)
-	Fox Rubin, B. (2019). UPS drone delivers medicine from CVS straight to customers’ homes. https://www.cnet.com/news/ups-drone-deliver-medicine-from-cvs-straight-to-customers-homes/
-	Keane, S. (2018). London’s Gatwick Airport closes again after drones force 33-hour shutdown. https://www.cnet.com/news/airport-suspends-all-flights-after-drones-get-too-close-to-airfield/
-	Kellermann, R., Biehle, T., & Fischer, L. (2020). Drones for parcel and passenger transportation: A literature review. Transportation Research Interdisciplinary Perspectives, 4, 100088. https://doi.org/10.1016/j.trip.2019.100088
-	Kim, J. J., Kim, I., & Hwang, J. (2021). A change of perceived innovativeness for contactless food delivery services using drones after the outbreak of COVID-19. International Journal of Hospitality Management, 93(June 2020), 102758. https://doi.org/10.1016/j.ijhm.2020.102758
-	Kim, S. H. (2020). Choice model based analysis of consumer preference for drone delivery service. Journal of Air Transport Management, 84(September 2019), 101785. https://doi.org/10.1016/j.jairtraman.2020.101785
-	Kolodny, L., & Feiner, L. (2019, June 5). Amazon debuts its new delivery drone. https://www.cnbc.com/2019/06/05/amazon-debuts-its-new-delivery-drone.html
-	Marshall, A. (2020, May 8). No, Amazon Won’t Deliver You a Burrito by Drone Anytime Soon | WIRED. https://www.wired.com/story/amazon-wont-deliver-burrito-drone-soon/
-	SESAR. (2016). European Drones Outlook Study. In Single European Sky ATM Research (Issue November). https://doi.org/10.2829/219851
-	UPS. (n.d.). UPS Drone Delivery Service | UPS - United States. Retrieved 4 February 2021, from https://www.ups.com/us/en/services/shipping-services/flight-forward-drones.page
-	Wang, X., Poikonen, S., & Golden, B. (2017). The vehicle routing problem with drones: several worst-case results. Optimization Letters, 11(4), 679–697. https://doi.org/10.1007/s11590-016-1035-3

## Electric vehicle delivery fleets {#electric_delivery_fleets}

### Synonyms {-}
*Electric commercial vehicle (ECV)*

### Definition {-}
The majority of current conventional transport vehicles are highly dependent on fossil fuels, leading to an increase in greenhouse gas (GHG) emissions. Although environmentally friendly transport alternatives exist, their number is still limited, especially in developing countries. Therefore, the governments, authorities and organisations need to take precautions to reduce the use of these fuels or limit emissions. In 2011, the European Commission published draft legislation on transport to define its goals for a competitive and efficient energy system by achieving zero-emission urban freight transport by 2030. According to this, the European Commission's goal is to halve the use of conventionally powered vehicles in urban freight transport in the medium term and phase them out by 2030. It is important to introduce the e-vehicles in commercial fleets because road transport constitutes a largest proportion of the overall inland freight while the e-vehicles are currently mainly used in ride-sharing and public transport (Bac & Erdem, 2021). <br/>

From November 2020, one in ten new cars sold in Europe will be a pure electric or plug-in hybrid to achieve a milestone of one million electric vehicles (EVs) being sold. This is a crucial turning point towards achieving 30-40% EV sales volumes by 2030, bringing Europe's carbon reduction targets within reach. The environmental benefits are the biggest prize. <br/>

In twenty in-depth interviews with executives from a wide range of industries, overwhelming optimism was found to electrify their fleet. One utility company, for example, plans to electrify its entire commercial fleet of 10,000 vehicles by 2026. Then, a battery manufacturer makes significant investments to build low-carbon, lithium-ion battery production site in Europe. Moreover, the electricity grid operators are pursuing the goal of adding value to mass transport electrification through the use of smart charging and vehicle-to-grid technologies (Colle et al., 2020).<br/>

However, there are many factors such as the limited mileage of these vehicles, long recharging times and low availability of charging stations (for more details on [electric charging stations](#charging_station)) that diminish the benefits of EVs in industrial and commercial logistics. Effective planning of EV routes and charging times is crucial for the future of the logistics sector (Bac & Erdem, 2021).


### Key stakeholders {-}

- **Affected**: Commercial vehicle drivers, freight industry
- **Responsible**: International, national and local authorities, Public and private transport companies, Automobile manufacturers

### Current state of art in research {-}
Recent studies on ECVs can be divided into two streams focusing on *(i)* aggregated cost analysis with TCO calculations to investigate the competitiveness of ECVs and *(ii)* decision support models for both integrated network design and fleet operation based on location routing problems (LRP) and vehicle routing problems (VRP) (Schiffer et al., 2021).

- Aggregated cost analyses

In order to assess the competitiveness of ECVs, it is common to perform aggregated TCO calculations to compare the life cycle costs of ECVs and ICEVs. Analyses by Davis & Figliozzi (2013), Lee et al., (2013) and Taefi et al., (2017) conclude that ECVs become more competitive with increasing distance travelled. In addition, characteristics of certain driving patterns such as frequent stops, congestion, engine idling and low speed increase the competitiveness of ECVs (Schiffer et al., 2021).

- Vehicle routing and location-routing problems

The major problems faced by the fleets are the optimisation of the electric vehicle charging schedule and the routing problem. For given problems, an electric vehicle routing problem with time windows (EVRPTW) is proposed, which is an extension of the well-known vehicle routing problem (VRP) (Bac & Erdem, 2021). Another EV-related problem is the location selection of charging stations. In the literature, there are a multiple studies proposing different approaches such as multi-criteria decision models (Feng et al., 2021; Wang et al., 2020) or genetic algorithm (GA) applications (Pan et al., 2020). Furthermore, Sachan et al., (2020) iexplored in this regard different charging infrastructures categorized into traditional charging infrastructures, fast-charging stations, and battery swapping stations.<br/>

Optimising electric vehicle charging to minimise marginal GHG emissions from electricity generation is also being investigated (Tu et al., 2020). A high penetration rate of EVs and night-time charging can potentially lead to periodic electricity demand congestion (Shaukat et al., 2018) and high marginal emissions from electricity generation. Optimal spatio-temporal distribution of EV charging activities should enable more environmentally friendly charging. The optimal charging schedule can be further improved with the help of a smart grid and vehicle-to-grid communication.

### Current state of art in practice {-}
In Europe, about 75% of fleet vehicles are spread across six key industries (Colle et al., 2020):

-	Wholesale and retail
-	Public administration and defence
-	Manufacturing
-	Construction
-	Transport including taxis, buses and coaches
-	Logistics
-	Vehicle rental and leasing

With currently 63 million vehicles, the fleet (comercial vehicles) accounts for 20% of all vehicles in Europe. Fleet vehicles cover more than 40% of total vehicle kilometres in Europe and are responsible for half of total road transport emissions. Six out of ten cars sold in Europe are company cars and in 2019, 96% of new company car registrations were petrol or diesel vehicles. On average, company cars travel 2.25 times further than private cars, contributing disproportionately to emissions. It is predicted that the total electrified fleet will increase 24-fold by 2030. National and local government incentives, the ability to negotiate discounts on vehicle contracts and route predictability make the choice of fleet electrification more popular (Colle et al., 2020).<br/>

All major delivery companies are starting to replace their gas-powered fleets with electric or low-emission vehicles, a switch will boost profits while fighting climate change and urban pollution. UPS has placed an order for 10,000 electric delivery vehicles. Amazon is buying 100,000 from start-up *Rivian*. DHL says one-fifth of its fleet is made up of zero-emission vehicles, with more to come. FedEx has just committed to replacing 100% of its pick-up and delivery fleet with battery-powered vehicles by 2040 (Domonoske, 2021).<br/>

One of the reasons why delivery fleets are increasingly investing in electric or zero-emission vehicles is the increasing government mandates to reduce emissions. Regulations can become even stricter in the future (Hughes, 2020). In Europe, 24 cities with a combined population of over 62 million plan to completely ban internal combustion engine (ICE) vehicles from urban areas by 2030 through a combination of regulations and incentives (Behrmann, 2019). However, more charging stations will have to be built to meet the increaseing demand. There are already about 285.800 charging stations in Europe (Statista, 2021). Nevertheless, logistics fleet operators have so far tended not to include public charging infrastructure in their planning in order to avoid operational disruptions that can occur in the event of blocked stations or vandalism (Schiffer et al., 2021).<br/>

Currently, electric cars, including company cars, pool cars, rental cars, ride-hailing vehicles and taxis, account for more than half (59%) of the total EV fleet. The light commercial vehicle (LCV) or van segment, which includes last-mile vehicles and delivery and service vehicles, accounts for 38%. The remainder is composed of heavy-duty vehicles (HDVs) and buses. The bus segment is expected to electrify the fastest, with 42% of the total bus fleet electrified by 2030. The passenger car segment follows with 17.5% by 2030, indicating a shift in company car policy towards electric vehicles and greater vehicle choice. More than 12% of the van segment will be electric by 2030, compared to only 2% of HDVs. Electric HDVs will, for the time being, be limited to vehicles under 7.5 tonnes and special and dedicated service vehicles, such as refuse collection and street sweeping vehicles (Colle et al., 2020).

### Relevant initiatives in Austria {-}
Austrian Post aims to exclusively use 100% electric vehicles by 2030, a process initiated in 2011. Today, Austrian Post already operates the largest e-vehicle fleet in Austria and has a comprehensive charging network at its disposal (Uyttebroeck, 2021). The commitment to convert their fleets was formalised when the two groups joined EV100, an initiative that aims to encourage and support fleets to convert their fleets to electric vehicles by 2030 (Field, 2019).

-	[theclimategroup.org](https://www.theclimategroup.org/ev100)
-	[cleantechnica.com](https://cleantechnica.com/2019/04/27/the-swiss-austrian-post-announce-commitment-to-100-evs-by-2030/)
-	[beoe.at](https://www.beoe.at/about-us/)
-	[ots.at](https://www.ots.at/pressemappe/17440/bundesverband-elektromobilitaet-oesterreich-beoe)

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table91, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Lower pollutant and noise emissions|**+**|Health & Wellbeing (*3*)|Bundesministerium fuer Umwelt, 2019|
|Systemic|Reduced overall level of emissions|**+**|Environmental sustainability *(7,12-13,15)*|Fritz et al., 2018|
|Systemic|Very limited use in large commercial vehicles|**~**|Sustainable economic development (*8,11*)|Colle et al., 2020|
|Systemic|Electric Vehicles Initiative is accelerating the introduction and adoption of electric vehicles worldwide|          **+**|Partnership & collaborations (*17*)|IEA, 2020|')
```

### Technology and societal readiness level {-}

```{r table92, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|8-9|')
```

### Open questions {-}
1. Could battery exchange stations play a role in electric city logistics?

### References {-}
-	Bac, U., & Erdem, M. (2021). Optimization of electric vehicle recharge schedule and routing problem with time windows and partial recharge: A comparative study for an urban logistics fleet. Sustainable Cities and Society, 102883. https://doi.org/10.1016/j.scs.2021.102883
-	Behrmann, E. (2019, July 26). A Dead End for Fossil Fuel in Europe’s City Centers - Bloomberg. https://www.bloomberg.com/news/articles/2019-07-26/a-dead-end-for-fossil-fuel-in-europe-s-city-centers
-	Bundesministerium für Umwelt, N. und nukleare S. (BMU). (2019). Wie umweltfreundlich sind Elektroautos? 1–20. www.bmu.de
-	Colle, S., Miller, R., Mortier, T., Coltelli, M., Horstead, A., Ruby, K., & Georgiev, P. (2020). Accelerating fleet electrification in Europe When does reinventing the wheel make perfect sense ?
-	Davis, B. A., & Figliozzi, M. A. (2013). A methodology to evaluate the competitiveness of electric delivery trucks. Transportation Research Part E: Logistics and Transportation Review, 49(1), 8–23. https://doi.org/10.1016/j.tre.2012.07.003
-	Domonoske, C. (2021, March 17). From Amazon To FedEx, The Delivery Truck Is Going Electric. https://text.npr.org/976152350
-	Feng, J., Xu, S. X., & Li, M. (2021). A novel multi-criteria decision-making method for selecting the site of an electric-vehicle charging station from a sustainable perspective. Sustainable Cities and Society, 65, 102623. https://doi.org/10.1016/j.scs.2020.102623
-	Field, K. (2019, April 27). The Swiss & Austrian Post Announce Commitment To 100% EVs By 2030. https://cleantechnica.com/2019/04/27/the-swiss-austrian-post-announce-commitment-to-100-evs-by-2030/
-	Fritz, D., Heinfellner, H., Lichtblau, G., Pölz, W., & Stranner, G. (2018). Update: Ökobilanz alternativer Antriebe. 1–16. https://umweltbundesamt.at/fileadmin/site/publikationen/DP152.pdf
-	Hughes, M. (2020, December 5). Electric Delivery Fleet Trends: How Green Are Your Deliveries? | ChargePoint. https://www.chargepoint.com/blog/electric-delivery-fleet-trends-how-green-are-your-deliveries/
-	IEA. (2020, October 26). Electric Vehicles Initiative – Programmes and partnerships - IEA. https://www.iea.org/areas-of-work/programmes-and-partnerships/electric-vehicles-initiative
-	Lee, D.-Y., Thomas, V. M., & Brown, M. A. (2013). Electric Urban Delivery Trucks: Energy Use, Greenhouse Gas Emissions, and Cost-Effectiveness. Environmental Science & Technology, 47(14), 8022–8030. https://doi.org/10.1021/es400179w
-	Pan, L., Yao, E., Yang, Y., & Zhang, R. (2020). A location model for electric vehicle (EV) public charging stations based on drivers’ existing activities. Sustainable Cities and Society, 59, 102192. https://doi.org/10.1016/j.scs.2020.102192
-	Sachan, S., Deb, S., & Singh, S. N. (2020). Different charging infrastructures along with smart charging strategies for electric vehicles. Sustainable Cities and Society, 60, 102238. https://doi.org/10.1016/j.scs.2020.102238
-	Schiffer, M., Klein, P. S., Laporte, G., & Walther, G. (2021). Integrated planning for electric commercial vehicle fleets: A case study for retail mid-haul logistics networks. European Journal of Operational Research, 291(3), 944–960. https://doi.org/10.1016/j.ejor.2020.09.054
-	Shaukat, N., Khan, B., Ali, S. M., Mehmood, C. A., Khan, J., Farid, U., Majid, M., Anwar, S. M., Jawad, M., & Ullah, Z. (2018). A survey on electric vehicle transportation within smart grid system. In Renewable and Sustainable Energy Reviews (Vol. 81, pp. 1329–1349). Elsevier Ltd. https://doi.org/10.1016/j.rser.2017.05.092
-	Statista. (2021). • Europe: number of electric vehicle charging stations 2010-2020 | Statista. https://www.statista.com/statistics/955443/number-of-electric-vehicle-charging-stations-in-europe/
-	Taefi, T. T., Stütz, S., & Fink, A. (2017). Assessing the cost-optimal mileage of medium-duty electric vehicles with a numeric simulation approach. Transportation Research Part D: Transport and Environment, 56, 271–285. https://doi.org/10.1016/j.trd.2017.08.015
-	Tu, R., Gai, Y. (Jessie), Farooq, B., Posen, D., & Hatzopoulou, M. (2020). Electric vehicle charging optimization to minimize marginal greenhouse gas emissions from power generation. Applied Energy, 277, 115517. https://doi.org/10.1016/j.apenergy.2020.115517
-	Uyttebroeck, B. (2021, March 21). The Mobility House to manage 100% EV fleet for Austrian Post | Fleet Europe. https://www.fleeteurope.com/en/new-energies/austria/features/mobility-house-manage-100-ev-fleet-austrian-post?a=BUY03&curl=1
-	Wang, R., Li, X., Xu, C., & Li, F. (2020). Study on location decision framework of electric vehicle battery swapping station: Using a hybrid MCDM method. Sustainable Cities and Society, 61, 102149. https://doi.org/10.1016/j.scs.2020.102149



## Multimodal transport management systems {#mtms}

## Freight hubs {#freight_hubs}

### Synonyms {-}
*Logistics hubs, Logistic hubs, Multimodal transportation hubs*

### Definition {-}
In logistics networks, logistics hubs are generally defined as interconnection points and primarily serve as transshipment points for flows of goods. Thus, not only storage activities take place, but also processes of ordering, bundling and unbundling. The variety of logistics nodes makes it difficult to clearly assign nodes to a specific type and class. However, a simplifying differentiation is often made based on a spatial or functional analysis. An example would be a spatial differentiation e.g. according to the spatial level (micro, meso, macro) according to which a transport logistics node can be defined as a hub (micro level) or a seaport (macro level). Another example would be a functional differentiation, according to which logistics hubs can consist of single modules (e.g., single shipping facility) or multiple modules (intermodal terminal with rail freight center and forwarders). A very simple but basic differentiation of logistics hubs can be found in the subdivision into transportation logistics hubs and distribution logistics hubs, as they differ fundamentally in numerous characteristics (see table below (Clausen & Thaller, 2013)) (Huber et al., 2015). <br/>

```{r , echo=FALSE, results='asis'}
cat('|Characteristics| Distribution logistics hub|Transport logistics hub|
|:----:|:----:|:----:|
|Number of sources|few|many|
|Number of sinks|many|many|
|Main function|storage, consolidation, distribution, packaging, value adding services|transhipment, certain buffer function|
|User|one or certain number of users|many different and changing customers|
|Operator|own account or logistics service provider (3PL)|logistics service provider, forwarding agency|
|Destination|sink is uncertain*|sink is certain**|
|Examples|distribution and/or consolidation centres, warehouses|intermodal freight terminals, locations of forwarding companies, seaports, inland ports, airports|')
```

*At the moment the goods arrive at the distribution logistics hub and the final destination is not defined. The final destination will be defined during the picking process only.<br/>
**Each shipment is labelled with an explicit destination before it reaches the transport logistics hub.<br/>

Distribution logistics hubs have warehouses where goods can be stored for an extended period of time. As a rule, these types of hubs combine a few sources with many sinks. Central or regional warehouses are an example - there are a few companies that deliver goods to the warehouse (sources) and many individuals and companies that order from the warehouse (sinks). Transportation logistics hubs, in contrast, have no storage function. There may be some buffer capacity due to the transshipment process, but this buffer is a side effect of the actual main function - the transshipment of goods. Transportation logistics hubs typically connect many sources to many sinks. Examples of this type of hub can be found at freight forwarder locations, combined transport terminals, airports, and rail transfer stations. The subdivision of logistics hubs into transportation logistics hubs and distribution logistics hubs is essential for freight demand modeling because the goods attracted and shipments made at these hubs differ significantly (Huber et al., 2015).

Intermodal logistics hubs (see [Intermodal Freight](#intermodal_freight)) attract significant volumes of different types of traffic. Trucks deliver and pick up goods, containers are moved, service and heavy-duty vehicles pass through, and sometimes public roads cross the site. The organic growth of these facilities and the multitude of parties involved often result in an inefficient and inadequate transportation infrastructure, unnecessary negative environmental impacts, and high costs for stakeholders. Many factors need to be known to optimize the infrastructure and operation of these nodes, but often are not. These factors include stakeholder requirements, hours of service at facilities, vehicle origins and destinations, and vehicle routes and speeds (Ehrler & Wolfermann, 2012).<br/>

Essaadi et al. (2016) shows that the location selection of logistic hubs is a strategic decision that is made after a multi-criteria analysis and first, it requires the analysis of quantitative or qualitative criteria (e.g. availability and quality of infrastructure, connectivity, land price, political stability, pollution, etc.). These can be independent or partially contradictory. The study proposes a generic structuring of criteria by geographical level and by category for hub location selection, taking into account the particular structure of location selection that is rarely found in the literature: sequential evaluation (choice of a country, then a region of this country) or simultaneous evaluation (direct choice of a location among several regions belonging to different countries).


### Key stakeholders {-}

- **Affected**: Courier Express Parcel Services, Package Recipients, Carriers
- **Responsible**: Courier Express Parcel Services, Logistics companies, Carriers, City administration, Funding provider, Ministries, Infrastructure managers, Chamber of Commerce 

### Current state of art in research {-}
The main research in the area of freight hubs focuses on the analysis of environmental impact of freight and proposal of sustainable solution as well as the investigation of the potential for improvement in efficiency considering freight hubs locations and demand patterns. <br/>

For example, with zero-emission vehicles (ZEVs) considered key technologies for reducing freight-related air pollution and greenhouse gas emissions, California's 2016 Sustainable Freight Action Plan calls for 100,000 zero-emission, renewable-fuel freight vehicles by 2030. However, as the current hydrogen infrastructure in California is sparse with about 25 stations installed, new infrastructure strategies will be crucial for the implementation of hydrogen freight applications. Li et al. (2021) analyzed the hydrogen infrastructure requirements with a focus on hydrogen fuel cells in freight applications. Using a California-specific EXCEL-based scenario model, the adoption and demand of hydrogen vehicles for trucking, rail, shipping and aviation was estimated for a range of scenarios through 2050. Findings show that ZEVs will need to grow very aggressively to meet state ZEV targets for freight. Moreover, the battery electric trucks (see sections on [Electric vehicle delivery fleets](#electric_delivery_fleets) and [Battery electric](#bev)) appear to have a head start on commercialization, and FCEVs (see section on [Hydrogen fuel cell](#FCEV)) on the other hand, have advantages that put them in a better position in certain markets, for example, long-haul trucking. By 2030, hundreds of tons of hydrogen will be needed per day to power the local hydrogen fleets projected in this study. For long-haul trucking, a nationwide "backbone" hydrogen refueling network could be developed with approximately 13 strategically located refueling stations. Further, there is potential for synergy between refueling infrastructure for local and long-range FCEVs. These synergies may be critical for the initial rollout of medium- and heavy-duty trucks hydrogen refueling stations, as many hydrogen refueling stations may be shared by fleets from more than one hub.<br/>

Further, Markvica et al. (2019) investigated the feasibility of a "quattromodal freight hub", a new hub concept with four modes of transport, in the city of Vienna. In doing so, they highlighted the strengths, weaknesses, opportunities, and threats of the concept from a theoretical and practical perspective and identified four options for creating a quattromodal freight hub in Vienna. They argue that cost and efficiency related decision criteria are crucial for the implementation and that it is attractive in terms of prestige and unique selling point for the region, and that at the same time, however, further research is required regarding legal aspects and impacts on the region.<br/>

Moreover, it has been showed that in the areas surrounding large freight hubs such as seaports, the truck flows can have a major impact on the motorways and their traffic management. Nadi et al. (2021) have therefore developed a model that predicts short-term changes in truck volumes generated by large container terminals in seaports. The model was developed, tested and demonstrated during a case study for the port of Rotterdam. It can predict the outgoing truck volume for the next hour with a high degree of accuracy.<br/>

On the other hand, COVID-19 outbreak had a significant impact on freight globally, where diverse adjustments along the logistic chains were required. Therefore, Huang et al. (2021) aimed to investigate the spatial patterns of freight demand network in six provinces of Central China since the beginning of the COVID-19 pandemic. For this purpose, the Big Data of "Cart Search" demand information provided by small and medium-sized freight companies on a freight information platform was analysed. The results show: 

- the choke point of the unbalanced freight demand network and the secondary choke points
- a chain reaction circle of unbalanced freight traffic
- a long-tail distribution with a wide range and unbalanced distribution
- that the import and export of freight varied significantly in different cities
- that the distribution was unbalanced 

Building on these findings, Huang et al. (2021) propose solutions in case of outbreaks in the freight market in six provinces of central China in the post-epidemic period. They argue that a "freight alliance" can be used to address the hidden dangers of freight imbalance and enable small and medium-sized freight enterprises in peripheral cities to break through and develop in the post-epidemic era.


### Current state of art in practice {-}
Due to the rapid population growth and the loss of business areas in the urban centres in favour of residential areas, the safeguarding and the increase of the utilisation efficiency of the remaining areas for activities related to logistics needs to be actively supported. Therefore, the project "Screening Logistics Areas", which was successfully carried out and completed under the auspices of the City of Vienna and the Province of Lower Austria. It focused on determining the location and area offers for the logistics sector in these two provinces. Areas starting at around 1 ha (Vienna) and 5 ha (Lower Austria) were considered. The available land concepts were used as a basis for the work, hence business parks and industrial zones were examined in detail. All relevant aspects in the area under consideration were examined, categorised and given concrete recommendations. It was determined that there are currently sufficient areas available, both quantitatively and qualitatively, although it is now necessary to find mechanisms to effectively secure the areas for business purposes. On the part of the provinces, the results flow into the examination of possible control instruments as well as various regional planning programmes (ARGE L2030+, n.d.).<br/>

In 2019, DHL Global Forwarding and DHL Freight combined their three previous locations at the new DHL Campus Vienna International Airport, bringing land, air and sea freight together in one location. This allows higher throughput rates for freight volumes and even better service. In addition to office space with around 3,500 square meters, the DHL Campus Vienna Airport includes two freight terminals with a total of around 12,000 square meters, divided into 5,000 square meters of storage and handling space for air and sea freight and around 7,000 square meters for land transport. Given its size, it acts as a strategically important logistics hub for Austria and Eastern Europe. In order to meet the constantly growing demand for logistics and transport services from the life science and healthcare sectors, the new building also created larger capacities for temperature-controlled goods (DHL International GmbH, n.d.). 


### Relevant initiatives in Austria {-}


- [Infrastruktur.oebb.at](https://infrastruktur.oebb.at/en/partners/terminals/locations/terminal-wien-sued) 
- [DHL-freight-connections.com](https://dhl-freight-connections.com/de/unternehmen/dhl-eroffnet-hochmodernes-logistikdrehkreuz-am-flughafen-wien/) 
- [Logistik2030.at](https://www.logistik2030.at/?page_id=63) 
- [Thinkportvienna.at](https://www.thinkportvienna.at/ueber-uns/projekte/) 
- [Hafen-wien.com](https://www.hafen-wien.com/de/home) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r , echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Changes in land use, increase density, and generation of more intensive local and cross-town traffic|          **-**|Sustainable economic development (*8,11*)|Rondinelli & Berry, 2000|
|Systemic|New developments due to novel vehicle technologies used|**+**|Innovation & Infrastructure (*9*)|Li et al., 2021|
|Systemic|Collaboration among stakeholders and states to achieve the greatest benefit for the region|          **+**|Partnership & collaborations (*17*)|ARGE L2030+, n.d.|')
```

### Technology and societal readiness level {-}

```{r, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-8|6-9|')
```

### Open questions {-}
1. What are the additional requirements for freight hubs in Europe with regard to electromobility and hydrogen-powered truck fleets?
2. How can efficiency be increased by combining different modes? 
3. How can the negative externalities of freight hubs - such as light and noise emissions or increased land consumption - be minimized and internalized?

### References {-}
-	ARGE L2030+. (n.d.). Pilotprojekte – Logistik 2030+. Central LogPOINT – DER Logistik HUB im Herzen von Wien. Retrieved April 28, 2021, from https://www.logistik2030.at/?page_id=268
-	Clausen, U., & Thaller, C. (Eds.). (2013). Wirtschaftsverkehr 2013: Datenerfassung und verkehrsträgerübergreifende Modellierung des Güterverkehrs als Entscheidungsgrundlage für die Verkehrspolitik. Springer-Verlag.
-	DHL International GmbH. (n.d.). DHL eröffnet hochmodernen Logistik-Hub am Flughafen Wien - DHL Freight Connections. Retrieved May 4, 2021, from https://dhl-freight-connections.com/de/unternehmen/dhl-eroffnet-hochmodernes-logistikdrehkreuz-am-flughafen-wien/
-	Ehrler, V., & Wolfermann, A. (2012). Traffic at intermodal logistic hubs: shedding light on the blind spot. Transportation research record, 2288(1), 1-8.
-	Essaadi, I., Grabot, B., & Fénies, P. (2016). Location of logistics hubs at national and subnational level with consideration of the structure of the location choice. IFAC-PapersOnLine, 49(31), 155-160.
-	Huang, Y., Liu, R., Huang, S., Yang, G., Zhang, X., Qin, Y., ... & Huang, B. (2021). Imbalance and breakout in the post-epidemic era: Research into the spatial patterns of freight demand network in six provinces of central China. Plos one, 16(4), e0250375.
-	Huber, S., Klauenberg, J., & Thaller, C. (2015). Consideration of transport logistics hubs in freight transport demand models. European Transport Research Review, 7(4), 1-14.
-	Li, G., Ogden, J., & Miller, M. (2021). Hydrogen Infrastructure Requirements for Zero-Emission Freight Applications in California.
-	Nadi, A., Sharma, S., Snelder, M., Bakri, T., van Lint, H., & Tavasszy, L. (2021). Short-term prediction of outbound truck traffic from the exchange of information in logistics hubs: A case study for the port of Rotterdam. Transportation Research Part C: Emerging Technologies, 127, 103111.
-	Markvica, K., Prandtstetter, M., Zajicek, J., Heilmann, B., Lenz, G., Hauger, G., ... & Eitler, S. (2019). Implementing a quattromodal freight hub: an approach for the city of Vienna. European Transport Research Review, 11(1), 1-16.
-	Rondinelli, D., & Berry, M. (2000). Multimodal transportation, logistics, and the environment: managing interactions in a global economy. European Management Journal, 18(4), 398-410.
