# Collective mobility vehicles {#collective}

## Demand responsive transit

## Personal rapid transit

### Synonyms {-}
*Automated Transit Networks, ATN, PRT, Podcars*

### Definition {-}
Personal rapid transit is a public transportation option in urban areas where fully automated vehicles provide on-demand, origin-to-destination service over an area network. PRT belongs to a larger family of automated guideway transit (AGT). It is a unique mode of transport that has the following features (Furman et al., 2014; Juster & Schonfeld, 2013): 

-	Direct service from origin to destination
-	Small vehicles that can be used exclusively by an individual or a small group
-	Service based on the needs of the user (on demand rather than fixed schedules)
-	Travelling speed up to 30 mph (48 km/h)
-	Fully automated vehicles
-	Vehicles restricted to a route reserved for their exclusive use, along the guideways
-	Small driveways that are usually elevated at or near ground level 

The components of the personal rapid transit system are:

-	**Software**: integrates all elements of the systems and ensures its flawless functioning 
-	**Hardware**: includes electric and electronic components along the rails and at the stations
-	**Vehicles**: are crucial for passenger experience. In some countries, eg. the USA, PRT vehicles must be suitable for use by disabled people
-	**Stations and guideways**: are the most important elements from the perspective of planners because they influence the urban architecture and invade significantly in the aesthetic of the urban environment. Guideways are physical structures that support and guide an AGT vehicle, consisting of rails and columns. This is the most expensive element of the PRT system
-	**Power sources**: all PRT are powered electrically
-	**Operations and maintenance facilities**: are necessary for control centre, vehicle and equipment storage and maintenance areas

PRT can be considered a more environmentally friendly, lighter and quieter alternative to [light rail transport](#lrt) (LRT) or [bus rapid transport](#brt) (BRT). It also travels slower than typical LRT or BRT, however, it compensates the time by only stopping at the requested stations. This, in turn, saves significant amount of deceleration and acceleration time. The capacity of PRL reaches 10 000 passenger per hour per direction, which is slightly lower compared to LRT and BRT. 

### Key stakeholders {-}

- **Affected**: Citizen, Airport Passengers
- **Responsible**: City government, Private Companies, Transport Planners

### Current state of art in research {-}
Recent research on PRT is scarce, where the majority of studies have been conducted around 2010. They mainly focused on the cost-benefit analysis and comparisons to LRT and BRT (Juster & Schonfeld, 2013). Furthermore, they considered the potential of different integration designs such as construction of guideways above the pavement with and without trees, above the parking lane, in the middle of the road, adjacent to building or integrated into the buildings or tunnels. Special considerations are required due to grade separation which can interfere with trees or cause obstruction at the pavements (Furman et al., 2014). An important aspect is also the aesthetic side of PRT which intrudes significantly into the urban landscape due to elevated rails. Therefore, concerns arise about social acceptability of PRT, where it is expected to affect adversely land use, property values, obstruct light and air as well as create unused spaces underneath the guideways (Staniscia, 2018). The most recent studies focus on the PRT in the context of airport shuttles where this concept has already been proven functional. For example, study by Mobolaji et al., (2021) explored the potential of extension of PRT to connect parking facilities and terminal at Budapest Airport, taking into account user expectations and system architecture.  

### Current state of art in practice {-}
The concept of PRT is more than 50 years old, nevertheless, they are only available in few places around the world. About 30% of the PRT are located at the airports, 30% in the institutional context such as university campus and the final 30% is a mass transit. The examples of existing PRTs are (Furman et al., 2014, p.12):

-	The Morgantown PRT at West Virginia University
-	The Parkshuttle Rivium metro-feeder outside Rotterdam 
-	The Terminal 5 shuttle at London Heathrow Airport 
-	The nature park shuttle in Suncheon Bay, South Korea 
-	The Masdar City PRT in Abu Dhabi 

The reasons for such a limited appearance are scare technical capabilities and scalability of PRT systems along with high capital and operating costs (Furman et al., 2014, p. 36). 


### Relevant initiatives in Austria {-}

- [17th International Conference on Personal Rapid Transit and System Design in 2023](https://waset.org/personal-rapid-transit-and-system-design-conference-in-june-2023-in-vienna) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table67, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Adverse impact on visual aspect of urban spaces|**-**|Health & Wellbeing (*3*)|Halprin, 1966; Pushkarev, 1963|
|Systemic|Grade separation increases safety|**+**|Health & Wellbeing (*3*)|Juster & Schonfeld, 2013; Furman et al., 2014|
|Systemic|Electric power used|**+**|Environmental sustainability (*7,12,13,15*)|Juster & Schonfeld, 2013|
|Systemic|Limited adoption worldwide|**-**|Innovation & Infrastructure (*9*)|Furman et al., 2014|')
```

### Technology and societal readiness level {-}

```{r table68, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-8|5-7|')
```

### Open questions {-}
1. More research is needed on actual analysis of costs and benefits associated with the construction of PRT.
2. Can PRT infrastructure be considered and used as a attractive urban furniture? How can art community be engaged in its development?
3. What are the major barriers for more widespread adoption of PRT?

### Further links {-}

- [MIT Report](https://transweb.sjsu.edu/sites/default/files/1227-automated-transit-networks.pdf)

### References {-}
- Furman, B., Fabian, L., Ellis, S., Muller, P., & Swenson, R. (2014). Automated transit networks (ATN): A review of the state of the industry and prospects for the future.
- Halprin, L. (1966). Freeways. NY.
- Juster, R., & Schonfeld, P. (2013). Comparative analysis of personal rapid transit as an urban transportation mode. Transportation research record, 2350(1), 128-135.
- Mobolaji, K., Földes, D., & Csiszár, C. (2021). Concept of Advanced Personal Rapid Transit at Airports. Periodica Polytechnica Civil Engineering, 65(1), 320-334.
- Staniscia, S. (2018). Aesthetic appreciation of Personal Rapid Transit: A new viewpoint. Cities, 79, 169-177.


## Bus rapid transit {#brt}

## Light rail transit {#lrt}

## Passenger drones

### Synonyms {-}
*urban air mobility (UAM), vertical take-off and landing (VTOL), unmanned aerial vehicles (UAVs)*

### Definition {-}
Drones or unmanned aerial vehicles (UAVs) could become the most iconic technology of the 21st century. Drones combine three key principles of technological modernity - computing, autonomy and limitless mobility. Capabilities that until now could only be used by the military are becoming accessible to most of the population. Potential use cases for drones range from surveillance and reconnaissance missions to novel forms of logistics and personal transport. The commercial use of drones is associated with enormous economic opportunities. However, even though drones are already common as surveillance/sensing devices in security services, geodesy or agriculture, their use as a means of transport is still at the beginning (Kellermann et al., 2020). <br/><br/>
Delivery drones are currently able to lift weights of up to 2-3 kg and carry out flight missions in an urban space (Kellermann et al., 2020). However, passenger drones have also already demonstrated the technical ability to transport passengers within or between cities (LeBeau, 2016; Holt, 2018; Hawkins, 2018). It is not only a historic turning point in aviation, but the beginning of a new era in which flat airspace could become the "third dimension" of transport (Kellermann et al., 2020).<br/><br/>
The name of the new type of vehicle is still far from being agreed internationally. There are several names to choose from, such as passenger drone, manned multicopter, Passenger Air Vehicle (PAV), Electric Vertical Take-off and Landing Aircraft (EVTOL), autonomous air taxis, unmanned aerial taxis, flying cars or even a new term (Pramer & Sommavilla, 2020). <br/><br/>
The autonomous air taxis will be a mixture of helicopter and drone. But this also means that they will be vertical take-off and landing (VTOL) vehicles.
There are several reasons why drone-related industries are being supported significantly. One of the reasons is that the airspace is still fairly free of traffic. The risk of collisions is (relatively) low and autopilots for aircraft have long been established. Industry experts therefore suspect that we could see self-flying air taxis even before self-driving cars. Being able to do without pilots would make an air taxi service even cheaper and make it possible for more people to afford it (UNIQA, 2019).<br/><br/>
The European Commission estimates the economic impact at €10 billion annually until 2035 and foresees the creation of more than 100,000 direct jobs. Taking into account indirect macroeconomic effects in drone-related industries, the Commission even projects 250,000 to 400,000 additional jobs (SESAR, 2016).

### Key stakeholders {-}

- **Affected**: Citizen, Insurers
- **Responsible**: National Governments, City government, Private Companies

### Current state of art in research {-}
Since all prototypes are owned by private companies and the technology is not really shared due to competition, there are few technical research papers.
The media analysis about drones for parcel and passenger transportation shows that currently the research is focused in the following areas (see table below):  

```{r table17, echo=FALSE, results='asis'}
cat('| Topics| Percentage|Number of studies|
|:------------:|:-----------:|:------------:|
|General Surveys    |   18.9%  |  21   |
|Logistics (general)  |  18.0%   |  20   |
|Attitude and Acceptance Research    | 13.5%    |  15   |
|Law and Regulations    | 11.7%    |   13  |
|Ethics and Technology Assessment    | 10.8%    |  12   |
|Sustainability Assessment    | 8.1%    |  9   |
|Urban and Transportation Planning    |  7.2%   |  8   |
|Political agenda/strategies    |  6.3%   |   7  |
|Passenger Transportation   |  2.7%   |  3   |
|Humanitarian Logistics    |2.7%     |   3  |') 
```

And although there are currently no autonomous air taxis, there is an initial research examining the factors of consumer willingness to fly in autonomous air taxis. This study identified four factors that turned out to be significantly positive: Familiarity, Value, Fun Factor, Sense of Happiness and two as significantly negative: Aversion to New Technology and Fear (Winter et al., 2020). <br/><br/>
When comparing the GHG emissions of conventional cars with VTOL passenger drones, the passenger drone actually performs a little better than the ordinary car from 35 km onwards. The reasons for this break-even point are, on the one hand, the energy-intensive take-off and landing hover mode and, on the other hand, the enormous efficiency of flying from point A to point B. Given the expected significant time savings over flying, passengers may have an incentive to share VTOL journeys. Therefore, it seems likely that the average occupancy of VTOLs will be greater than that of conventional passenger cars (Kasliwal et al., 2019).

### Current state of art in practice {-}
There are quite a few providers working to develop the unmanned aerial taxis. Munich-based start-up *Lilium* successfully launched a one-and-a-half-ton prototype vertically into the air and hovered in place in early May 2019. It sounds like a helicopter, but it doesn't look like one: The Lilium Jet has wings and 36 all-electric jet motors. That makes it quieter and more energy efficient (and makes it look more futuristic) than a helicopter. Lilium expects to be able to start commercial everyday operation not before 2025.<br/><br/>
But *Lilium* is far from the only manufacturer. *Joby Aviation, Volocopter, AeroMobil, Kittyhawk* and *Zee.Aero* are just a few of the companies hoping to place the most successful model on the market. Already established companies are also taking the concept very seriously. Mobility company *Uber*, aircraft manufacturers *Airbus* and *Boeing*, and car companies such as *Daimler* and *Porsche* are in the race (UNIQA, 2019).
Some manufacturers such as Boeing (LeBeau, 2016), Airbus (Hawkins, 2018) and Volocopter (Holt, 2018) have already conducted the first flight tests of their prototypes.<br/><br/>
The European Union proposes that support services such as flight planning, flight permits and clearances, and dynamic airspace information will be available for drone flight from 2022. From 2027, services such as collision avoidance and capacity management in congested areas will follow. From 2035, according to the European Commission's timetable, the full integration of unmanned aerial vehicles into controlled airspace with manned aviation should be completed (Wiener Zeitung, 2019).<br/><br/>
The air taxi will not be shuttling crowds of tourists between Schwechat and Stephansplatz in the next ten years. But thanks to the decreasing noise levels of the rotors, which are no longer noticeable in the "normal city noise", as Volocopter claims, they will be used more and more often in one or the other metropolis with millions of inhabitants - possibly also in Vienna (Pramer & Sommavilla, 2020).

The first approved test track in Austria for an unmanned aerial taxi is located in Linz. There is already a functioning prototype in Austria that was developed by *Ehang* in China and built by *FACC* in Innviertel. The aircraft costs around 300,000 euros, weighs 360 kilos, is equipped with 16 electric motors and 16 rotors and is designed to autonomously transport two people. The batteries of the eight-armed drone are sufficient for about 50 kilometres. In its own production line in Ried, 300 units are to be delivered by the end of next year. In order for them to be able to take off in Europe and also for Linz AG, they are working with Austro Control on the "approval regulations", explains FACC board member Robert Machtlinger. 
In Linz it was once again said that this type of passenger transport is seen as a supplement to bus and train. The air enables the fastest connection from A to B in urban areas. However, before the first test route is set up in the Upper Austrian capital for this autonomous transport, 5G mobile radio must first be installed, which was planned for spring 2020 (DerStandard, 2019).

### Relevant initiatives in Austria {-}

- [derbrutkasten.com](https://www.derbrutkasten.com/autonome-lufttaxis-linz-ag-facc-ehang/) 
- [derstandard.at-1](https://www.derstandard.at/story/2000103120464/erste-teststrecke-fuer-e-lufttaxis-2020-in-linz) 
- [derstandard.at-2](https://www.derstandard.at/story/2000122402408/flugtaxis-wann-kommt-der-tesla-der-luefte) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table18, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|It is expected to be too costly for general use|**-**|Equality (*5,10*)|Pramer & Sommavilla, 2020|
|Individual|The flight price is expected to settle in the range of a very expensive taxi|          **-**|Sustainable economic development (*8,11*)|Pramer & Sommavilla, 2020|
|Systemic|Slightly reduced GHG emissions compared to conventional cars from 35 km onwards|**~**|Environmental sustainability (*7,12,13,15*)|Kasliwal et al., 2019|
|Systemic|Increased investment until 2035 and creation of job opportunities|          **+**|Sustainable economic development (*8,11*)|SESAR, 2016|
|Systemic|Stationary areas for safety checks|          **~**|Innovation & Infrastructure (*9*)|Pramer & Sommavilla, 2020|')
```

### Technology and societal readiness level {-}

```{r table19, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-6|1-3|')
```

### Open questions {-}

1. Who will develop the regulations for passenger drones?
2. How much space is needed for take-off and landing, and will this differ between the various providers?
3. What types of routes will be replaced with flying taxis?
4. Will some companies work together and share their technologies?
5. What other areas of application will there be?
6. What will be the thematic priorities for development in the coming years?
7. What name will be agreed internationally for this type of vehicles?


### References {-}

- DerStandard (2019). ‘Teststrecke: Erstes unbemanntes Lufttaxi hebt 2020 in Linz ab - Unternehmen - derStandard.at › Wirtschaft’, 14 May. Available at: https://www.derstandard.at/story/2000103120464/erste-teststrecke-fuer-e-lufttaxis-2020-in-linz (Accessed: 21 January 2021).
- Holt, K. (2018). Volocopter will test its autonomous air taxis in Singapore next year | Engadget. Available at: https://www.engadget.com/2018-10-24-volocopter-air-taxi-test-singapore-autonomous-drone-helicopter.html (Accessed: 21 January 2021).
- J. Hawkins, A. (2018). ‘Airbus’ autonomous “air taxi” Vahana completes its first test flight - The Verge’, 1 February. Available at: https://www.theverge.com/2018/2/1/16961688/airbus-vahana-evtol-first-test-flight (Accessed: 21 January 2021).
- Kasliwal, A. et al. (2019). ‘Role of flying cars in sustainable mobility’, Nature Communications, 10(1). doi: 10.1038/s41467-019-09426-0.
- Kellermann, R., Biehle, T. and Fischer, L. (2020) ‘Drones for parcel and passenger transportation: A literature review’, Transportation Research Interdisciplinary Perspectives. The Author(s), 4, p. 100088. doi: 10.1016/j.trip.2019.100088.
- LeBeau, P. (2016). Boeing’s first test flight of air taxi a success as it works on Uber Air. Available at: https://www.cnbc.com/2019/01/23/boeing-takes-step-in-developing-uber-air--with-successful-test-flight.html (Accessed: 21 January 2021).
- Pramer, P. and Sommavilla, F. (2020) Flugtaxis: Wann kommt der Tesla der Lüfte? - PodcastEditionZukunft - derStandard.at › EditionZukunft. Available at: https://www.derstandard.at/story/2000122402408/flugtaxis-wann-kommt-der-tesla-der-luefte (Accessed: 21 January 2021).
- SESAR (2016). European Drones Outlook Study, Single European Sky ATM Research. doi: 10.2829/219851.
- UNIQA (2019). Lufttaxis: Die Überflieger im Verkehr | UNIQA Österreich | UNIQA Österreich. Available at: https://www.uniqa.at/versicherung/mobilitaet/lufttaxis.html (Accessed: 21 January 2021).
- Wiener Zeitung (2019). ‘Regelwerk für autonome Lufttaxis noch offen - Wiener Zeitung Online’, 20 September. Available at: https://www.wienerzeitung.at/nachrichten/wirtschaft/oesterreich/2030182-Fliegen-statt-fahren.html (Accessed: 21 January 2021).
- Winter, S. R., Rice, S. and Lamb, T. L. (2020). ‘A prediction model of Consumer’s willingness to fly in autonomous air taxis’, Journal of Air Transport Management. Elsevier Ltd, 89(August), p. 101926. doi: 10.1016/j.jairtraman.2020.101926.




## Automatic train operations