# Collective mobility vehicles {#collective}

## Demand responsive transit {#drt}

### Synonyms {-}
*Micro-public transport, Sustainable Urban Mobility Plans (SUMP), European Structural and Investment Funds (ESIF), Information and communications technology (ICT)*

### Definition {-}
Demand-responsive transit (DRT) systems provide transport services in response to a passenger's request. Modern telecommunication technologies are used to receive requests (via websites, SMS, mobile apps, and less frequently phone calls). These systems are a middle ground between routing systems and taxi services and hardly differ from them in extreme implementations. DRT systems offer flexibility in transport services provided according to demand by using different technologies and organisational principles (Gorev et al., 2020).
DRT has been seen primarily as a solution for people with reduced mobility, but targeting new user groups and combining it with new ICT systems and traffic management tools is an innovative and effective approach. This new form of mobility can bring many benefits to European regions, such as cost savings, less congestion on the roads and emission reductions, while improving the quality of life of citizens. DRT systems are adaptable to almost any region, with numerous parameters that can be changed to suit different regional scenarios (Hunkin & Krell, 2018):

-	DRT systems usually need to be initiated by a public authority. By their very nature, potential user groups are often scattered and fragmented, making them difficult for private operators to identify. Regions should take stock of the performance of their transport systems and consider where public transport is used and at what cost to see where it might be cheaper and more environmentally friendly to use a DRT system.
-	Sustainable Urban Mobility Plans should be developed or modified to include DRT, considering linkages with other transport modes. The process should be overseen by a single transport authority that sets clear low-carbon transport targets to show the long-term direction of travel.
-	The lead partner of DRT initiatives must bring together all stakeholders and manage the process of collaboration. It is the role of the authority to consider the long-term goals and set targets to achieve broader public policy objectives.
-	DRT needs to be made attractive and convenient if it is to have a broad impact. Communication should focus on the multiple benefits of DRT, and ICT should be used as much as possible to effectively integrate services into public transport information systems.
-	The integration of [smart cards](#contactless_cards) and electronic payments can help improve the convenience of the DRT solution, but care must be taken not to alienate users who may not have access to these technologies by avoiding app/mobile-only services.
-	Systems should aim to provide both instant and pre-order services for maximum convenience.
-	There is support for the development and implementation of DRT systems such as ESIFs (European Structural and Investment Funds).

### Key stakeholders {-}

- **Affected**: Public transport users 
- **Responsible**: National governments, Local government, Public transport authorities

### Current state of art in research {-}
There are many variables that have a direct or indirect influence on the performance of DRT services. These factors can be divided into three groups: *Network*, *Operations* and *Demand Characteristics*. Most research uses different approaches to study the impact of these factors on the performance of the service. These approaches can be broadly categorised into two groups: simulations and approximate mathematical methods (Amirgholy & Gonzales, 2016). <br/><br/>
A study on the impact of pricing and service area design on modal shift towards demand responsive transport has shown that a small service area and too low prices can lead to an undesirable shift effect from walking and cycling to DRT. In order to design a DRT concept that leads to a reduction of car trips, the service area should not be limited to the city centre but cover a typical commuter area. Furthermore, the pricing system should have a relatively high minimum charge to make DRT less attractive for short-distance travellers who can also walk or cycle. The simulation experiments have also shown that even with a city-wide DRT coverage area and a minimum charge of 3.00 EUR, without car toll only 11% of DRT users are shifted from car to DRT. With car tolls, the simulation experiments result in a larger share of users shifting from car to DRT. Similar to Kaddoura et al., (2020a) this shows the importance of combining "*pull*" measures (introducing the DRT as a new mode of transport) with "*push*" measures, such as introducing road pricing, increasing parking fees or banning private cars in certain areas. Several simulation experiments show that even with a city-wide DRT and a minimum charge, a significant proportion of DRT users come from public transport. From a user perspective, this is a positive effect as users who switch from public transport to DRT benefit from, for instance, a shorter travel time (Kaddoura et al., 2020b).
Demand responsive transport (DRT) services appear to be more effective in minimising emissions than scheduled services when demand density is not excessive and good service levels are targeted. In particular, demand responsive services perform better in a ring network and the possibility to use smaller vehicles makes them perform better than scheduled services in almost all scenarios studied (Amirgholy & Gonzales, 2016). <br/><br/>
The analysis on the introduction of flexible modes of public transport services recommends the following order of measures for the implementation of the DRT system: 

-	the analysis of the service area in terms of demand for public passenger transport services (population and job density, transport accessibility, condition of the road network, age structure of the population, income level, motorisation level); 
-	the forecast of passenger transport intensity and identification of areas with low demand; 
-	the analysis of the demand structure, identification of factors that negatively affect the DRT system, such as time-dependent demand (seasonal demand, etc.); 
-	the development of an organisational plan and DRT technology, selection of the service type; 
-	the analysis of the possibility to combine the service with social services for people with reduced mobility; 
-	the calculation of costs;
-	the determination of the possibility to receive subsidies from local authorities; 
-	the development of a detailed service plan, including a business plan and a marketing policy. 

As the DRT system needs significant subsidies from local authorities to cover the significant difference between the costs and the revenues of the transport operator, it is advisable to use combined solutions for existing routes that provide for the use of DRT at certain intervals, which can be beneficial for all entities (transport operators, local authorities, passengers) (Gorev et al., 2020).Finally, the productivity and costs of DRT services of two management practices (time window size setting and a centralised vs. decentralised strategy) is investigated by Quadrifoglio et al., (2008).

### Current state of art in practice {-}
In the last decades, several companies have launched app-based on-demand mobility services (e.g. [UBER](https://www.uber.com/at/en/), [Lyft]( https://www.lyft.com/), [BerlKönig]( https://www.berlkoenig.de/), [CleverShuttle]( https://www.clevershuttle.de/en), [MOIA]( https://www.moia.io/en)). Most of these services combine ride requests of several passengers who then share a ride (pooling, ridesharing). These services form a new category within public transport and are often referred to as Demand Responsive Transit (DRT) or ride-hailing. Most of the existing services are still based on conventional driver-controlled vehicles; only a few services are experimenting with autonomous vehicles (Harris, 2015; Hsu, 2016). A broad market introduction of (shared) autonomous vehicles is expected to reduce operating costs. Lower estimates are in the range of 0.30 to 0.38 EUR per passenger-km (Bösch et al., 2018). Lower operating costs can translate into lower user prices, which can boost demand for such innovative on-demand mobility services (Kaddoura et al., 2020).
The global Demand-Responsive Transit market can be segmented on the basis of transport service, transit service type, booking type and region. In terms of transit service, the Demand-Responsive Transit market can be segmented into bus and taxi. DRT service is generally served by buses. The demand for DRT service for various applications such as school buses, para-transit, medical rides and more is increasing. <br/><br/>
Based on booking type, the Demand-responsive Transit market can be segmented into online booking and offline booking. DRT service offers multiple booking modes such as offline and online booking. Majority of consumers use online booking as it offers higher level of flexibility and real-time information. The online segment is expected to grow rapidly.
In terms of regions, the global demand responsive transit market can be segmented into North America, Europe, Asia Pacific, and Middle East and Africa. Demand-response transit in the United Kingdom and Europe has largely catered to the needs of mobility-impaired passengers since 1997. Europe is likely to dominate the global Demand-Responsive Transit market for the next few years.<br/><br/>
The global market is in its early stages; however, the number of service providers entering the market in North America and Asia-Pacific is increasing (Demand-responsive Transit (DRT) Market Analysis Report with Investment Feasibility & Trends 2026 – FLA News, 2021)
A report by Frost & Sullivan predicts that DRT shuttles will account for 50% of the shared mobility market by 2030. They forecast the DRT market to grow from 2.8 billion US dollars in 2017 to 551.61 billion US dollars and fleet size to increase from 24,100 units to 5.8 million by 2030.<br/><br/>
DRT adoption is being driven by smart city initiatives, government and mobility policies, reduced car ownership and a shift towards multimodal and intermodal transport.
Much of this growth will take place in Europe, currently the largest market for shared mobility, and China. Asia-Pacific, where shuttles are already replacing public transport, is one of the largest DRT markets and will be responsible for over 60% of fleet size growth. The US is also experimenting with different models and it is expected to see strong growth in the medium to long term. However, in other regions such as Latin America, shuttles compete with existing shuttle buses and the potential market in Africa is still very small. 
DRT shuttles combine the affordability of public transport with the convenience of single rides. Accessible transport promotes social inclusion as citizens of all income levels are able to commute to work in a reasonable time.<br/><br/>
The ITF Forum proposes to replace all car trips with demand responsive shuttles.
As DRT shuttles generally complement public transport, Frost & Sullivan recommend public-private partnerships to expand the market and deploy resources effectively. They suggest that asset-intensive services could be operated in collaboration between different service providers or through specially developed technology platforms (Martret, 2019).
The degree of cost recovery from ticket sales of flexible public transport services varies, according to an evaluation of different mobility services in Germany, between 6% and 41% (BMVBS & BBSR, 2009). In rural areas in Austria, the degree of cost recovery is about 15% due to low utilisation rate (VCÖ, 2011). This circumstance requires increased subsidies, which cannot be provided without restrictions due to the need to save money in the public sector (Sammer & Klementschitz, 2010).

### Relevant initiatives in Austria {-}
The term micro-public transport is currently not found in any legal documents in Austria. In the „Kraftfahrliniengesetz“ (KflG), the term Anrufsammeltaxis (§ 38 para. 3 Z 2 KflG) is explained in more detail and many micro-public transport services correspond to this legal definition. In addition, the commercial operation of passenger transport requires a licence under the “Gelegenheitsverkehrsgesetz” (§ 3 para. 1 Z 3 GelverkG). Nevertheless, there are micro public transport services, which cannot be assigned to the KflG nor the GelverkG. For the future development, with regard to legal certainty, more in-depth research work is recommended (Brandl, 2020).

- [bedarfsverkehr.at](https://www.bedarfsverkehr.at/content/Literatur)
- [repositum.tuwien.at](https://repositum.tuwien.at/handle/20.500.12708/1312)
- [projekte.ffg.at](https://projekte.ffg.at/projekt/2929323)


### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table80, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Increases affordability, social inclusion and accessibility of transport.|**+**|Equality (*5,10*)|Martret, 2019|
|Systemic|Potential for cost savings, congestion and emissions reductions|**+**|Health & Wellbeing (*3*)|Hunkin and Krell, 2018|
|Systemic|Only 11% of car users are predicted to shift to DRT|**~**|Environmental sustainability (*7,12,13,15*)|Kaddoura et al., 2020b|
|Systemic|Predicted market growth worldwide; low utilisation rate in Austrain rural areas|**~**|Sustainable economic development (*8,11*))|VCOE, 2011; Martret, 2019|')
```

### Technology and societal readiness level {-}

```{r table81, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|6-8|')
```

### Open questions {-}
1. Which service concepts lead to which modal shift effect?
2. How to deal with volunteering in the context of micro-public (demand responsive) transport ? What framework conditions (legal basis) need to be created?
3. Can DRT be included in the 1-2-3 ticket in Austria, and how to operationalise it?

### Further links {-}

- [maas-alliance.eu](https://maas-alliance.eu/)

### References {-}
- Amirgholy, M., & Gonzales, E. J. (2016). Demand responsive transit systems with time-dependent demand: User equilibrium, system optimum, and management strategy. Transportation Research Part B: Methodological, 92, 234–252. https://doi.org/10.1016/j.trb.2015.11.006
-	BMVBS, & BBSR. (2009). Mobilitätskonzepte zur Sicherung der Daseinsvorsorge in nachfrageschwachen Räumen - Evaluationsreport. BBSR-Online-Publikation, 10, 47. http://www.bbsr.bund.de/BBSR/DE/Veroeffentlichungen/BBSROnline/2009/DL_ON102009.pdf?__blob=publicationFile&v=2
-	Bösch, P. M., Becker, F., Becker, H., & Axhausen, K. W. (2018). Cost-based analysis of autonomous mobility services. Transport Policy, 64(August 2017), 76–91. https://doi.org/10.1016/j.tranpol.2017.09.005
-	Brandl, H. (2020). Mobilität – Daseinsgrundfunktion in ländlichen Räumen? Erfolgsfaktoren für den idealtypischen Prozess zur Planung und Implementierung von Mikro-ÖV Angeboten.
-	Demand-responsive Transit (DRT) Market Analysis Report with Investment Feasibility & Trends 2026 – FLA News. (2021, March 12). https://www.flanewsonline.com/demand-responsive-transit-drt-market-analysis-report-with-investment-feasibility-trends-2026/
-	Gorev, A., Popova, O., & Solodkij, A. (2020). Demand-responsive transit systems in areas with low transport demand of “smart city”. Transportation Research Procedia, 50(2019), 160–166. https://doi.org/10.1016/j.trpro.2020.10.020
-	Harris, M. (2015, September 14). Uber Could Be First to Test Completely Driverless Cars in Public - IEEE Spectrum. https://spectrum.ieee.org/cars-that-think/transportation/self-driving/uber-could-be-first-to-test-completely-driverless-cars-in-public
-	Hsu, J. (2016, January 4). GM and Lyft Team Up for Robot Taxi Service - IEEE Spectrum. https://spectrum.ieee.org/cars-that-think/transportation/self-driving/gm-and-lyft-team-up-for-robot-taxi-service
-	Hunkin, S., & Krell, K. (2018). Demand responsive transportation. Interreg Europe Policy Learning Platform on Low-Carbon Economy. https://doi.org/10.4324/9781315118321
-	Kaddoura, I., Bischoff, J., & Nagel, K. (2020a). Towards welfare optimal operation of innovative mobility concepts: External cost pricing in a world of shared autonomous vehicles. Transportation Research Part A: Policy and Practice, 136(February), 48–63. https://doi.org/10.1016/j.tra.2020.03.032
-	Kaddoura, I., Leich, G., & Nagel, K. (2020b). The impact of pricing and service area design on the modal shift towards demand responsive transit. Procedia Computer Science, 170, 807–812. https://doi.org/10.1016/j.procs.2020.03.152
-	Martret, O. (2019, March 25). Demand-Responsive Transit Is the Future of the Mobility Market. https://shotl.com/news/responsive-transit-is-the-future-of-the-mobility-market
-	Quadrifoglio, L., Dessouky, M. M., & Ordóñez, F. (2008). A simulation study of demand responsive transit system design. Transportation Research Part A: Policy and Practice, 42(4), 718–737. https://doi.org/10.1016/j.tra.2008.01.018
-	Sammer, G., & Klementschitz, R. (2010). Finanzierung des öffentlichen Personennahverkehrs in Österreich. ÖGZ - Österreichische Gemeindezeitung, 26 – 28.
-	VCÖ. (2011). Erfolgreicher Öffentlicher Verkehr. In VCÖ Schriftenreihe ‘Mobilität mit Zukunft’ (Issue 4).

## Personal rapid transit {#prt}

### Synonyms {-}
*Automated Transit Networks, ATN, PRT, Podcars*

### Definition {-}
Personal rapid transit is a public transportation option in urban areas where fully automated vehicles provide on-demand, origin-to-destination service over an area network. PRT belongs to a larger family of automated guideway transit (AGT). It is a unique mode of transport that has the following features (Furman et al., 2014; Juster & Schonfeld, 2013): 

-	Direct service from origin to destination
-	Small vehicles that can be used exclusively by an individual or a small group
-	Service based on the needs of the user (on demand rather than fixed schedules)
-	Travelling speed up to 30 mph (48 km/h)
-	Fully automated vehicles
-	Vehicles restricted to a route reserved for their exclusive use, along the guideways
-	Small driveways that are usually elevated at or near ground level 

The components of the personal rapid transit system are:

-	**Software**: integrates all elements of the systems and ensures its flawless functioning 
-	**Hardware**: includes electric and electronic components along the rails and at the stations
-	**Vehicles**: are crucial for passenger experience. In some countries, eg. the USA, PRT vehicles must be suitable for use by disabled people
-	**Stations and guideways**: are the most important elements from the perspective of planners because they influence the urban architecture and invade significantly in the aesthetic of the urban environment. Guideways are physical structures that support and guide an AGT vehicle, consisting of rails and columns. This is the most expensive element of the PRT system
-	**Power sources**: all PRT are powered electrically
-	**Operations and maintenance facilities**: are necessary for control centre, vehicle and equipment storage and maintenance areas

PRT can be considered a more environmentally friendly, lighter and quieter alternative to [light rail transport](#lrt) (LRT) or [bus rapid transport](#brt) (BRT). It also travels slower than typical LRT or BRT, however, it compensates the time by only stopping at the requested stations. This, in turn, saves significant amount of deceleration and acceleration time. The capacity of PRL reaches 10 000 passenger per hour per direction, which is slightly lower compared to LRT and BRT. 

### Key stakeholders {-}

- **Affected**: Citizen, Airport Passengers
- **Responsible**: City government, Private Companies, Transport Planners

### Current state of art in research {-}
Recent research on PRT is scarce, where the majority of studies have been conducted around 2010. They mainly focused on the cost-benefit analysis and comparisons to LRT and BRT (Juster & Schonfeld, 2013). Furthermore, they considered the potential of different integration designs such as construction of guideways above the pavement with and without trees, above the parking lane, in the middle of the road, adjacent to building or integrated into the buildings or tunnels. Special considerations are required due to grade separation which can interfere with trees or cause obstruction at the pavements (Furman et al., 2014). An important aspect is also the aesthetic side of PRT which intrudes significantly into the urban landscape due to elevated rails. Therefore, concerns arise about social acceptability of PRT, where it is expected to affect adversely land use, property values, obstruct light and air as well as create unused spaces underneath the guideways (Staniscia, 2018). The most recent studies focus on the PRT in the context of airport shuttles where this concept has already been proven functional. For example, study by Mobolaji et al., (2021) explored the potential of extension of PRT to connect parking facilities and terminal at Budapest Airport, taking into account user expectations and system architecture.  

### Current state of art in practice {-}
The concept of PRT is more than 50 years old, nevertheless, they are only available in few places around the world. About 30% of the PRT are located at the airports, 30% in the institutional context such as university campus and the final 30% is a mass transit. The examples of existing PRTs are (Furman et al., 2014, p.12):

-	The Morgantown PRT at West Virginia University
-	The Parkshuttle Rivium metro-feeder outside Rotterdam 
-	The Terminal 5 shuttle at London Heathrow Airport 
-	The nature park shuttle in Suncheon Bay, South Korea 
-	The Masdar City PRT in Abu Dhabi 

The reasons for such a limited appearance are scare technical capabilities and scalability of PRT systems along with high capital and operating costs (Furman et al., 2014, p. 36). 


### Relevant initiatives in Austria {-}

- [17th International Conference on Personal Rapid Transit and System Design in 2023](https://waset.org/personal-rapid-transit-and-system-design-conference-in-june-2023-in-vienna) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table67, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Adverse impact on visual aspect of urban spaces|**-**|Health & Wellbeing (*3*)|Staniscia, 2018|
|Systemic|Grade separation increases safety|**+**|Health & Wellbeing (*3*)|Juster & Schonfeld, 2013; Furman et al., 2014|
|Systemic|Electric power used|**+**|Environmental sustainability (*7,12,13,15*)|Juster & Schonfeld, 2013|
|Systemic|Limited adoption worldwide|**-**|Innovation & Infrastructure (*9*)|Furman et al., 2014|')
```

### Technology and societal readiness level {-}

```{r table68, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-8|5-7|')
```

### Open questions {-}
1. More research is needed on actual analysis of costs and benefits associated with the construction of PRT.
2. Can PRT infrastructure be considered and used as a attractive urban furniture? How can art community be engaged in its development?
3. What are the major barriers for more widespread adoption of PRT?

### Further links {-}

- [MIT Report](https://transweb.sjsu.edu/sites/default/files/1227-automated-transit-networks.pdf)

### References {-}
- Furman, B., Fabian, L., Ellis, S., Muller, P., & Swenson, R. (2014). Automated transit networks (ATN): A review of the state of the industry and prospects for the future.
- Juster, R., & Schonfeld, P. (2013). Comparative analysis of personal rapid transit as an urban transportation mode. Transportation research record, 2350(1), 128-135.
- Mobolaji, K., Földes, D., & Csiszár, C. (2021). Concept of Advanced Personal Rapid Transit at Airports. Periodica Polytechnica Civil Engineering, 65(1), 320-334.
- Staniscia, S. (2018). Aesthetic appreciation of Personal Rapid Transit: A new viewpoint. Cities, 79, 169-177.


## Bus rapid transit {#brt}

## Light rail transit {#lrt}

### Synonyms {-}
*Light Rail Transit (LRT), Heavy Rail Transit (HRT)*

### Definition {-}
Light rail (also referred to as tram) has become a popular measure among urban planners to address the social and economic complexities associated with urban growth (Baker & Lee, 2019). Light rail developments go beyond improving access to public transport to surrounding neighbourhoods. Impacts can also address property values, neighbourhood demographics, employment opportunities and access to services (Hess, 2020). <br/>

Proponents argue that the establishment of new LRT lines can promote environmental sustainability by reducing private car use and improve public health by increasing neighbourhood walkability and air quality (D. Knowles & Ferbrache, 2016). Other arguments are that LRTs address income inequality positively by encouraging business development and providing direct access to employment centres in a city or suburb (Hess, 2020). <br/>

Many of these impacts have implications for the health and well-being of residents. These are assessed through the social determinants of health (SDOH). The World Health Organization (WHO) defines the social determinants of health (SDOH) as the "*conditions under which people are born, grow up, live, work and age*" (e.g. income, housing, employment) and "*the underlying drivers of these conditions*" (e.g. economic and social policies and political systems) (Braveman & Gottlieb, 2014). The SDOH framework asserts that the way political, social and economic resources are distributed within communities can have an impact on health and well-being (Solar & Irwin, 2010). <br/>

LRT has become a popular form of transit due to low construction costs compared to underground systems and large perceived economic benefits. LRT systems are typically built along existing streets, eliminating the need for expensive tunnel or elevated rail infrastructure. While LRT shares road space with vehicles and pedestrians, portions of the routes are used with priority, allowing for higher speeds and fewer delays than buses. Unlike bus transit, the need for rails, an above-ground power source and platforms ensures that LRT is a long-term local investment (Tyndall, 2021). <br/>

Laird (2019) defines the differences between light rail systems and commuter rail systems. They function differently, have different requirements and use different systems for operation:

**Commuter rail systems**
Commuter rail systems are passenger trains that run on diesel-electric or electrically powered engines. They operate on existing railway tracks on the same routes used by intercity goods trains. They are operated by state authorities or private companies on their own or third-party tracks. They typically have speeds of 80 to 130 km/h, shorter routes and are usually only found in larger conurbations. Commuter rail systems have fewer stops than light rail and typically run through suburbs and central cities. Because commuter rail systems are designed for commuters, the schedule frequency may be different than light rail and run less frequently throughout the day. Commuter rail systems often run most frequently during standard commuting hours for the average workday. 

**Light rail systems**
Light rail systems are passenger trains powered by overhead electric wires. They have lighter frames and smaller bodies than other trains due to the turning radius required. Because light rail systems operate in city streets and urban corridors with frequent stops, they have a larger turning radius to weave in and out of busy areas and can accelerate and decelerate faster than commuter rail. While commuter rail systems can run over existing goods train tracks, light rail systems generally require their own tracks.

### Key stakeholders {-}

- **Affected**: Pedestrians, Public transport users, Local residents, Drivers
- **Responsible**: City governments, Private Transport Companies, Transport Authorities

### Current state of art in research {-}
Research mainly examines the social impacts (Baker & Kim, 2020; Baker & Lee, 2019; Deyas & Woldeamanuel, 2020; Hess, 2020; Tyndall, 2021) and the economic impacts (D. Knowles & Ferbrache, 2016) of light rail transit. <br/>

Knowles & Ferbrache (2016) assess the wider economic impacts of light rail investment on cities. They focused on evidence from UK, European and North American case studies, as there is still a lack of light rail research from less developed cities around the world. They found that investments in light rail systems can have positive economic impacts on cities. However, similar light rail investments in different locations do not necessarily have the same impact, therefore the geography matters. Similar to other forms of transport infrastructure, light rail investments alone are unlikely to be a sufficient catalyst for economic change without additional supporting measures. <br/>

Light rail can spur economic growth by improving accessibility to previously inaccessible areas, encouraging foreign investment, triggering new growth, expanding labour market catchment areas and influencing property prices. However, while many areas experience price increases following the construction of light rail, these are often not absorbed and the capital to fund the infrastructure has to be raised elsewhere. Therefore, as with other forms of public transport, the economic impact of light rail is enhanced when land use and transport planning policies are coordinated, and this is highly dependent on other contextual factors. <br/>

Tyndall (2021) notes that LRT effectively increases overall public transport use because it attracts higher-skilled workers who would be unlikely to use other forms of public transport, while low-skilled workers continue to use public transport. <br/>

**Automation of LRT**

-  Level 2 automation refers to a system where trains run automatically from station to station, but there is a driver in the cab who is responsible for closing the doors, detecting obstacles on the track in front of the train and handling emergency situations. 
- In automation level 3, trains run automatically from station to station, but there is always a staff member on the train who is responsible for handling emergency situations. 
- In automation level 4, trains are able to run automatically at all times, including door closure, obstacle detection and emergency situations. On-board staff may be provided for other purposes, such as customer service, but are not required for safe operation. Controls are often provided to manually operate the train in the event of a computer failure. Examples so far are only Metros in Paris, Barcelona, Sydney and Copenhagen (UITP, n.d.). Trams have only been tested for autonomous driving since 2018.

### Current state of art in practice {-}

Trams and light rail systems exist in 389 cities worldwide, more than half (204) of them in Europe. Between 2015 and 2018, the light rail infrastructure in Europe grew by 3.9% from 8943 km to 9296 km, and passenger numbers increased by 6.9% from 9740 million to 10,422 million passengers. Light rail now carries as many passengers as metros and regional/suburban rail lines and ten times more passengers than air travel in Europe. 
The average light rail journey in Europe is 3.27 km. The busiest light rail network in Europe is in Budapest, Hungary, with 411 million passengers, while Berlin has the longest light rail network in Europe with 193 km. Passenger growth varies from region to region, ranging from 17.5% in the British Isles to 1.5% in Poland. <br/>

While the average European line is 7.3 km long, they tend to be longer on average in countries with newer systems and a limited number of lines, while older, more complex systems have a lower average line length. Europe's fleet consists of 20,750 trams and light rail vehicles, with 51% of this fleet consisting of partial or full low-floor vehicles, ranging from countries with almost 100% such as France, Spain, Ireland, the UK and Norway to countries with much lower proportions. The average annual mileage per vehicle in Europe is 52,000 km, ranging from 38,700 km to 77,500 km (this value is theoretical and based on the assumption that all vehicles are used equally). According to UITP, light rail will continue to receive support from decision-makers and the travelling public in Europe, given the ongoing pressure to reduce congestion, tackle poor air quality in cities and reduce greenhouse gas emissions contributing to climate change.However, much attention and resources will go into maintaining, upgrading and replacing facilities to keep ageing systems attractive and operational. For this reason, the growth of greenfield projects in Europe will continue to slow down (Burroughs, 2020; UITP, 2019)

**Examples of recent investments**

- The Jerusalem tram project which involves building 27 km of new tram lines, 53 stations and a number of depots (Railway-News, 2019).

- Lisbon public transport company, and CAF signed a contract for a new LRT line. The low-floor tram will have a length of 28.5 metres and will be able to run at a speed of 70 km/h (Railwaypro, 2021b).

-	Paris opens the new T9 LRT line between Paris and the city of Orly. The 40 km line has 19 stations and is expected to carry 70,000 to 80,000 passengers per day. The light rail vehicles have eight double doors per side and wider gangways to improve passenger flow and reduce boarding times (Burroughs, 2021).
European cities will continue to invest in cleaner transport and logistics, including a boost for rail travel and clean mobility in cities and regions (European Commission, 2020). Also the World Bank, 2020 states that investing in reliable mass transit systems such as metro, LRT and Bus Rapid Transit can help keep cities moving while reducing the carbon footprint of urban transport. <br/>

- US cities also have made significant investments in Light Rail Transit (LRT) in recent years. A common justification for LRT is that transit infrastructure improves urban commuter networks by creating spatial connections between workers and jobs. Between 2000 and 2015, the number of LRT stations in four metropolitan regions in the US grew by 56% (Tyndall, 2021). <br/>

- Moscow announced that the tram network will be managed by the Moscow Metro. It is expected that the centralised management and consistent modernisation will increase tram speeds, improve track maintenance, halve the number of repairs and reduce maintenance costs. In 2019, the tram network carried 212 million passengers, which is 12 times more than the population of Moscow. By 2023, Moscow plans to replace its entire existing old tram fleet with low-floor trams. It is expected that by 2024 only low-floor trams will operate on all Moscow routes (Railwaypro, 2021a). <br/>

Furthermore, the investments are also being made in tramtrains. A combination of Light rail and train. While as a train it has the travel advantages of a railway in the surrounding areas, such as speeds, safety standards, ride comfort, sanitary facilities and in the city centre it functions as a tram. These multi-system vehicles, with their equipment and operating characteristics, are ideally suited to the needs of passengers who travel longer distances by train, as well as for passengers who make only a few stops in the city centre. Above all, the tramtrain vehicles also enable direct connections from surrounding areas to the city without changing trains. From the region to the city limits, passengers, thus, have the advantage of fast travel times and high passenger comfort (Seyringer, 2020). In Karlsruhe there are some examples of tram trains existing since 1992 (Stadtwiki Karlsruhe, 2016). They already have continuous rail connections between the inner cities and the regions (Stadtwiki Karlsruhe, 2020). Further, the Badner Bahn is also a well-known tram-train example. The 30.4 km long Badner Bahn tram system runs between Vienna and Baden. Every year, Wiener Lokalbahnen transports around 12 million passengers between Vienna and Baden. Commuters from the south of Vienna are particularly prone to use the Badner Bahn every day (RailwayPro, 2020). <br/>

**Automation**

For several decades now, the tram has been experiencing a renaissance worldwide. Cars and buses are becoming increasingly intelligent and independent thanks to advanced sensor and automation technology. If the tram wants to keep up and secure its attractiveness and competitiveness in the long term, it must evolve into a smart and autonomous means of transport. The automation technology works by using lidar, radar and cameras (Siemens Mobility, 2019):

*Lidar (Light detection and ranging)*

-	Enables 3D environmental detection and positioning
-	Scans objects vertically and horizontally with laser beams and horizontally; uses the reflected waves to perceive the surroundings
-	Enables the tram to "see" at an angle of up to 270&deg;

*Radar (Radio Detection and Ranging)*

-	Measures distance and speed with high accuracy - especially of metallic objects
-	Emits radio waves and uses the reflected waves to locate objects
-	Covers a large area in front of the tram 

*Cameras*

-	Are trained in intelligent object and signal recognition
-	Can detect and classify objects in thousands of contours and positions - e.g. as people, signals or infrastructure elements
-	Cover a large optical area around the vehicle

A few years before 2020, Siemens Mobility launched the driver assistance system "*Siemens Tram Assistant*" - a collision warning and protection system to support the driver. The system is already being used successfully in Siemens trams in Den Haag (Netherlands) and Ulm (Germany). The next commercially viable step planned is depot automation based on an autonomous tram. This will make it possible to automate time-consuming shunting operations in the depot, such as service runs through a car wash to a siding. The extensive isolation from public traffic simplifies technical control and approval (Hofmann, 2020; Zasiadko, 2019b).<br/>

The three leading countries in the development of autonomous trams are Germany, Russia and China (Intelligent Transport, 2019). Siemens Mobility is testing the first autonomous tram in an automated depot in Germany since 2018 (Hofmann, 2020). PC Transport Systems and Cognitive Technologies have announced a joint project that will develop a fully autonomous tram for the Russian and foreign markets by 2021-2022 (Intelligent Transport, 2019; Zasiadko, 2019a). Following the successful tests of the autonomous tram in Moscow, the Russian company Cognitive Technologies is planning to develop an AI-based computer vision system for the Chinese market together with Fuxin Intelligent Transportation Solutions (FITSCO), one of the largest suppliers of signalling and communication solutions in China. This is needed for the testing and introduction of the self-driving trams in Shanghai. The project is to be implemented in 2021 (Zasiadko, 2020).<br/>

Unlike similar solutions for self-driving cars, the system for rail has a number of simplifications that will allow trams to run on public roads sooner. But despite these simplifications, there are only a few countries in the world that are ready for the real deployment of autonomous trams. The developed autonomous driving system recognises vehicles and other trams, traffic lights, pedestrians, tram and bus stops, railway switches and various obstacles. The tram is also able to stop in front of the obstacles, maintain a safe distance from the vehicles ahead, accelerate and stop (Intelligent Transport, 2019).

### Relevant initiatives in Austria {-}
Vienna has a number of expansion plans for the tram network planned for the coming decade, including connecting the Lower Austrian surrounding area with three lines. Such routes are now becoming more realistic. This is because both provinces are examining the possibilities and framework conditions for implementation in spring 2020 (Wiener Linien, 2020).

- [Tram network extension plans](https://wien.orf.at/stories/3068577/)

"*In addition to the 1-2-3-Klimaticket itself, Upper Austrians will also get an improved, innovative public transport infrastructure with the Linz-Gallneukirchen-Pregarten regional light rail system in the coming years and thus further incentives for commuters to travel to Linz by 'tram-train' instead of standing in traffic jams on the S10 and A7*" explains Weratschnig (Transport spokesman of the Green Party) (Grüner Klub im Parlament, 2021).Further, Badner Bahn will receive new trams. These will gradually replace the old high-floor models starting in the second half of 2021 (RailwayPro, 2020).

- [Badner Bahn](https://www.railwaypro.com/wp/interior-design-for-wiener-lokalbahnen-new-tram-unveiled/)

TramTrains for Linz (Upper Austria) are expected in July 2026. The vehicles will have regional train seats, seating groups and luggage storage facilities in the middle section, while multi-purpose compartments for wheelchair, pram and bicycle transport are planned for the front and rear sections. The vehicles intended for use on the longer light rail routes will be equipped with toilets (Seyringer, 2020).In Salzburg, too, at least 20 new tram-trains are planned for Salzburg until 2026 (SALZBURG24, 2020). 

- [Salzburg tram](https://www.salzburg24.at/news/salzburg/salzburger-lokalbahn-bekommt-neue-tram-trains-91192069)
- [Linz tramtrain](https://www.ots.at/presseaussendung/OTS_20210312_OTS0128/grueneweratschnig-vom-muehlkreis-aus-bald-per-tram-train-nach-linz-statt-im-stau-auf-der-a7)
- [Tramtrain in Austria](https://kommunal.at/eisenbahn-und-stadtbahn-einem)

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table104, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Increase in neighbourhood walkability and air quality|**+**|Health & Wellbeing  (*3*)|D. Knowles & Ferbrache, 2016|
|Systemic|Encouraged business development and provision of direct access to employment centres |**+**|Equality  (*5,10*)|Hess, 2020|
|Systemic|Reduced car use|**+**|Environmental sustainability (*7,12,13,15*)|D. Knowles & Ferbrache, 2016|
|Systemic|Low construction costs compared to underground systems and large perceived economic benefits|**+**|Sustainable economic development  (*8,11*)|Tyndall, 2021|
|Systemic|Support from decision-makers in Europe but still a lack of significant LRT efforts in less developed cities around the world| **~**|Partnership & collaborations (*17*)|Burroughs, 2020; D. Knowles & Ferbrache, 2016; UITP, 2019|')
```

### Technology and societal readiness level {-}

```{r table105, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|8-9|6-9|')
```

### Open questions {-}
1. What is the social impact of light rail transit?
2. What is the potential of LRT to influence residential choices?

### Further links

- [railjournal.com](https://www.railjournal.com/category/passenger/light-rail/)
- [uitp.org](https://www.uitp.org/topics/light-rail/)
- [railwaypro.com](https://www.railwaypro.com/wp/)
- [railway-news.com](https://railway-news.com/)
- [co2-emmisions](https://ourworldindata.org/co2-emissions-from-transport)
- [lrta.org](http://www.lrta.org/) 
- [mobility.siemens.com](https://www.mobility.siemens.com/global/en/portfolio/rail/rolling-stock/trams-and-light-rail/autonomous-tram.html)
- [metroautomation.org](https://metroautomation.org/what-makes-automation-possible/)


### References {-}

- Baker, D. M., & Kim, S. (2020). What remains? The influence of light rail transit on discretionary income. Journal of Transport Geography, 85, 102709. https://doi.org/10.1016/j.jtrangeo.2020.102709
-	Baker, D. M., & Lee, B. (2019). How Does Light Rail Transit (LRT) Impact Gentrification? Evidence from Fourteen US Urbanized Areas. Journal of Planning Education and Research, 39(1), 35–49. https://doi.org/10.1177/0739456X17713619
-	Braveman, P., & Gottlieb, L. (2014). The social determinants of health: It’s time to consider the causes of the causes. Public Health Reports, 129(SUPPL. 2), 19–31. https://doi.org/10.1177/00333549141291s206
-	Burroughs, D. (2020, January 16). Light rail growth strong in Europe, UITP says | International Railway Journal. Railjournal. https://www.railjournal.com/passenger/light-rail/light-rail-sees-strong-growth-in-europe-uitp-says/
-	Burroughs, D. (2021, April 24). Paris T9 light rail line opens | International Railway Journal. https://www.railjournal.com/passenger/light-rail/paris-t9-light-rail-line-opens/
-	D. Knowles, R., & Ferbrache, F. (2016). Evaluation of wider economic impacts of light rail investment on cities. Journal of Transport Geography, 54, 430–439. https://doi.org/10.1016/j.jtrangeo.2015.09.002
-	Deyas, G. T., & Woldeamanuel, M. G. (2020). Social and economic impacts of public transportation on adjacent communities: The case of the Addis Ababa light rail transit. Research in Transportation Economics, 84, 100970. https://doi.org/10.1016/j.retrec.2020.100970
-	European Commission. (2020, May 26). Europe’s moment: Repair and prepare for the next generation. https://ec.europa.eu/commission/presscorner/detail/en/ip_20_940
-	Grüner Klub im Parlament. (2021, March 12). Grüne/Weratschnig: Vom Mühlkreis aus bald per Tram-Train nach Linz, statt im Stau auf der A7 | Grüner Klub im Parlament, 12.03.2021. https://www.ots.at/presseaussendung/OTS_20210312_OTS0128/grueneweratschnig-vom-muehlkreis-aus-bald-per-tram-train-nach-linz-statt-im-stau-auf-der-a7
-	Hess, C. L. (2020). Light-Rail Investment in Seattle: Gentrification Pressures and Trends in Neighborhood Ethnoracial Composition. In Urban Affairs Review (Vol. 56, Issue 1). https://doi.org/10.1177/1078087418758959
-	Hofmann, M. (2020). Siemens Mobility und VIP Potsdam auf dem Weg zur autonomen Tram. „bahn Manager Magazin“, Edition 02/2020.
-	Intelligent Transport. (2019, February 13). Russia’s first autonomous tram will be launched in Moscow. https://www.intelligenttransport.com/transport-news/75915/autonomous-tram-development-russia/
-	Laird, K. (2019, August 28). Spot the Difference: Commuter vs. Light Rail. https://hoponboardblog.com/2019/08/spot-the-difference-commuter-vs-light-rail/
-	Railway-News. (2019, August 9). CAF Group Consortium Wins Jerusalem Tram Contract | Railway-News. https://railway-news.com/caf-group-consortium-wins-jerusalem-tram-contract/
-	Railwaypro. (2021a, April 8). Additional Vityaz-M trams expected to enter operation in Moscow. https://www.railwaypro.com/wp/moscow-to-receive-100-low-floor-trams/
-	Railwaypro. (2021b, April 22). Urbos tram contract signed by Lisbon transport company. https://www.railwaypro.com/wp/lisbon-signs-tram-contract/
-	RailwayPro. (2020, May 20). The new trams for Badner Bahn have a comfort-oriented design. https://www.railwaypro.com/wp/interior-design-for-wiener-lokalbahnen-new-tram-unveiled/
-	SALZBURG24. (2020, August 7). Salzburger Lokalbahn bekommt neue Tram-Trains - SALZBURG24. https://www.salzburg24.at/news/salzburg/salzburger-lokalbahn-bekommt-neue-tram-trains-91192069
-	Seyringer, K. (2020, August 12). Das ist ein nie dagewesenes Projekt, von dem wir sehr profitieren werden. https://www.tips.at/nachrichten/linz/wirtschaft-politik/513716-das-ist-ein-nie-dagewesenes-projekt-von-dem-wir-sehr-profitieren-werden
-	Siemens Mobility. (2019). Teaching trams to drive . Der Trend : Städte fahren autonom.
-	Solar, O., & Irwin, A. (2010). A conceptual framework for action on the social determinants of health. 79. http://www.who.int/sdhconference/resources/ConceptualframeworkforactiononSDH_eng.pdf
-	Stadtwiki Karlsruhe. (2016). Karlsruher Modell – Stadtwiki Karlsruhe. https://ka.stadtwiki.net/Karlsruher_Modell
-	Stadtwiki Karlsruhe. (2020, December 21). Stadtbahn – Stadtwiki Karlsruhe. https://ka.stadtwiki.net/Stadtbahn
-	The World Bank. (2020, April 23). Earth Day 2020: Could COVID-19 Be the Tipping Point for Transport Emissions? https://www.worldbank.org/en/news/feature/2020/04/22/earth-day-2020-could-covid-19-be-the-tipping-point-for-transport-emissions
-	Tyndall, J. (2021). The local labour market effects of light rail transit. Journal of Urban Economics, 124, 103350. https://doi.org/10.1016/j.jue.2021.103350
-	UITP. (n.d.). Automation Essentials | Automated Metros Observatory. Retrieved May 3, 2021, from https://metroautomation.org/automation-essentials/
-	UITP. (2019). Light Rail and Tram : the European Outlook. November. https://www.uitp.org/publications/light-rail-and-tram-the-european-outlook/
-	Wiener Linien. (2020, September 25). Pläne für Straßenbahn nach Niederösterreich - wien.ORF.at. https://wien.orf.at/stories/3068577/
-	Zasiadko, M. (2019a, February 13). Russia tests its first self-driving tram | RailTech.com. RailTech.Com. https://www.railtech.com/digitalisation/2019/02/13/russia-tests-its-first-self-driving-tram/
-	Zasiadko, M. (2019b, October 9). Germany will test first autonomous tram in automated depot | RailTech.com. RailTech.Com. https://www.railtech.com/rolling-stock/2019/10/09/germany-will-test-first-autonomous-tram-in-automated-depot/
-	Zasiadko, M. (2020, April 23). Computer vision for self-driving trams in Shanghai | RailTech.com. RailTech.Com. https://www.railtech.com/digitalisation/2020/04/23/computer-vision-for-self-driving-trams-in-shanghai/





## Passenger drones {#passenger_drones}

### Synonyms {-}
*urban air mobility (UAM), vertical take-off and landing (VTOL), unmanned aerial vehicles (UAVs)*

### Definition {-}
Drones or unmanned aerial vehicles (UAVs) could become the most iconic technology of the 21st century. Drones combine three key principles of technological modernity - computing, autonomy and limitless mobility. Capabilities that until now could only be used by the military are becoming accessible to most of the population. Potential use cases for drones range from surveillance and reconnaissance missions to novel forms of logistics and personal transport. The commercial use of drones is associated with enormous economic opportunities. However, even though drones are already common as surveillance/sensing devices in security services, geodesy or agriculture, their use as a means of transport is still at the beginning (Kellermann et al., 2020). <br/><br/>
Delivery drones are currently able to lift weights of up to 2-3 kg and carry out flight missions in an urban space (Kellermann et al., 2020). However, passenger drones have also already demonstrated the technical ability to transport passengers within or between cities (LeBeau, 2016; Holt, 2018; Hawkins, 2018). It is not only a historic turning point in aviation, but the beginning of a new era in which flat airspace could become the "third dimension" of transport (Kellermann et al., 2020).<br/><br/>
The name of the new type of vehicle is still far from being agreed internationally. There are several names to choose from, such as passenger drone, manned multicopter, Passenger Air Vehicle (PAV), Electric Vertical Take-off and Landing Aircraft (EVTOL), autonomous air taxis, unmanned aerial taxis, flying cars or even a new term (Pramer & Sommavilla, 2020). <br/><br/>
The autonomous air taxis will be a mixture of helicopter and drone. But this also means that they will be vertical take-off and landing (VTOL) vehicles.
There are several reasons why drone-related industries are being supported significantly. One of the reasons is that the airspace is still fairly free of traffic. The risk of collisions is (relatively) low and autopilots for aircraft have long been established. Industry experts therefore suspect that we could see self-flying air taxis even before self-driving cars. Being able to do without pilots would make an air taxi service even cheaper and make it possible for more people to afford it (UNIQA, 2019).<br/><br/>
The European Commission estimates the economic impact at €10 billion annually until 2035 and foresees the creation of more than 100,000 direct jobs. Taking into account indirect macroeconomic effects in drone-related industries, the Commission even projects 250,000 to 400,000 additional jobs (SESAR, 2016).

### Key stakeholders {-}

- **Affected**: Citizen, Insurers
- **Responsible**: National Governments, City government, Private Companies

### Current state of art in research {-}
Since all prototypes are owned by private companies and the technology is not really shared due to competition, there are few technical research papers.
The media analysis about drones for parcel and passenger transportation shows that currently the research is focused in the following areas (see table below):  

```{r table17, echo=FALSE, results='asis'}
cat('| Topics| Percentage|Number of studies|
|:------------:|:-----------:|:------------:|
|General Surveys    |   18.9%  |  21   |
|Logistics (general)  |  18.0%   |  20   |
|Attitude and Acceptance Research    | 13.5%    |  15   |
|Law and Regulations    | 11.7%    |   13  |
|Ethics and Technology Assessment    | 10.8%    |  12   |
|Sustainability Assessment    | 8.1%    |  9   |
|Urban and Transportation Planning    |  7.2%   |  8   |
|Political agenda/strategies    |  6.3%   |   7  |
|Passenger Transportation   |  2.7%   |  3   |
|Humanitarian Logistics    |2.7%     |   3  |') 
```

And although there are currently no autonomous air taxis, there is an initial research examining the factors of consumer willingness to fly in autonomous air taxis. This study identified four factors that turned out to be significantly positive: Familiarity, Value, Fun Factor, Sense of Happiness and two as significantly negative: Aversion to New Technology and Fear (Winter et al., 2020). <br/><br/>
When comparing the GHG emissions of conventional cars with VTOL passenger drones, the passenger drone actually performs a little better than the ordinary car from 35 km onwards. The reasons for this break-even point are, on the one hand, the energy-intensive take-off and landing hover mode and, on the other hand, the enormous efficiency of flying from point A to point B. Given the expected significant time savings over flying, passengers may have an incentive to share VTOL journeys. Therefore, it seems likely that the average occupancy of VTOLs will be greater than that of conventional passenger cars (Kasliwal et al., 2019).

### Current state of art in practice {-}
There are quite a few providers working to develop the unmanned aerial taxis. Munich-based start-up *Lilium* successfully launched a one-and-a-half-ton prototype vertically into the air and hovered in place in early May 2019. It sounds like a helicopter, but it doesn't look like one: The Lilium Jet has wings and 36 all-electric jet motors. That makes it quieter and more energy efficient (and makes it look more futuristic) than a helicopter. Lilium expects to be able to start commercial everyday operation not before 2025.<br/><br/>
But *Lilium* is far from the only manufacturer. *Joby Aviation, Volocopter, AeroMobil, Kittyhawk* and *Zee.Aero* are just a few of the companies hoping to place the most successful model on the market. Already established companies are also taking the concept very seriously. Mobility company *Uber*, aircraft manufacturers *Airbus* and *Boeing*, and car companies such as *Daimler* and *Porsche* are in the race (UNIQA, 2019).
Some manufacturers such as Boeing (LeBeau, 2016), Airbus (Hawkins, 2018) and Volocopter (Holt, 2018) have already conducted the first flight tests of their prototypes.<br/><br/>
The European Union proposes that support services such as flight planning, flight permits and clearances, and dynamic airspace information will be available for drone flight from 2022. From 2027, services such as collision avoidance and capacity management in congested areas will follow. From 2035, according to the European Commission's timetable, the full integration of unmanned aerial vehicles into controlled airspace with manned aviation should be completed (Wiener Zeitung, 2019).<br/><br/>
The air taxi will not be shuttling crowds of tourists between Schwechat and Stephansplatz in the next ten years. But thanks to the decreasing noise levels of the rotors, which are no longer noticeable in the "normal city noise", as Volocopter claims, they will be used more and more often in one or the other metropolis with millions of inhabitants - possibly also in Vienna (Pramer & Sommavilla, 2020).

The first approved test track in Austria for an unmanned aerial taxi is located in Linz. There is already a functioning prototype in Austria that was developed by *Ehang* in China and built by *FACC* in Innviertel. The aircraft costs around 300,000 euros, weighs 360 kilos, is equipped with 16 electric motors and 16 rotors and is designed to autonomously transport two people. The batteries of the eight-armed drone are sufficient for about 50 kilometres. In its own production line in Ried, 300 units are to be delivered by the end of next year. In order for them to be able to take off in Europe and also for Linz AG, they are working with Austro Control on the "approval regulations", explains FACC board member Robert Machtlinger. 
In Linz it was once again said that this type of passenger transport is seen as a supplement to bus and train. The air enables the fastest connection from A to B in urban areas. However, before the first test route is set up in the Upper Austrian capital for this autonomous transport, 5G mobile radio must first be installed, which was planned for spring 2020 (DerStandard, 2019).

### Relevant initiatives in Austria {-}

- [derbrutkasten.com](https://www.derbrutkasten.com/autonome-lufttaxis-linz-ag-facc-ehang/) 
- [derstandard.at-1](https://www.derstandard.at/story/2000103120464/erste-teststrecke-fuer-e-lufttaxis-2020-in-linz) 
- [derstandard.at-2](https://www.derstandard.at/story/2000122402408/flugtaxis-wann-kommt-der-tesla-der-luefte) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table18, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|It is expected to be too costly for general use|**-**|Equality (*5,10*)|Pramer & Sommavilla, 2020|
|Individual|The flight price is expected to settle in the range of a very expensive taxi|          **-**|Sustainable economic development (*8,11*)|Pramer & Sommavilla, 2020|
|Systemic|Slightly reduced GHG emissions compared to conventional cars from 35 km onwards|**~**|Environmental sustainability (*7,12,13,15*)|Kasliwal et al., 2019|
|Systemic|Increased investment until 2035 and creation of job opportunities|          **+**|Sustainable economic development (*8,11*)|SESAR, 2016|
|Systemic|Stationary areas for safety checks|          **~**|Innovation & Infrastructure (*9*)|Pramer & Sommavilla, 2020|')
```

### Technology and societal readiness level {-}

```{r table19, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-6|1-3|')
```

### Open questions {-}

1. Who will develop the regulations for passenger drones?
2. How much space is needed for take-off and landing, and will this differ between the various providers?
3. What types of routes will be replaced with flying taxis?
4. Will some companies work together and share their technologies?
5. What other areas of application will there be?
6. What will be the thematic priorities for development in the coming years?
7. What name will be agreed internationally for this type of vehicles?


### References {-}
-	DerStandard. (2019, May 14). Teststrecke: Erstes unbemanntes Lufttaxi hebt 2020 in Linz ab - Unternehmen - derStandard.at › Wirtschaft. https://www.derstandard.at/story/2000103120464/erste-teststrecke-fuer-e-lufttaxis-2020-in-linz
-	Holt, K. (2018, October 24). Volocopter will test its autonomous air taxis in Singapore next year | Engadget. https://www.engadget.com/2018-10-24-volocopter-air-taxi-test-singapore-autonomous-drone-helicopter.html
-	J. Hawkins, A. (2018, February 1). Airbus’ autonomous ‘air taxi’ Vahana completes its first test flight - The Verge. https://www.theverge.com/2018/2/1/16961688/airbus-vahana-evtol-first-test-flight
-	Kasliwal, A., Furbush, N. J., Gawron, J. H., McBride, J. R., Wallington, T. J., De Kleine, R. D., Kim, H. C., & Keoleian, G. A. (2019). Role of flying cars in sustainable mobility. Nature Communications, 10(1). https://doi.org/10.1038/s41467-019-09426-0
-	Kellermann, R., Biehle, T., & Fischer, L. (2020). Drones for parcel and passenger transportation: A literature review. Transportation Research Interdisciplinary Perspectives, 4, 100088. https://doi.org/10.1016/j.trip.2019.100088
-	LeBeau, P. (2016, January 23). Boeing’s first test flight of air taxi a success as it works on Uber Air. https://www.cnbc.com/2019/01/23/boeing-takes-step-in-developing-uber-air--with-successful-test-flight.html
-	Pramer, P., & Sommavilla, F. (2020, December 11). Flugtaxis: Wann kommt der Tesla der Lüfte? - PodcastEditionZukunft - derStandard.at › EditionZukunft. https://www.derstandard.at/story/2000122402408/flugtaxis-wann-kommt-der-tesla-der-luefte
-	SESAR. (2016). European Drones Outlook Study. In Single European Sky ATM Research (Issue November). https://doi.org/10.2829/219851
-	UNIQA. (2019, October 24). Lufttaxis: Die Überflieger im Verkehr | UNIQA Österreich | UNIQA Österreich. https://www.uniqa.at/versicherung/mobilitaet/lufttaxis.html
-	Wiener Zeitung. (2019, September 20). Regelwerk für autonome Lufttaxis noch offen - Wiener Zeitung Online. https://www.wienerzeitung.at/nachrichten/wirtschaft/oesterreich/2030182-Fliegen-statt-fahren.html
-	Winter, S. R., Rice, S., & Lamb, T. L. (2020). A prediction model of Consumer’s willingness to fly in autonomous air taxis. Journal of Air Transport Management, 89(August), 101926. https://doi.org/10.1016/j.jairtraman.2020.101926

## Automatic train operations {#automatic_train}

### Synonyms {-}
*high-speed railways (HSR), automatic train control system (ATC), Grade of Automation (GoA), unattended train operation (UTO or GoA4)*

### Definition {-}
Over the past decades, rail transport systems have undergone significant changes in terms of technology level, overall length, travel speed and service qualities. The management of train operations to achieve safe and efficient operation for a railway system has been a long-term concern. In traditional railway systems, this is usually achieved by *(i)* a timetable and a vehicle plan and *(ii)* real-time train operation by drivers using fixed signalling equipment. However, this partially manual work has many disadvantages for train operation due to increasing traffic demand and limited railway infrastructure. Manual driving is usually based on training and experience, which lacks rigorous calculations and systematic considerations. Therefore, it is difficult to ensure safety, quality of service (e.g. transport capacity, punctuality, accuracy of stops) and operating costs (e.g. energy consumption, infrastructure utilisation). This problem is particularly serious in urban rail systems, where passenger demand is extremely high and train departure times are very short (Yin et al., 2017).<br/>

In manual operation, drivers must keep track of the forward signals and then give appropriate train control commands. Essentially, the track logic and dispatchers only know which track circuit the train is on and prevent other trains from entering the same track circuit. Furthermore, as this process is not strictly monitored and rigorously calculated, it is usually susceptible to external factors (e.g. mental state of the driver, extreme weather that can affect driver visibility), leading to uncertainty and inefficiency (Yin et al., 2017). With the development of communication, control and computer technologies in recent decades, automatic train operation (ATO) is considered an emerging technology to replace traditional manual driving in many urban rail systems (Dong et al., 2010; Miyatake & Ko, 2010). Typically, ATO aims to improve the efficiency of railway operations by automatically making real-time decisions on optimised train acceleration, deceleration and braking commands. In the face of increasingly serious environmental and energy concerns, ATO is also widely recognised as a promising approach to reduce energy consumption and carbon emissions through optimised train control decisions while improving service quality (Yin et al., 2017).<br/>

In the railway transportation system, a distinction is made between two basic control loops: the outer control loop as the control of railway traffic and the inner control loop as the train operation itself (Rao, 2015). These two control loops are closely linked and both are important for the safe and efficient operation of a railway system (Yin et al., 2017).

**Rail traffic control**

Rail traffic control aims to monitor the status of traffic and infrastructure, identify deviations and conflicts, and develop a conflict-free plan for rescheduling trains to make supporting decisions for railway dispatchers to optimise the capacity and punctuality of arriving trains and avoid conflicts with other trains (Corman and Meng, 2015). 

**Train operations**

Train operation (i.e. the inner loop) focuses on the safe and efficient train movements in each block at a microscopic level during scheduled (or rescheduled) planning by defining all the basic train control commands (i.e. accelerate, run, coast and brake). 

**For defining the terminology**: ATO is a subsystem of the automatic train control system (ATC). ATC contains three subsystems, namely, automatic train protection (ATP), automatic train operation (ATO) and automatic train supervision (ATS) (Yin et al., 2017).

-	The ATS system is responsible for monitoring train movement by performing the following tasks: Train status monitoring, automatic route selection, automatic schedule generation, automatic operation logging, statistics and reporting, and automatic system status monitoring.
-	The ATP system is a fail-safe system responsible for the safe movement of each train. ATP regulates speed limits not only to maintain a safe operating distance between trains, but also to meet safety and speed requirements. Once the train exceeds the speed limit, ATP automatically applies braking (or emergency braking) to stop the train and ensure safety.
-	The ATO system takes over the on-board functions of a driver to ensure smooth acceleration of the train to the running speed, speed control and precise stopping of the train at the destination platform. In normal situations, the ATO system is responsible for all control commands for traction and braking control of the train and is therefore key to the operational efficiency and profitability of train operation systems (Dong et al., 2010).

Essentially, ATO consists of computer programming and control techniques that support (or completely replace) the driver in the automatic control of train movements under the supervision of ATP and ATS. Typically, train operation is currently realised in two ways: <br/>

*(1)*	Manual driving, supported by a driver advisory system (DAS), which can give the driver a recommended speed to help him with better driving strategies.<br/>

*(2)*	Semi- or fully automatic mode through ATO system in light rail lines with higher level of automation, which can partially (or fully) replace manual driving. In some automated light rail lines, the ATO system first generates a recommended speed profile before the train departs, which allows the train to arrive on time at the next station. After the train has left the station, ATO's speed controller automatically adjusts the control commands for train acceleration, coasting, running or braking through a feedback control loop. Furthermore, ATO is responsible for the automatic precise stopping of the train in the station. Other functions include the automatic reversing of trains at terminals (Yin et al., 2017).<br/>

An ATO system consists of two components (Tasler & Knollmann, 2018):

-	The trackside component ATO-TS (ATO trackside) collects static and dynamic route and timetable data from existing and timetable data from existing trackside TMS (Traffic Management System) and transmits it to the ATO on-board units.
-	The on-board component ATO-OB (ATO onboard) calculates the optimal the optimal driving profile at any time based on the data on infrastructure, route and timetable information, and timetable information and controls the vehicle's traction and braking braking devices of the vehicle for automated driving.

The automation levels of train operating systems are defined by the international standard IEC 62290-1 2014. The five levels of automation range from GoA0 to GoA4. According to this standard, GoA0 and GoA1 are essentially non-automated train operation levels that require manual operation of the trains by the crew in the driver's cab. In GoA2, acceleration and braking are automated, while the driver in the cab is responsible for the safe departure of the train and door control. GoA3 is driverless train operation, where there is no driver in the front cab of the train. Instead, there is only one member of the operating staff responsible for the safe departure of the train. Most of the existing ATO systems usually achieve GoA2 or GoA3. The highest level of train operation automation is unaccompanied train operation (Fully automated [or unattended] train operation (termed as UTO or GoA4), where there is no driver or operations staff at all and trains are operated fully automatically (Yin et al., 2017).<br/>

A wide range of automation functions are already in use on the railways today: electronic interlockings control and secure the routes, automatic train control systems ensure that train movements are secured with the greatest possible track capacity and intelligent operations control systems take over the timely setting of routes and ensure that the existing capacities are better utilised by exploiting all performance reserves and that a high level of performance and quality of rail traffic is guaranteed even in exceptional operational situations. With digitalisation, however, a further increase in performance is possible through the optimisation of complete operational processes in rail traffic according to operational and economic indicators. In the case of underground railways, fully automated rail operations have long been a reality, as their routes in closed tunnel systems make them particularly well suited for driverless operation (Tasler & Knollmann, 2018).<br/>

The expected advantages of ATO in long-distance rail transport are (Tasler & Knollmann, 2018):

-	Increased line and transport capacity by reducing the intervals between trains
-	Improvement of timetable stability and punctuality through uniform and plannable travel times between stations
-	Energy savings through optimised driving
-	Reduction of mechanical stress and wear in the drive and braking system with lower maintenance costs
-	Noise reduction, especially in freight transport, through smooth and even driving smooth driving with fewer braking operations
-	Increased passenger comfort due to smooth and even driving and consistent and consistent ride quality
-	Increased flexibility for demand-responsive train services (with GoA3/4)
-	Improvement of operating costs by increasing the effectiveness of the personnel (for GoA3/4)


### Key stakeholders {-}

- **Affected**: Passengers, Train operators
- **Responsible**: National Governments, City Governments, Private Transport Companies, Public Transport Authorities, Policymakers

### Current state of art in research {-}
A majority of ongoing research is considered to be in private hands, however, recent literature review shows some important trends in rail automation research.<br/>

Firstly, the concept of rail platooning is being developed by Schwerdfeger et al. (2021). It is to enable the coupling of several smaller trains (together with their locomotives) to form a platoon. Radio communication and sensors ensure coordination between the locomotives so that all trains in a platoon brake and accelerate at the same time (Castagnetti et al., 2016). Unlike truck platooning, where the trucks are only connected via wireless vehicle-to-vehicle communication technology, in the rail platooning concept the trains are actually physically connected. In marshalling yards, goods trains are supposed to wait for other trains to be coupled into longer train formations. However, the construction of these fast marshalling yards requires a massive investment in rail infrastructure. Germany's largest rail operator, Deutsche Bahn, has initiated a project to evaluate whether the gains promised by train platooning are worth the huge investment. However, train platooning has the potential to significantly increase train throughput in a congested rail corridor and minimise total train kilometres if the shunting process is fully automated and takes only a few minutes. Therefore, technological developments to realise fast shunting operations are a critical success factor for the rail platooning concept (Schwerdfeger et al., 2021).<br/>

Secondly, the obstacle detection plays an important role in automatic train operation. To overcome the low accuracy and poor real-time performance of conventional detection methods and better detect obstacles over medium and long distances, obstacle detection in rail transport based on Deep Learning was investigated (He et al., 2021).<br/>

Although there is currently no common model or approach aimed at controlling the entire railway network, there is an emerging trend to use integrated optimisation models to simultaneously generate train control actions (e.g. recommended speed profile) and rescheduling. Furthermore, as LTE-R (Long Term Evolution for Railway) technology is emerging, enabling cross-border train-to-ground and train-to-train communication, integrated rail traffic control can also benefit from the future application of this new technology and contribute to the improvement of rail transport systems (Yin et al., 2017).


### Current state of art in practice {-}
Currently, ATO is being deployed on many newly established metro mass transit lines. Almost a quarter of the world's metro systems use at least one line in unattended train operation (UTO) mode, for example in Paris, Beijing, Dubai and Sydney (Fraszczyk & Mulley, 2017) and it has shown its great benefits in reducing manual labour, increasing the transport capacity of the infrastructure and improving the quality of service for passengers (eg. higher punctuality and more precise stops) (Yin et al., 2017). <br/>

However, UTO operation is not yet possible in regional and long-distance traffic. Although research and facilities for the forced braking of trains and the influencing of traction units began as early as the 1930s, it was not until December 1989 that the representatives of all EU states met to decide on the creation of a uniform train control system. The planners were primarily concerned with cross-border traffic. With the uniform European Train Control System (ETCS), an internationally valid standard was to be created that other states and railway companies could then also adopt. So far, the EU Commission has mainly pushed the equipment of the international corridors, which are to be operated with ETCS throughout by the mid-2020s. In addition, new locomotives are to be generally equipped with this system, but for the time being they will also be able to run without the system with conventional signals. The achievements of the ETCS are remarkable and show the direction in which European rail traffic will go in the future. However, the introduction of the complex and technically advanced systems is associated with enormous investments (DB, 2019). <br/>

The train control system (ETCS), the train radio system (GSM-R) and parts of the uniform regulations (Technical Specifications for Interoperability (TSI)) together form the European Rail Traffic Management System (ERTMS) (ÖBB, n.d.). The ETCS enables driving without main and advance signals and increases safety at the same time. The system is comparable to an autopilot, which has been common in aviation for decades. The system monitors the train using information from the route atlas, a precise position determination and predefined command variables and can, thus, make the right decisions in good time to ensure the train's journey even at high speeds. As in the aircraft cockpit, however, the train driver has the final say, he can drive and control the train even if the system fails or other unforeseen events occur. So although, it would technically be possible to operate the train without a driver, in the current state, it is the driver who controls the safety of the systems on the track. He still bears the responsibility for the train and its passengers or its freight. He observes the track and makes sure that there is no damage or obstacles and is the last and most important safety instance in the system. In addition, so far only a few routes are equipped with the technically demanding and price-intensive system. As soon as the train is to continue on other sections of track, the driver is also indispensable (DB, 2019).<br/>

**Ongoing projects**

Together with other partners, *Siemens Mobility* has won two innovative research projects of the German Centre for Rail Transport Research (DZSF) to investigate the safety of automated rail operations at the end of 2020. The aim is to define the necessary criteria for the approval of fully automated regional and long-distance transport. The two projects are funded with 1.7 million euros and have a duration of 30 months. The studies will focus on the highest levels of automation (GoA 3 and GoA 4), which classify fully automated trains with and without attendants. It is expected that automated regional and long-distance rail transport will shape the future of mobility. In Hamburg, Siemens Mobility is conducting a pilot project on highly automated driving with the S-Bahn, which is scheduled to start passenger operations in 2021 for the ITS World Congress. In London, Siemens Mobility has combined a system for automated train operation (ATO) with the European Train Control System (ATCS) for the first time on the ThamesLink line (Global Railway Review, 2020). <br/>

*Alstom* has been awarded an innovation prize for a planned test project to implement Automatic Train Operation (ATO) in the daily operation of regional passenger transport. Alstom will launch this project in 2021 together with the Regional Association "*Großraum Braunschweig*", the German Aerospace Centre (DLR) and the Technical University of Berlin (TU Berlin). It will run for three years. The test trains will run between Braunschweig and Wolfsburg (Sapién, 2020).<br/>

*Nokia* will work with Deutsche Bahn to develop 5G for ATO in 2021 on a 23-km section of Line 21 of the Hamburg S-Bahn from Berliner Tor to Bergedorf.  Line 21 will initially demonstrate driverless shunting of empty trains in the Bergedorf station area based on the transmission of train control information via the 5G network. Siemens will install a similar system in Hamburg (Briginshaw, 2019).<br/>

*EAST Japan Railway (JR East)* has announced plans to start automatic train operation (ATO) on an initial 30km section of the Tokyo Joban commuter line between Ayase and Toride in 2021 (Cuenca, 2021).<br/>

*China Railway Corporation* is developing a train control system that will allow Fuxing high-speed trains to run automatically at speeds of up to 350 km per hour. The automated trains are scheduled to enter service in 2022 ahead of the opening of the Winter Olympics (China Plus, 2019).<br/>

With its smartrail 4.0, the *Swiss railway SBB* is only aiming for the GoA2 automation level. The ATO assistance system developed and tested by SBB is based on the European safety standard ETCS (SBB, n.d.).The train manufacturer Bombardier wants to put the first autonomous trains on the rails in long-distance transport in 2025. On closed lines, self-driving rail vehicles have already existed since the 1980s, for example in suburban and underground trains. On open tracks, however, the demands on the technology are much higher because it has to detect signals and obstacles such as fallen trees or animals without any doubt. However, development is progressing. Oz Ural (head of digital at Bombardier Transportation in Berlin) says: "*Most of the technology is already there and tests are already being carried out. What is still missing for the use of autonomous trains on open tracks is the legal framework and public acceptance.*" (Schwär, 2020)

### Relevant initiatives in Austria {-}
The Digital Automatic Coupling (DAK) is a pioneering innovation in European rail freight transport. It not only enables largely automated coupling for the first time, but also creates the conditions for the automation and digitalisation of rail freight transport in Europe and is, thus, regarded as an important step towards increasing its attractiveness and productivity.

- [DAK](https://presse.oebb.at/de/presseinformationen/20200908-oebb-fokus-auf-digitaler-automatischer-kupplung-im-gueterverkehr)

Currently, about 300 km of ETCS Level 2 are in operation in Austria. In three implementation phases, the ETCS network is to be extended to 3,700 km of ETCS Level 2. From 2025, only railways that can operate with this standard will be allowed to use it.

- [ETCS in Austria](https://infrastruktur.oebb.at/de/geschaeftspartner/schienennetz/dokumente-und-daten/etcs-zugbeeinflussung/etcs-ausbau)

From 2025, driverless metro trains are to run on the new U5 line in Vienna.

- [Driverless trains 1](https://www.wienerzeitung.at/nachrichten/chronik/wien-chronik/2023966-Erster-Blick-auf-fahrerlose-U-Bahn.html)
- [Driverless trains 2](https://www.wienerlinien.at/web/wiener-linien/erste-u-bahn-testfahrt-mit-dem-neuen-x-wagen)
- [Automatic trains](https://www.unsereoebb.at/de/artikel/2019/automatisierter-bahnbetrieb)


### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table102, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Systemic|Noise reduction|**+**|Health & Wellbeing  (*3*)|Tasler & Knollmann, 2018|
|Systemic|Reduced energy consumption|**+**|Environmental sustainability (*7,12,13,15*)|Tasler & Knollmann, 2018; Yin et al., 2017|
|Systemic|Parallel efforts in different countries on the development of ATO|**+**|Innovation & Infrastructure (*9*)|China Plus, 2019; Cuenca, 2021; Briginshaw, 2019|
|Systemic|Many collaboration and joint projects emerge| **+**|Partnership & collaborations (*17*)|Sapien, 2020; Briginshaw, 2019|')
```

### Technology and societal readiness level {-}

```{r table103, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|7-9|')
```

### Open questions {-}
1. How to deal with the social impact of train drivers being able to lose their jobs? 
2. How to increase social acceptance of top level automated trains in Europe?

### References {-}
- Briginshaw, D. (2019, December 12). DB appoints Nokia to develop 5G network for automatic operation | International Railway Journal. https://www.railjournal.com/infrastructure/db-appoints-nokia-to-develop-5g-network-for-automatic-operation/
-	Castagnetti, F., Toubol, A., & Rizzi, G. (2016). C4R Project Increases Rail Capacity without Laying Down New Tracks. Transportation Research Procedia, 14, 672–678. https://doi.org/10.1016/j.trpro.2016.05.329
-	China Plus. (2019, January 4). China developing world’s first 350 km/h automated bullet trains - China Plus. http://chinaplus.cri.cn/news/china/9/20190103/231172.html
-	Cuenca, O. (2021, February 22). JR East to introduce ATO on the Joban Line next month | International Railway Journal. https://www.railjournal.com/regions/asia/jr-east-to-introduce-ato-on-the-joban-line-next-month/
-	DB. (2019, March 19). ETCS: Das Europäische Zugsicherungssystem | DB Inside Bahn. https://inside.bahn.de/etcs-europaeisches-zugsicherheitssystem/
-	Dong, H., Ning, B., Cai, B., & Hou, Z. (2010). Automatic Train Control System Development and Simulation for High-Speed Railways. IEEE Circuits and Systems Magazine, 10(2), 6–18. https://doi.org/10.1109/MCAS.2010.936782
-	Fraszczyk, A., & Mulley, C. (2017). Public Perception of and Attitude to Driverless Train: A Case Study of Sydney, Australia. Urban Rail Transit, 3(2), 100–111. https://doi.org/10.1007/s40864-017-0052-6
-	Global Railway Review. (2020, November 24). Siemens Mobility and partners to study automated rail operations. https://www.globalrailwayreview.com/news/114422/siemens-mobility-partners-automated-rail-operations/
-	He, D., Zou, Z., Chen, Y., Liu, B., Yao, X., & Shan, S. (2021). Obstacle detection of rail transit based on deep learning. Measurement: Journal of the International Measurement Confederation, 176, 109241. https://doi.org/10.1016/j.measurement.2021.109241
-	Miyatake, M., & Ko, H. (2010). Optimization of train speed profile for minimum energy consumption. IEEJ Transactions on Electrical and Electronic Engineering, 5(3), 263–269. https://doi.org/10.1002/tee.20528
-	ÖBB. (n.d.). ETCS Zugbeeinflussung - ÖBB-Infrastruktur AG. Retrieved May 14, 2021, from https://infrastruktur.oebb.at/de/geschaeftspartner/schienennetz/dokumente-und-daten/etcs-zugbeeinflussung
-	Rao, X. (2015). Holistic rail network operation by integration of train automation and traffic management. PhD Thesis, ETH Zurich, 22706.
-	Sapién, J. C. (2020, May 28). Alstom Wins Prize for World-First: ATO Testing for Regional Trains | Railway-News. https://railway-news.com/alstom-wins-german-innovation-prize-ato/
-	SBB. (n.d.). Automatic Train Operation (ATO): die innovative Automatisierung | SBB. Retrieved May 5, 2021, from https://bahninfrastruktur.sbb.ch/de/digitale-bahn/ato.html
-	Schwär, H. (2020, July 24). Autonome Züge ab 2025: Wie sich Bombardier zukunftsfit machen will - Business Insider. https://www.businessinsider.de/wirtschaft/mobility/autonome-zuege-ab-2025-und-schnelles-internet-im-abteil-wie-dieser-30-jaehrige-den-zughersteller-bombardier-fit-fuer-die-zukunft-machen-will/
-	Schwerdfeger, S., Otto, A., & Boysen, N. (2021). Rail platooning: Scheduling trains along a rail corridor with rapid-shunting facilities. European Journal of Operational Research. https://doi.org/10.1016/j.ejor.2021.02.019
-	Tasler, G., & Knollmann, V. (2018). Einführung des hochautomatisierten Fahrens -- auf dem Weg zum vollautomatischen Bahnbetrieb. Signal + Draht, 110(6), 6–14.
-	Yin, J., Tang, T., Yang, L., Xun, J., Huang, Y., & Gao, Z. (2017). Research and development of automatic train operation for railway transportation systems: A survey. In Transportation Research Part C: Emerging Technologies (Vol. 85, pp. 548–572). Elsevier Ltd. https://doi.org/10.1016/j.trc.2017.09.009

