# Collective mobility vehicles {#collective}

## Demand responsive transit {#drt}

### Synonyms {-}
*Micro-public transport, Sustainable Urban Mobility Plans (SUMP), European Structural and Investment Funds (ESIF), Information and communications technology (ICT)*

### Definition {-}
Demand-responsive transit (DRT) systems provide transport services in response to a passenger's request. Modern telecommunication technologies are used to receive requests (via websites, SMS, mobile apps, and less frequently phone calls). These systems are a middle ground between routing systems and taxi services and hardly differ from them in extreme implementations. DRT systems offer flexibility in transport services provided according to demand by using different technologies and organisational principles (Gorev et al., 2020).
DRT has been seen primarily as a solution for people with reduced mobility, but targeting new user groups and combining it with new ICT systems and traffic management tools is an innovative and effective approach. This new form of mobility can bring many benefits to European regions, such as cost savings, less congestion on the roads and emission reductions, while improving the quality of life of citizens. DRT systems are adaptable to almost any region, with numerous parameters that can be changed to suit different regional scenarios (Hunkin & Krell, 2018):

-	DRT systems usually need to be initiated by a public authority. By their very nature, potential user groups are often scattered and fragmented, making them difficult for private operators to identify. Regions should take stock of the performance of their transport systems and consider where public transport is used and at what cost to see where it might be cheaper and more environmentally friendly to use a DRT system.
-	Sustainable Urban Mobility Plans should be developed or modified to include DRT, considering linkages with other transport modes. The process should be overseen by a single transport authority that sets clear low-carbon transport targets to show the long-term direction of travel.
-	The lead partner of DRT initiatives must bring together all stakeholders and manage the process of collaboration. It is the role of the authority to consider the long-term goals and set targets to achieve broader public policy objectives.
-	DRT needs to be made attractive and convenient if it is to have a broad impact. Communication should focus on the multiple benefits of DRT, and ICT should be used as much as possible to effectively integrate services into public transport information systems.
-	The integration of [smart cards](#contactless_cards) and electronic payments can help improve the convenience of the DRT solution, but care must be taken not to alienate users who may not have access to these technologies by avoiding app/mobile-only services.
-	Systems should aim to provide both instant and pre-order services for maximum convenience.
-	There is support for the development and implementation of DRT systems such as ESIFs (European Structural and Investment Funds).

### Key stakeholders {-}

- **Affected**: Public transport users 
- **Responsible**: National governments, Local government, Public transport authorities

### Current state of art in research {-}
There are many variables that have a direct or indirect influence on the performance of DRT services. These factors can be divided into three groups: *Network*, *Operations* and *Demand Characteristics*. Most research uses different approaches to study the impact of these factors on the performance of the service. These approaches can be broadly categorised into two groups: simulations and approximate mathematical methods (Amirgholy & Gonzales, 2016). <br/><br/>
A study on the impact of pricing and service area design on modal shift towards demand responsive transport has shown that a small service area and too low prices can lead to an undesirable shift effect from walking and cycling to DRT. In order to design a DRT concept that leads to a reduction of car trips, the service area should not be limited to the city centre but cover a typical commuter area. Furthermore, the pricing system should have a relatively high minimum charge to make DRT less attractive for short-distance travellers who can also walk or cycle. The simulation experiments have also shown that even with a city-wide DRT coverage area and a minimum charge of 3.00 EUR, without car toll only 11% of DRT users are shifted from car to DRT. With car tolls, the simulation experiments result in a larger share of users shifting from car to DRT. Similar to Kaddoura et al., (2020a) this shows the importance of combining "*pull*" measures (introducing the DRT as a new mode of transport) with "*push*" measures, such as introducing road pricing, increasing parking fees or banning private cars in certain areas. Several simulation experiments show that even with a city-wide DRT and a minimum charge, a significant proportion of DRT users come from public transport. From a user perspective, this is a positive effect as users who switch from public transport to DRT benefit from, for instance, a shorter travel time (Kaddoura et al., 2020b).
Demand responsive transport (DRT) services appear to be more effective in minimising emissions than scheduled services when demand density is not excessive and good service levels are targeted. In particular, demand responsive services perform better in a ring network and the possibility to use smaller vehicles makes them perform better than scheduled services in almost all scenarios studied (Amirgholy & Gonzales, 2016). <br/><br/>
The analysis on the introduction of flexible modes of public transport services recommends the following order of measures for the implementation of the DRT system: 

-	the analysis of the service area in terms of demand for public passenger transport services (population and job density, transport accessibility, condition of the road network, age structure of the population, income level, level of motorisation): 
-	the forecast of passenger transport intensity and identification of areas with low demand; 
-	the analysis of the demand structure, identification of factors that negatively affect the DRT system, such as time-dependent demand (seasonal demand, etc.); 
-	the development of an organisational plan and DRT technology, selection of the service type; 
-	the analysis of the possibility to combine the service with social services for people with reduced mobility; 
-	the calculation of costs;
-	the determination of the possibility to receive subsidies from local authorities; 
-	the development of a detailed service plan, including a business plan and a marketing policy. 

As the DRT system needs significant subsidies from local authorities to cover the significant difference between the costs and the revenues of the transport operator, it is advisable to use combined solutions for existing routes that provide for the use of DRT at certain intervals, which can be beneficial for all entities (transport operators, local authorities, passengers) (Gorev et al., 2020).Finally, the productivity and costs of DRT services of two management practices (time window size setting and a centralised vs. decentralised strategy) is investigated by Quadrifoglio et al., (2008).

### Current state of art in practice {-}
In the last decades, several companies have launched app-based on-demand mobility services (e.g. [UBER](https://www.uber.com/at/en/), [Lyft]( https://www.lyft.com/), [BerlKönig]( https://www.berlkoenig.de/), [CleverShuttle]( https://www.clevershuttle.de/en), [MOIA]( https://www.moia.io/en)). Most of these services combine ride requests of several passengers who then share a ride (pooling, ridesharing). These services form a new category within public transport and are often referred to as Demand Responsive Transit (DRT) or ride-hailing. Most of the existing services are still based on conventional driver-controlled vehicles; only a few services are experimenting with autonomous vehicles (Harris, 2015; Hsu, 2016). A broad market introduction of (shared) autonomous vehicles is expected to reduce operating costs. Lower estimates are in the range of 0.30 to 0.38 EUR per passenger-km (Bösch et al., 2018). Lower operating costs can translate into lower user prices, which can boost demand for such innovative on-demand mobility services (Kaddoura et al., 2020).
The global Demand-Responsive Transit market can be segmented on the basis of transport service, transit service type, booking type and region. In terms of transit service, the Demand-Responsive Transit market can be segmented into bus and taxi. DRT service is generally served by buses. The demand for DRT service for various applications such as school buses, para-transit, medical rides and more is increasing. <br/><br/>
Based on booking type, the Demand-responsive Transit market can be segmented into online booking and offline booking. DRT service offers multiple booking modes such as offline and online booking. Majority of consumers use online booking as it offers higher level of flexibility and real-time information. The online segment is expected to grow rapidly.
In terms of regions, the global demand responsive transit market can be segmented into North America, Europe, Asia Pacific, and Middle East and Africa. Demand-response transit in the United Kingdom and Europe has largely catered to the needs of mobility-impaired passengers since 1997. Europe is likely to dominate the global Demand-Responsive Transit market for the next few years.<br/><br/>
The global market is in its early stages; however, the number of service providers entering the market in North America and Asia-Pacific is increasing (Demand-responsive Transit (DRT) Market Analysis Report with Investment Feasibility & Trends 2026 – FLA News, 2021)
A report by Frost & Sullivan predicts that DRT shuttles will account for 50% of the shared mobility market by 2030. They forecast the DRT market to grow from 2.8 billion US dollars in 2017 to 551.61 billion US dollars and fleet size to increase from 24,100 units to 5.8 million by 2030.<br/><br/>
DRT adoption is being driven by smart city initiatives, government and mobility policies, reduced car ownership and a shift towards multimodal and intermodal transport.
Much of this growth will take place in Europe, currently the largest market for shared mobility, and China. Asia-Pacific, where shuttles are already replacing public transport, is one of the largest DRT markets and will be responsible for over 60% of fleet size growth. The US is also experimenting with different models and it is expected to see strong growth in the medium to long term. However, in other regions such as Latin America, shuttles compete with existing shuttle buses and the potential market in Africa is still very small. 
DRT shuttles combine the affordability of public transport with the convenience of single rides. Accessible transport promotes social inclusion as citizens of all income levels are able to commute to work in a reasonable time.<br/><br/>
The ITF Forum proposes to replace all car trips with demand responsive shuttles.
As DRT shuttles generally complement public transport, Frost & Sullivan recommend public-private partnerships to expand the market and deploy resources effectively. They suggest that asset-intensive services could be operated in collaboration between different service providers or through specially developed technology platforms (Martret, 2019).
The degree of cost recovery from ticket sales of flexible public transport services varies, according to an evaluation of different mobility services in Germany, between 6% and 41% (BMVBS & BBSR, 2009). In rural areas in Austria, the degree of cost recovery is about 15% due to low utilisation rate (VCÖ, 2011). This circumstance requires increased subsidies, which cannot be provided without restrictions due to the need to save money in the public sector (Sammer & Klementschitz, 2010).

### Relevant initiatives in Austria {-}
The term micro-public transport is currently not found in any legal documents in Austria. In the „Kraftfahrliniengesetz“ (KflG), the term Anrufsammeltaxis (§ 38 para. 3 Z 2 KflG) is explained in more detail and many micro-public transport services correspond to this legal definition. In addition, the commercial operation of passenger transport requires a licence under the “Gelegenheitsverkehrsgesetz” (§ 3 para. 1 Z 3 GelverkG). Nevertheless, there are micro public transport services, which cannot be assigned to the KflG nor the GelverkG. For the future development, with regard to legal certainty, more in-depth research work is recommended (Brandl, 2020).

- [bedarfsverkehr.at](https://www.bedarfsverkehr.at/content/Literatur)
- [repositum.tuwien.at](https://repositum.tuwien.at/handle/20.500.12708/1312)
- [projekte.ffg.at](https://projekte.ffg.at/projekt/2929323)


### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table80, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Increases affordability, social inclusion and accessibility of transport.|**+**|Equality (*5,10*)|Martret, 2019|
|Systemic|Potential for cost savings, congestion and emissions reductions|**+**|Health & Wellbeing (*3*)|Hunkin and Krell, 2018|
|Systemic|Only 11% of car users are predicted to shift to DRT|**~**|Environmental sustainability (*7,12,13,15*)|Kaddoura et al., 2020b|
|Systemic|Predicted market growth worldwide; low utilisation rate in Austrain rural areas|**~**|Sustainable economic development (*8,11*))|VCOE, 2011; Martret, 2019|')
```

### Technology and societal readiness level {-}

```{r table81, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|6-8|')
```

### Open questions {-}
1. Which service concepts lead to which modal shift effect?
2. How to deal with volunteering in the context of micro-public (demand responsive) transport ? What framework conditions (legal basis) need to be created?
3. Can DRT be included in the 1-2-3 ticket in Austria, and how to operationalise it?

### Further links {-}

- [maas-alliance.eu](https://maas-alliance.eu/)

### References {-}

- Amirgholy, M. and Gonzales, E. J. (2016) ‘Demand responsive transit systems with time-dependent demand: User equilibrium, system optimum, and management strategy’, Transportation Research Part B: Methodological. Elsevier Ltd, 92, pp. 234–252. doi: 10.1016/j.trb.2015.11.006.
- BMVBS and BBSR (2009) ‘Mobilitätskonzepte zur Sicherung der Daseinsvorsorge in nachfrageschwachen Räumen - Evaluationsreport’, BBSR-Online-Publikation, (10), p. 47. Available at: http://www.bbsr.bund.de/BBSR/DE/Veroeffentlichungen/BBSROnline/2009/DL_ON102009.pdf?__blob=publicationFile&v=2.
- Bösch, P. M. et al. (2018) ‘Cost-based analysis of autonomous mobility services’, Transport Policy, 64(August 2017), pp. 76–91. doi: 10.1016/j.tranpol.2017.09.005.
Brandl, H. (2020) ‘Mobilität – Daseinsgrundfunktion in ländlichen Räumen? Erfolgsfaktoren für den idealtypischen Prozess zur Planung und Implementierung von Mikro-ÖV Angeboten’.
Demand-responsive Transit (DRT) Market Analysis Report with Investment Feasibility & Trends 2026 – FLA News (2021). Available at: https://www.flanewsonline.com/demand-responsive-transit-drt-market-analysis-report-with-investment-feasibility-trends-2026/ (Accessed: 20 March 2021).
- Gorev, A., Popova, O. and Solodkij, A. (2020) ‘Demand-responsive transit systems in areas with low transport demand of “smart city”’, Transportation Research Procedia. Elsevier B.V., 50(2019), pp. 160–166. doi: 10.1016/j.trpro.2020.10.020.
- Harris, M. (2015) Uber Could Be First to Test Completely Driverless Cars in Public - IEEE Spectrum. Available at: https://spectrum.ieee.org/cars-that-think/transportation/self-driving/uber-could-be-first-to-test-completely-driverless-cars-in-public (Accessed: 1 March 2021).
- Hsu, J. (2016) GM and Lyft Team Up for Robot Taxi Service - IEEE Spectrum. Available at: https://spectrum.ieee.org/cars-that-think/transportation/self-driving/gm-and-lyft-team-up-for-robot-taxi-service (Accessed: 1 March 2021).
- Hunkin, S. and Krell, K. (2018) ‘Demand responsive transportation’, Interreg Europe Policy Learning Platform on Low-carbon economy. doi: 10.4324/9781315118321.
- Kaddoura, I., Bischoff, J. and Nagel, K. (2020a) ‘Towards welfare optimal operation of innovative mobility concepts: External cost pricing in a world of shared autonomous vehicles’, Transportation Research Part A: Policy and Practice, 136(February), pp. 48–63. doi: 10.1016/j.tra.2020.03.032.
- Kaddoura, I., Leich, G. and Nagel, K. (2020b) ‘The impact of pricing and service area design on the modal shift towards demand responsive transit’, Procedia Computer Science. Elsevier B.V., 170, pp. 807–812. doi: 10.1016/j.procs.2020.03.152.
- Martret, O. (2019) Demand-Responsive Transit Is the Future of the Mobility Market. Available at: https://shotl.com/news/responsive-transit-is-the-future-of-the-mobility-market (Accessed: 20 March 2021).
- Quadrifoglio, L., Dessouky, M. M. and Ordóñez, F. (2008) ‘A simulation study of demand responsive transit system design’, Transportation Research Part A: Policy and Practice. Pergamon, 42(4), pp. 718–737. doi: 10.1016/j.tra.2008.01.018.
- Sammer, G. and Klementschitz, R. (2010) ‘Finanzierung des öffentlichen Personennahverkehrs in Österreich.’, ÖGZ - Österreichische Gemeindezeitung, pp. 26 – 28.
VCÖ (2011) Erfolgreicher Öffentlicher Verkehr, VCÖ Schriftenreihe ‘Mobilität mit Zukunft’.




## Personal rapid transit {#prt}

### Synonyms {-}
*Automated Transit Networks, ATN, PRT, Podcars*

### Definition {-}
Personal rapid transit is a public transportation option in urban areas where fully automated vehicles provide on-demand, origin-to-destination service over an area network. PRT belongs to a larger family of automated guideway transit (AGT). It is a unique mode of transport that has the following features (Furman et al., 2014; Juster & Schonfeld, 2013): 

-	Direct service from origin to destination
-	Small vehicles that can be used exclusively by an individual or a small group
-	Service based on the needs of the user (on demand rather than fixed schedules)
-	Travelling speed up to 30 mph (48 km/h)
-	Fully automated vehicles
-	Vehicles restricted to a route reserved for their exclusive use, along the guideways
-	Small driveways that are usually elevated at or near ground level 

The components of the personal rapid transit system are:

-	**Software**: integrates all elements of the systems and ensures its flawless functioning 
-	**Hardware**: includes electric and electronic components along the rails and at the stations
-	**Vehicles**: are crucial for passenger experience. In some countries, eg. the USA, PRT vehicles must be suitable for use by disabled people
-	**Stations and guideways**: are the most important elements from the perspective of planners because they influence the urban architecture and invade significantly in the aesthetic of the urban environment. Guideways are physical structures that support and guide an AGT vehicle, consisting of rails and columns. This is the most expensive element of the PRT system
-	**Power sources**: all PRT are powered electrically
-	**Operations and maintenance facilities**: are necessary for control centre, vehicle and equipment storage and maintenance areas

PRT can be considered a more environmentally friendly, lighter and quieter alternative to [light rail transport](#lrt) (LRT) or [bus rapid transport](#brt) (BRT). It also travels slower than typical LRT or BRT, however, it compensates the time by only stopping at the requested stations. This, in turn, saves significant amount of deceleration and acceleration time. The capacity of PRL reaches 10 000 passenger per hour per direction, which is slightly lower compared to LRT and BRT. 

### Key stakeholders {-}

- **Affected**: Citizen, Airport Passengers
- **Responsible**: City government, Private Companies, Transport Planners

### Current state of art in research {-}
Recent research on PRT is scarce, where the majority of studies have been conducted around 2010. They mainly focused on the cost-benefit analysis and comparisons to LRT and BRT (Juster & Schonfeld, 2013). Furthermore, they considered the potential of different integration designs such as construction of guideways above the pavement with and without trees, above the parking lane, in the middle of the road, adjacent to building or integrated into the buildings or tunnels. Special considerations are required due to grade separation which can interfere with trees or cause obstruction at the pavements (Furman et al., 2014). An important aspect is also the aesthetic side of PRT which intrudes significantly into the urban landscape due to elevated rails. Therefore, concerns arise about social acceptability of PRT, where it is expected to affect adversely land use, property values, obstruct light and air as well as create unused spaces underneath the guideways (Staniscia, 2018). The most recent studies focus on the PRT in the context of airport shuttles where this concept has already been proven functional. For example, study by Mobolaji et al., (2021) explored the potential of extension of PRT to connect parking facilities and terminal at Budapest Airport, taking into account user expectations and system architecture.  

### Current state of art in practice {-}
The concept of PRT is more than 50 years old, nevertheless, they are only available in few places around the world. About 30% of the PRT are located at the airports, 30% in the institutional context such as university campus and the final 30% is a mass transit. The examples of existing PRTs are (Furman et al., 2014, p.12):

-	The Morgantown PRT at West Virginia University
-	The Parkshuttle Rivium metro-feeder outside Rotterdam 
-	The Terminal 5 shuttle at London Heathrow Airport 
-	The nature park shuttle in Suncheon Bay, South Korea 
-	The Masdar City PRT in Abu Dhabi 

The reasons for such a limited appearance are scare technical capabilities and scalability of PRT systems along with high capital and operating costs (Furman et al., 2014, p. 36). 


### Relevant initiatives in Austria {-}

- [17th International Conference on Personal Rapid Transit and System Design in 2023](https://waset.org/personal-rapid-transit-and-system-design-conference-in-june-2023-in-vienna) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table67, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Adverse impact on visual aspect of urban spaces|**-**|Health & Wellbeing (*3*)|Staniscia, 2018|
|Systemic|Grade separation increases safety|**+**|Health & Wellbeing (*3*)|Juster & Schonfeld, 2013; Furman et al., 2014|
|Systemic|Electric power used|**+**|Environmental sustainability (*7,12,13,15*)|Juster & Schonfeld, 2013|
|Systemic|Limited adoption worldwide|**-**|Innovation & Infrastructure (*9*)|Furman et al., 2014|')
```

### Technology and societal readiness level {-}

```{r table68, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-8|5-7|')
```

### Open questions {-}
1. More research is needed on actual analysis of costs and benefits associated with the construction of PRT.
2. Can PRT infrastructure be considered and used as a attractive urban furniture? How can art community be engaged in its development?
3. What are the major barriers for more widespread adoption of PRT?

### Further links {-}

- [MIT Report](https://transweb.sjsu.edu/sites/default/files/1227-automated-transit-networks.pdf)

### References {-}
- Furman, B., Fabian, L., Ellis, S., Muller, P., & Swenson, R. (2014). Automated transit networks (ATN): A review of the state of the industry and prospects for the future.
- Juster, R., & Schonfeld, P. (2013). Comparative analysis of personal rapid transit as an urban transportation mode. Transportation research record, 2350(1), 128-135.
- Mobolaji, K., Földes, D., & Csiszár, C. (2021). Concept of Advanced Personal Rapid Transit at Airports. Periodica Polytechnica Civil Engineering, 65(1), 320-334.
- Staniscia, S. (2018). Aesthetic appreciation of Personal Rapid Transit: A new viewpoint. Cities, 79, 169-177.


## Bus rapid transit {#brt}

## Light rail transit {#lrt}

## Passenger drones {#passenger_drones}

### Synonyms {-}
*urban air mobility (UAM), vertical take-off and landing (VTOL), unmanned aerial vehicles (UAVs)*

### Definition {-}
Drones or unmanned aerial vehicles (UAVs) could become the most iconic technology of the 21st century. Drones combine three key principles of technological modernity - computing, autonomy and limitless mobility. Capabilities that until now could only be used by the military are becoming accessible to most of the population. Potential use cases for drones range from surveillance and reconnaissance missions to novel forms of logistics and personal transport. The commercial use of drones is associated with enormous economic opportunities. However, even though drones are already common as surveillance/sensing devices in security services, geodesy or agriculture, their use as a means of transport is still at the beginning (Kellermann et al., 2020). <br/><br/>
Delivery drones are currently able to lift weights of up to 2-3 kg and carry out flight missions in an urban space (Kellermann et al., 2020). However, passenger drones have also already demonstrated the technical ability to transport passengers within or between cities (LeBeau, 2016; Holt, 2018; Hawkins, 2018). It is not only a historic turning point in aviation, but the beginning of a new era in which flat airspace could become the "third dimension" of transport (Kellermann et al., 2020).<br/><br/>
The name of the new type of vehicle is still far from being agreed internationally. There are several names to choose from, such as passenger drone, manned multicopter, Passenger Air Vehicle (PAV), Electric Vertical Take-off and Landing Aircraft (EVTOL), autonomous air taxis, unmanned aerial taxis, flying cars or even a new term (Pramer & Sommavilla, 2020). <br/><br/>
The autonomous air taxis will be a mixture of helicopter and drone. But this also means that they will be vertical take-off and landing (VTOL) vehicles.
There are several reasons why drone-related industries are being supported significantly. One of the reasons is that the airspace is still fairly free of traffic. The risk of collisions is (relatively) low and autopilots for aircraft have long been established. Industry experts therefore suspect that we could see self-flying air taxis even before self-driving cars. Being able to do without pilots would make an air taxi service even cheaper and make it possible for more people to afford it (UNIQA, 2019).<br/><br/>
The European Commission estimates the economic impact at €10 billion annually until 2035 and foresees the creation of more than 100,000 direct jobs. Taking into account indirect macroeconomic effects in drone-related industries, the Commission even projects 250,000 to 400,000 additional jobs (SESAR, 2016).

### Key stakeholders {-}

- **Affected**: Citizen, Insurers
- **Responsible**: National Governments, City government, Private Companies

### Current state of art in research {-}
Since all prototypes are owned by private companies and the technology is not really shared due to competition, there are few technical research papers.
The media analysis about drones for parcel and passenger transportation shows that currently the research is focused in the following areas (see table below):  

```{r table17, echo=FALSE, results='asis'}
cat('| Topics| Percentage|Number of studies|
|:------------:|:-----------:|:------------:|
|General Surveys    |   18.9%  |  21   |
|Logistics (general)  |  18.0%   |  20   |
|Attitude and Acceptance Research    | 13.5%    |  15   |
|Law and Regulations    | 11.7%    |   13  |
|Ethics and Technology Assessment    | 10.8%    |  12   |
|Sustainability Assessment    | 8.1%    |  9   |
|Urban and Transportation Planning    |  7.2%   |  8   |
|Political agenda/strategies    |  6.3%   |   7  |
|Passenger Transportation   |  2.7%   |  3   |
|Humanitarian Logistics    |2.7%     |   3  |') 
```

And although there are currently no autonomous air taxis, there is an initial research examining the factors of consumer willingness to fly in autonomous air taxis. This study identified four factors that turned out to be significantly positive: Familiarity, Value, Fun Factor, Sense of Happiness and two as significantly negative: Aversion to New Technology and Fear (Winter et al., 2020). <br/><br/>
When comparing the GHG emissions of conventional cars with VTOL passenger drones, the passenger drone actually performs a little better than the ordinary car from 35 km onwards. The reasons for this break-even point are, on the one hand, the energy-intensive take-off and landing hover mode and, on the other hand, the enormous efficiency of flying from point A to point B. Given the expected significant time savings over flying, passengers may have an incentive to share VTOL journeys. Therefore, it seems likely that the average occupancy of VTOLs will be greater than that of conventional passenger cars (Kasliwal et al., 2019).

### Current state of art in practice {-}
There are quite a few providers working to develop the unmanned aerial taxis. Munich-based start-up *Lilium* successfully launched a one-and-a-half-ton prototype vertically into the air and hovered in place in early May 2019. It sounds like a helicopter, but it doesn't look like one: The Lilium Jet has wings and 36 all-electric jet motors. That makes it quieter and more energy efficient (and makes it look more futuristic) than a helicopter. Lilium expects to be able to start commercial everyday operation not before 2025.<br/><br/>
But *Lilium* is far from the only manufacturer. *Joby Aviation, Volocopter, AeroMobil, Kittyhawk* and *Zee.Aero* are just a few of the companies hoping to place the most successful model on the market. Already established companies are also taking the concept very seriously. Mobility company *Uber*, aircraft manufacturers *Airbus* and *Boeing*, and car companies such as *Daimler* and *Porsche* are in the race (UNIQA, 2019).
Some manufacturers such as Boeing (LeBeau, 2016), Airbus (Hawkins, 2018) and Volocopter (Holt, 2018) have already conducted the first flight tests of their prototypes.<br/><br/>
The European Union proposes that support services such as flight planning, flight permits and clearances, and dynamic airspace information will be available for drone flight from 2022. From 2027, services such as collision avoidance and capacity management in congested areas will follow. From 2035, according to the European Commission's timetable, the full integration of unmanned aerial vehicles into controlled airspace with manned aviation should be completed (Wiener Zeitung, 2019).<br/><br/>
The air taxi will not be shuttling crowds of tourists between Schwechat and Stephansplatz in the next ten years. But thanks to the decreasing noise levels of the rotors, which are no longer noticeable in the "normal city noise", as Volocopter claims, they will be used more and more often in one or the other metropolis with millions of inhabitants - possibly also in Vienna (Pramer & Sommavilla, 2020).

The first approved test track in Austria for an unmanned aerial taxi is located in Linz. There is already a functioning prototype in Austria that was developed by *Ehang* in China and built by *FACC* in Innviertel. The aircraft costs around 300,000 euros, weighs 360 kilos, is equipped with 16 electric motors and 16 rotors and is designed to autonomously transport two people. The batteries of the eight-armed drone are sufficient for about 50 kilometres. In its own production line in Ried, 300 units are to be delivered by the end of next year. In order for them to be able to take off in Europe and also for Linz AG, they are working with Austro Control on the "approval regulations", explains FACC board member Robert Machtlinger. 
In Linz it was once again said that this type of passenger transport is seen as a supplement to bus and train. The air enables the fastest connection from A to B in urban areas. However, before the first test route is set up in the Upper Austrian capital for this autonomous transport, 5G mobile radio must first be installed, which was planned for spring 2020 (DerStandard, 2019).

### Relevant initiatives in Austria {-}

- [derbrutkasten.com](https://www.derbrutkasten.com/autonome-lufttaxis-linz-ag-facc-ehang/) 
- [derstandard.at-1](https://www.derstandard.at/story/2000103120464/erste-teststrecke-fuer-e-lufttaxis-2020-in-linz) 
- [derstandard.at-2](https://www.derstandard.at/story/2000122402408/flugtaxis-wann-kommt-der-tesla-der-luefte) 

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table18, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|It is expected to be too costly for general use|**-**|Equality (*5,10*)|Pramer & Sommavilla, 2020|
|Individual|The flight price is expected to settle in the range of a very expensive taxi|          **-**|Sustainable economic development (*8,11*)|Pramer & Sommavilla, 2020|
|Systemic|Slightly reduced GHG emissions compared to conventional cars from 35 km onwards|**~**|Environmental sustainability (*7,12,13,15*)|Kasliwal et al., 2019|
|Systemic|Increased investment until 2035 and creation of job opportunities|          **+**|Sustainable economic development (*8,11*)|SESAR, 2016|
|Systemic|Stationary areas for safety checks|          **~**|Innovation & Infrastructure (*9*)|Pramer & Sommavilla, 2020|')
```

### Technology and societal readiness level {-}

```{r table19, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|5-6|1-3|')
```

### Open questions {-}

1. Who will develop the regulations for passenger drones?
2. How much space is needed for take-off and landing, and will this differ between the various providers?
3. What types of routes will be replaced with flying taxis?
4. Will some companies work together and share their technologies?
5. What other areas of application will there be?
6. What will be the thematic priorities for development in the coming years?
7. What name will be agreed internationally for this type of vehicles?


### References {-}

- DerStandard (2019). ‘Teststrecke: Erstes unbemanntes Lufttaxi hebt 2020 in Linz ab - Unternehmen - derStandard.at › Wirtschaft’, 14 May. Available at: https://www.derstandard.at/story/2000103120464/erste-teststrecke-fuer-e-lufttaxis-2020-in-linz (Accessed: 21 January 2021).
- Holt, K. (2018). Volocopter will test its autonomous air taxis in Singapore next year | Engadget. Available at: https://www.engadget.com/2018-10-24-volocopter-air-taxi-test-singapore-autonomous-drone-helicopter.html (Accessed: 21 January 2021).
- J. Hawkins, A. (2018). ‘Airbus’ autonomous “air taxi” Vahana completes its first test flight - The Verge’, 1 February. Available at: https://www.theverge.com/2018/2/1/16961688/airbus-vahana-evtol-first-test-flight (Accessed: 21 January 2021).
- Kasliwal, A. et al. (2019). ‘Role of flying cars in sustainable mobility’, Nature Communications, 10(1). doi: 10.1038/s41467-019-09426-0.
- Kellermann, R., Biehle, T. and Fischer, L. (2020) ‘Drones for parcel and passenger transportation: A literature review’, Transportation Research Interdisciplinary Perspectives. The Author(s), 4, p. 100088. doi: 10.1016/j.trip.2019.100088.
- LeBeau, P. (2016). Boeing’s first test flight of air taxi a success as it works on Uber Air. Available at: https://www.cnbc.com/2019/01/23/boeing-takes-step-in-developing-uber-air--with-successful-test-flight.html (Accessed: 21 January 2021).
- Pramer, P. and Sommavilla, F. (2020) Flugtaxis: Wann kommt der Tesla der Lüfte? - PodcastEditionZukunft - derStandard.at › EditionZukunft. Available at: https://www.derstandard.at/story/2000122402408/flugtaxis-wann-kommt-der-tesla-der-luefte (Accessed: 21 January 2021).
- SESAR (2016). European Drones Outlook Study, Single European Sky ATM Research. doi: 10.2829/219851.
- UNIQA (2019). Lufttaxis: Die Überflieger im Verkehr | UNIQA Österreich | UNIQA Österreich. Available at: https://www.uniqa.at/versicherung/mobilitaet/lufttaxis.html (Accessed: 21 January 2021).
- Wiener Zeitung (2019). ‘Regelwerk für autonome Lufttaxis noch offen - Wiener Zeitung Online’, 20 September. Available at: https://www.wienerzeitung.at/nachrichten/wirtschaft/oesterreich/2030182-Fliegen-statt-fahren.html (Accessed: 21 January 2021).
- Winter, S. R., Rice, S. and Lamb, T. L. (2020). ‘A prediction model of Consumer’s willingness to fly in autonomous air taxis’, Journal of Air Transport Management. Elsevier Ltd, 89(August), p. 101926. doi: 10.1016/j.jairtraman.2020.101926.


## Automatic train operations {#automatic_train}