# Shared mobility {#shared}

## Car sharing

### Synonyms {-}
*Car-Sharing scheme, CSS*

### Definition {-}
In recent years, the growth of car sharing services as a new and more sustainable way of travelling has led to a shift in private mobility from ownership to use of services. The basic idea of car sharing is quite simple: the sharing of a fleet of vehicles by members to make trips on a per-trip basis. Although, the first car sharing scheme for economic reasons dates back to 1948 in the city of Zurich, Switzerland, other attempts at public car sharing schemes in the following years were not successful. Several successful car sharing schemes were launched in the 1980s, with a consolidation in the early 1990s, thanks to an increasing awareness of citizens and a real boom due to a greater diffusion of ICT and mobile services in the 2000s. Car sharing increases the mobility of community members to reach destinations otherwise inaccessible by public transport, walking or cycling, while raising citizens' awareness of the social and environmental impacts of using private cars. It encourages and supports multimodal communities by providing an additional transport option. From the point of view of building a sustainable city, the vehicles used in car sharing are usually fuel efficient and lead to positive effects in reducing urban emissions and urban congestion (Martin and Shaheen, 2011). <br/>

Nowadays, there are different variants of car sharing available on the market. These include:

- Station based

In station-based CarSharing, the cars are parked in fixed parking spaces as close to home as possible. Customers pick up the car there and return it after the journey. Only with this variant, the reservations are possible several days or weeks in advance, but the end time of the booking must also usually be planned in advance. This ensures a high degree of predictability in vehicle availability. Station-based CarSharing is also the cheapest CarSharing variant. The largest providers in Germany (by fleet size) are stadtmobil, cambio, teilAuto and book-n-drive.<br/>

- Free-floating

With free-floating CarSharing, the cars are randomly distributed within a defined business area. Users locate and book them via smartphone. The booking is only possible shortly before the start of the journey and until booking, availability and exact location of the vehicle are uncertain. After the journey, the cars can be parked within the business area. All bookings are open-ended. With this variant, reservations in advance are not possible. Both the availability and the location of the vehicle are therefore difficult to predict. Free-floating, however, allows one-way journeys within the business area. Prices are higher than those of station-based CarSharing. The largest providers in Germany are ShareNow, Sixt share and We share.<br/>

- Combined sharing

Since 2011, combined CarSharing offers were established that offer station-based and free-floating vehicles from a single source. Combined offers in Germany are available, for example, from stadtmobil, book-n-drive, teilAuto and cambio. The prices are usually based on the lower prices of station-based CarSharing.
Free-floating users, on the other hand, largely keep their car. Their motorisation at the time of the study was 485 private cars per 1,000 inhabitants (Bundesverband CarSharing e.V., 2020).

### Key stakeholders {-}

- **Affected**: Citizens
- **Responsible**: Authorities, Municipalities, International lobbyists, Private Companies 

### Current state of art in research {-}
The CarSharing variants have different traffic-reducing effects. The EU research project STARS investigated the traffic-relieving effect of different CarSharing variants under uniform framework conditions. The study shows that many users of station-based and combined CarSharing get rid of private cars shortly before or during CarSharing participation. At the time of the study, the households, therefore, only had a motorisation rate of 108 and 104 cars per 1,000 people in the surveyed households. These values are already below the target of 150 cars per 1,000 people recommended by the Federal Environment Agency Germany for climate and environmentally friendly urban transport in the future.<br/>

The replacement rates in different CarSharing studies from Germany vary. On the one hand, this is due to different survey methods. Only in the studies since 2018 have used a largely uniform survey method in Germany. On the other hand, the latest research has shown that the replacement rate depends strongly on the CarSharing variant studied. For station-based CarSharing and combined CarSharing, there are exclusively positive replacement rates. For pure free-floating CarSharing, both positive and negative replacement rates can be observed. In some cases, fewer private cars were removed by free-floating CarSharing than were put on the road by the CarSharing service. <br/>

According to calculations made by Finanztip together with the ADAC, car sharing is already profitable if the number of kilometres driven per year is less than 10,000, or less than 800 kilometres per month. The costs for a private vehicle with 10,000 annual kilometres driven are identical to the costs that would be incurred for car sharing. Other studies see the limit only at 11,250 kilometres (Hoyer, 2013) or 15,600 kilometres (Seipp, 2014). At 5,000 kilometres per year with one's own medium-sized vehicle, one would save on average between 900 and 1,500 euros per year with a car-sharing provider.
In summary, accrding to Evers (2018) car sharing is profitable, if one: <br/>

-	does not depend on a car every day
-	does not regularly drive longer distances over 100 kilometres
-	drives a total of less than 10,000 kilometres a year

### Current state of art in practice {-}

Europe is currently the most important market for car sharing providers. In 2016, 5.8 million people used the 68,000 carsharing vehicles here. Recently, car manufacturers also started to enter the market directly, such as Daimler, BMW and the FCA Group, which are directly involved in car sharing activities, in order to find new channels to market the cars they produce. The market is growing fast and with this increasing demand comes the need for better understanding and control of the system. In fact, car sharing is not just a matter of business or fleet optimisation, but forms a complex system consisting of different actors, including citizens, authorities and municipalities, businesses. The system becomes complex because of the strong links between the actors as well as the impact on the governance of a city when a large car sharing service is introduced, such as the integration with the existing public transport network and the policies that allow different companies to compete in the same urban area (Ferrero et al., 2018).


### Relevant initiatives in Austria {-}

- [VCÖ](https://www.vcoe.at/presse/presseaussendungen/detail/carsharing-haushalte-potential-2018)
- [ÖBB](https://www.carsharing-wien.com/anbieter/oebb-rail-and-drive)

### Impacts with respect to Sustainable Development Goals (SDGs) {-}

```{r table9, echo=FALSE, results='asis'}
cat('| Impact level| Indicator|Impact direction| Goal description and number|Source|
|:------------:|:-----------:|:------------:|:------------:|:------------:|
|Individual|Uniform access to car in the population|**+**|Equality (*5,10*)|VCOE - Mobilitaet mit Zukunft, 2018|
|Individual|Cost reduced|**+**|Sustainable economic development (*8,11*)|Evers, 2018|
|Systemic|Reduced traffic and improved air quality|**+**|Health & Wellbeing (*3*)|Martin and Shaheen, 2011|
|Systemic|Car-free households are no longer disadvantaged|         **+**|Equality (*5,10*)|VCOE - Mobiliteat mit Zukunft, 2018|
|Systemic|Reduced emissions|          **+**|Environmental sustainability (*7,12-13,15*)|Martin and Shaheen, 2011|
|Systemic|Car sharing fleet grows steadily|    **+** |Innovation & Infrastructure (*9*)|Stadt Wien, Strassenverwaltung und Strassenbau, no date|')
```

### Technology and societal readiness level {-}

```{r table10, echo=FALSE, results='asis'}
cat('|TRL| SRL|
|:----:|:----:|
|7-9|5-7|')
```

### Open questions {-}
1. What is the role of policymakers and minicipalities in supporting car sharing in addressing challenges associated with long-term strategic decisions such as operation area, parking locations or size and type of the fleet, considering specific characteristics of a given city?

### Further links {-}

- [share_now](https://www.share-now.com/at/en/faq/)


### References {-}

- Bundesverband CarSharing (2020) CarSharing in Deutschland 2020.
- Bundesverband CarSharing e.V. (2020) Verkehrsentlastung durch CarSharing - Factsheet.
- Evers, H. (2018) Carsharing günstiger als eigenes Auto? Available at: https://www.computerbild.de/artikel/cb-Tipps-Connected-Car-Kostenvergleich-Ab-wann-lohnt-sich-Carsharing-20041131.html (Accessed: 14 January 2021).
- Ferrero, F. et al. (2018) ‘Car-sharing services: An annotated review’, Sustainable Cities and Society, 37, pp. 501–518. doi: https://doi.org/10.1016/j.scs.2017.09.020.
- Hoyer, N. (2013) Mobilität: Lohnt sich Ihr Auto? Available at: https://www.wiwo.de/technologie/mobilitaet/mobilitaet-lohnt-sich-ihr-auto/8681062-all.html (Accessed: 14 January 2021).
- Martin, E. W. and Shaheen, S. A. (2011) ‘Greenhouse Gas Emission Impacts of Carsharing in North America’, IEEE Transactions on Intelligent Transportation Systems, 12(4), pp. 1074–1086. doi: 10.1109/TITS.2011.2158539.
- Seipp, B. (2014) Carsharing: Für wen sich das geteilte Auto wirklich lohnt - WELT. Available at: https://www.welt.de/motor/article128304929/Fuer-wen-sich-das-geteilte-Auto-wirklich-lohnt.html (Accessed: 14 January 2021).
- Stadt Wien | Straßenverwaltung und Straßenbau (no date) Carsharing in Wien: Nutzung nimmt zu. Available at: https://www.wien.gv.at/verkehr/kfz/carsharing/evaluierung.html (Accessed: 17 January 2021).
- VCÖ - Mobilität mit Zukunft (2018) Mehr als 100.000 Carsharing-Haushalte in Österreich – Potenzial um ein Vielfaches höher - Mobilität mit Zukunft. Available at: https://www.vcoe.at/presse/presseaussendungen/detail/carsharing-haushalte-potential-2018 (Accessed: 18 January 2021).


## Bicycle and e-bicycle hire
### Synonyms {-}
*bike-sharing schemes (BSS), station-based bike-sharing schemes (SBBSS), free-floating bike-sharing schemes (FFBSS)*

### Definition {-}
Bike-sharing schemes have become a key component of urban transport policy over the past decade, as shown by the increase in the number of bicycles recently seen in major cities around the world. The concept of bike-sharing is a service for individuals to get around comfortably by bike without owning one. The bicycle is an energy-efficient, safe, CO~2~-neutral and space-saving means of transport. It has a low environmental footprint (when used). In urban areas it is a good alternative to the car for short journeys. For longer journeys or for getting to work in an urban environment, it is an excellent complement to public transport. Although at the beginning of the 21st century most BSSs were docked, today's BSSs consist of both docked and dockless BSSs, which have recently emerged in several cities such as London, New York, San Francisco, Beijing and many others (El Arbi and Stephane, 2020).<br/><br/>
Modern urban short-term bicycle rental systems or public BSSs offer 24-hour access to bicycles, can be picked up and returned at self-service docking stations, and are distributed throughout the city (Midgley, 2011). As technology advances, global positioning systems (GPS) allow operators to track the bikes and reposition them if necessary, while user registration and credit card identification reduce anonymity and theft.<br/><br/>
The development of, mainly European, BSSs has typically been categorized into four different generations, which in some properties merge into each other. The use of the 1965 "White Bikes" in Amsterdam was possible without personal registration and the bikes could be found all over the city without fixed stations. The model collapsed within a few days due to vandalism and theft. The second generation used locks and heavy bikes. Vandalism rates decreased, but bikes were still stolen because of the anonymity of the customers (DeMaio, 2009). The third generation became smarter and more attractive through technological improvements such as automated smart cards, electronic bike locks and payment systems. Users received a code via SMS to unlock the bikes. 
The current fourth generation, could include movable solar-powered docking stations, GPS-based real-time availability apps on mobile phones and more electric bikes (Zademach and Musch, 2018).<br/><br/>
In Austria, there are multiple companies currently offering bicycle or e-bicycle sharing services such as *Citybike, Ofo or oBike*. Nevertheless, about a year after their launch, the Asian providers of stationless rental bikes in Vienna (start-ups *Ofo* and *oBike*) are in retreat from the federal capital (Rachbauer, 2018b). This has been a result of strict rules announced stationless rental bikes in Vienna on 1st August 2018. A failure to comply with the rules results in the bikes being removed for a fee. Previously, the traffic road act (*StVO*) allowed the placement of rental bikes in public spaces and the city lacked the means of action against the providers. With the help of a so-called local police ordinance, the city hall requires that the bike-sharing companies pick up objectionable bikes on weekdays within four hours and at night and on weekends within twelve hours after notification. If they do not comply, the bikes are removed for a fee. In addition, administrative fines of up to 700 euros are possible.
Moreover, a maximum number of 1500 bicycles per provider is also set. The bikes are given a number and each one has to be registered with the city. According to Blum, setting up several companies to circumvent the upper limit is not allowed. The rental companies themselves also have to meet some criteria now. In particular, there must be a company headquarters in Vienna and a service hotline. *Ofo* and *oBike* already meet these requirements (Rachbauer, 2018d). <br/><br/>
Across Austria the situation with rental bikes differs significantly, for example in Innsbruck, renatal bikes were first introduced in 2014 and and are succesfully opearting. Meanwhile, in Linz rental bike system has been approved in 2017 and the docking stations are currently under construction. Further, in Salzburg the S-Bike rental system is operating and linked to federal funding (Affenzeller, 2020, Citybike Salzburg, no date a). On the other hand in Graz there is no uniform bike rental system and these services are provided by shops and hotels (Rachbauer, 2016).

### Key stakeholders {-}

- **Affected**: Mobile citizen, pedestrians, 
- **Responsible**: National Governments, International lobbyists, (Public) Transport Agency, Non-Profit-Organizations, Private For-Profit Companies, Private Companies (e. g. outdoor advertising companies)

### Current state of art in research {-}


### Current state of art in practice {-}




### Relevant initiatives in Austria {-}

- [citybikewien](https://www.citybikewien.at/de)
- [citybikesalzburg](http://www.citybikesalzburg.at/)
- [nextbike](https://www.nextbike.at/de/niederoesterreich/)
- [tpis.at](https://www.tips.at/nachrichten/linz/land-leute/523512-linzer-radverleih-startet-im-fruehjahr-an-40-standorten)

### Impacts with respect to Sustainable Development Goals (SDGs) {-}



### Technology and societal readiness level {-}


### Open questions {-}


### Further links {-}




### References {-}
- Affenzeller, J. (2020) Linzer Radverleih startet im Frühjahr an 40 Standorten. Available at: https://www.tips.at/nachrichten/linz/land-leute/523512-linzer-radverleih-startet-im-fruehjahr-an-40-standorten (Accessed: 18 January 2021).
- El Arbi, A. A. and Stephane, C. K. T. (2020) ‘Intelligent Management of Bike Sharing in Smart Cities using Machine Learning and Internet of Things’, Sustainable Cities and Society. Elsevier B.V., p. 135907. doi: 10.1016/j.scs.2020.102702.
- Campbell, A. A. et al. (2016) ‘Factors influencing the choice of shared bicycles and shared electric bikes in Beijing’, Transportation Research Part C: Emerging Technologies. Elsevier Ltd, 67, pp. 399–414. doi: 10.1016/j.trc.2016.03.004.
- Cheng, L. et al. (2020) ‘How could the station-based bike sharing system and the free-floating bike sharing system be coordinated?’, Journal of Transport Geography. Elsevier Ltd, 89(March 2019), p. 102896. doi: 10.1016/j.jtrangeo.2020.102896.
- Cheng, X. and Gao, Y. (2018) ‘The Optimal Monthly Strategy Pricing of Free-Floating Bike Sharing Platform’, Modern Economy. Scientific Research Publishing, Inc, 09(02), pp. 318–338. doi: 10.4236/me.2018.92021.
- Citybike Salzburg (no date a) Citybike Salzburg. Available at: http://www.citybikesalzburg.at/hanuschplatz.php (Accessed: 12 January 2021).
- Citybike Salzburg (no date b) Citybike Salzburg - Tarife. Available at: http://www.citybikesalzburg.at/tarife.php (Accessed: 18 January 2021).
- Citybike Wien (2019) 10 Millionen Fahrten bei Citybike Wien! Available at: https://www.citybikewien.at/de/news/595-neuer-meilenstein-bei-citybike-wien-erreicht (Accessed: 12 January 2021).
- Citybike Wien (no date) Tarife - Citybike Wien. Available at: https://www.citybikewien.at/de/tarife (Accessed: 18 January 2021).
- DeMaio, P. (2009) ‘Bike-sharing: History, Impacts, Models of Provision, and Future’, Journal of Public Transportation. University of South Florida Libraries, 12(4), pp. 41–56. doi: 10.5038/2375-0901.12.4.3.
- Eillie, A. (2016) Oxford Test Drives Peer-to-Peer Bike Sharing. Available at: https://www.bloomberg.com/news/articles/2016-09-13/cycle-land-is-a-new-peer-to-peer-bike-sharing-platform (Accessed: 8 January 2021).
- Fishman, E. et al. (2014) ‘Barriers to bikesharing: An analysis from Melbourne and Brisbane’, Journal of Transport Geography. Elsevier Ltd, 41, pp. 325–337. doi: 10.1016/j.jtrangeo.2014.08.005.
- Fishman, E. (2016) ‘Bikeshare: A Review of Recent Literature’, Transport Reviews. Routledge, 36(1), pp. 92–113. doi: 10.1080/01441647.2015.1033036.
- der Grazer (2019) 100 Millionen Euro für Fahrrad-Offensive im Großraum Graz – Der Grazer. Available at: https://grazer.at/de/uHAnk5t2/100-millionen-euro-fuer-fahrrad-offensive-im-graz/ (Accessed: 18 January 2021).
- Gu, T., Kim, I. and Currie, G. (2019) ‘To be or not to be dockless: Empirical analysis of dockless bikeshare development in China’, Transportation Research Part A: Policy and Practice. Elsevier Ltd, 119, pp. 122–147. doi: 10.1016/j.tra.2018.11.007.
- Hillebrand, T. (2019) Mietradsystem ‘Stadtrad Innsbruck’ - VCÖ Vorbildhafte Mobilitätsprojekte. Available at: https://mobilitaetsprojekte.vcoe.at/mietradsystem-stadtrad-innsbruck-2019 (Accessed: 18 January 2021).
- Li, H. et al. (2019) ‘Effects of dockless bike-sharing systems on the usage of the London Cycle Hire’, Transportation Research Part A: Policy and Practice. Elsevier Ltd, 130, pp. 398–411. doi: 10.1016/j.tra.2019.09.050.
- Ma, X. et al. (2020) ‘A comparison in travel patterns and determinants of user demand between docked and dockless bike-sharing systems using multi-sourced data’, Transportation Research Part A: Policy and Practice. Elsevier, 139(June), pp. 148–173. doi: 10.1016/j.tra.2020.06.022.
- Midgley, P. (2011) ‘Bicycle-Sharing Schemes: Enhancing Sustainable Mobility in Urban Areas’, Commission on Sustainable Development, Nine teent(8), p. 24. Available at: http://www.un.org/esa/dsd/resources/res_pdfs/csd-19/Background-Paper8-P.Midgley-Bicycle.pdf.
- Nextbike Niederösterreich (no date) Nextbike Niederösterreich - Tarife. Available at: https://www.nextbike.at/de/niederoesterreich/preise/ (Accessed: 18 January 2021).
- Rachbauer, S. (2016) Bikesharing in Österreich: Leihräder auf der Überholspur | kurier.at. Available at: https://kurier.at/chronik/oesterreich/bikesharing-in-oesterreich-leihraeder-auf-der-ueberholspur/400067369 (Accessed: 12 January 2021).
- Rachbauer, S. (2018a) Bikesharing in Österreich: Leihräder auf der Überholspur. Available at: https://kurier.at/chronik/oesterreich/bikesharing-in-oesterreich-leihraeder-auf-der-ueberholspur/400067369.
- Rachbauer, S. (2018b) Leihräder: Wiener wollen es aufgeräumt. Available at: https://kurier.at/chronik/wien/leihraeder-wiener-wollen-es-aufgeraeumt/400067357 (Accessed: 12 January 2021).
- Rachbauer, S. (2018c) Leihräder: Wiener wollen es aufgeräumt. Available at: https://kurier.at/chronik/wien/leihraeder-wiener-wollen-es-aufgeraeumt/400067357.
- Rachbauer, S. (2018d) Wien führt strenge Regeln für stationslose Leihräder ein. Available at: https://kurier.at/chronik/wien/wien-fuehrt-strenge-regeln-fuer-stationslose-leihraeder-ein/312.944.617 (Accessed: 12 January 2021).
- Schöffel, R., Zierer, M. and Kühne, S. (2017) Nutzerdaten offen im Netz: BR deckt Datenleck beim Fahrradverleiher Obike auf. Available at: https://www.br.de/nachricht/datenleck-obike-100.html (Accessed: 8 January 2021).
- Sun, S. and Ertz, M. (2021) ‘Contribution of bike-sharing to urban resource conservation: The case of free-floating bike-sharing’, Journal of Cleaner Production. Elsevier Ltd, 280, p. 124416. doi: 10.1016/j.jclepro.2020.124416.
- Tech & Nature (2020) Niederösterreich modernisiert Sharing-Bike-Flotte ‘nextbike’ - Tech & Nature. Available at: https://www.techandnature.com/niederosterreich-modernisiert-sharing-bike-flotte-nextbike/ (Accessed: 18 January 2021).
- Zademach, H. M. and Musch, A. K. (2018) ‘Bicycle-sharing systems in an alternative/diverse economy perspective: a sympathetic critique’, Local Environment. Taylor & Francis, 23(7), pp. 734–746. doi: 10.1080/13549839.2018.1434494.
- Zhang, Y., Lin, D. and Mi, Z. (2019) ‘Electric fence planning for dockless bike-sharing services’, Journal of Cleaner Production, 206, pp. 383–393. doi: 10.1016/j.jclepro.2018.09.215.



## E-scooter hire

## Ride-hailing